### 2020-03-18 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
remove older files to merge with new laptop files

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F)
```

### Load data and packages
```{r}
library(randomForest)
library(pdp)
library(caret)
library(tidyverse)
library(ggpubr)
library(caret)
library(stringr)
library(here)
library(janitor)
library(ggfortify)

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  select('drainage_assoication', 'ridge', 'mortality', 
         'prefire','mortality_qual', 'canopy_loss_success', 
         'canopy_loss_failure', 'postfire_canopy', 'postfire', 
         'Change_in_canopy_success', 'Change_in_canopy_failure', 
         'Change_Qualitative','high_mortality', 'area_scaled', 
         'dem_10m_zaca', 'flow_accumulation', 'sr', 'sr_new', 
         'rain', 'slope', 'tmax', 'tmean', 'tmin', 'TPI_five', 
         'TPI_queen', 'TWI', 'fri', 'dem_scaled', 'southness', "cwd", "vpd_jan", "vpd_aug" ) 
 #mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High")))
 # keep getting error (Outer names are only allowed for unnamed scalar atomic inputs) when order the levels, but the reordering works anyway (although it doesnt work with the ssame error in the binomial code. The code directly below works to reorder the levels without error)
mcf$mortality_qual  <- factor(mcf$mortality_qual, levels = c("None", "Low", "Moderate", "High"))

```

### Sample Size for each mortality class
```{r}
mortality_count <- mcf %>% 
  group_by(mortality_qual) %>% 
  count()

```

### Data Spread

```{r}
solar_rad_mcf <- ggplot(mcf, aes(x=mortality_qual, y=sr, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="Solar Radiation") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

tmax_mcf <- ggplot(mcf, aes(x=mortality_qual, y=tmax, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="Max Aug Temp (celcius)") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

rain_mcf <- ggplot(mcf, aes(x=mortality_qual, y=rain, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="Precipitation") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

tpi_mcf <- ggplot(mcf, aes(x=mortality_qual, y=TPI_five, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="TPI") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

twi_mcf <- ggplot(mcf, aes(x=mortality_qual, y=TWI, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="TWI") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

slope_mcf <- ggplot(mcf, aes(x=mortality_qual, y=slope, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="Slope") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

elev_mcf <- ggplot(mcf, aes(x=mortality_qual, y=dem_10m_zaca, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="Elevation") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

cwd_mcf <- ggplot(mcf, aes(x=mortality_qual, y=cwd, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="CWD") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

aug_vpd_mcf <- ggplot(mcf, aes(x=mortality_qual, y=vpd_aug, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="Aug VPD") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

jan_vpd_mcf <- ggplot(mcf, aes(x=mortality_qual, y=vpd_jan, fill=mortality_qual))+
  geom_violin(alpha=0.5, show.legend=F) +
  theme_classic() +
  labs(x="Mortality", y="Jan VPD") +
  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
  coord_flip()

ggarrange(solar_rad_mcf, rain_mcf, slope_mcf, elev_mcf, tmax_mcf, tpi_mcf, twi_mcf, cwd_mcf, aug_vpd_mcf, jan_vpd_mcf)
#ggsave("zaca MCF most important variable violin plots.png")
```

Most distinction between classes in slope, elevation, and solar radiation. Could explain why those variables tend to have the highest variable importance. 
Interestingly, stands with higher elevation had higher mortality and stands with steeper slope and higher max temp had lower/no mortality. Both of which are counter-intuitive.

### PCA

```{r}
#sub-data
mcf_sub <- mcf %>% 
  select (dem_10m_zaca:sr, rain:tmax, TPI_five, TWI, cwd:vpd_jan)

mcf_sub_pca <- prcomp(mcf_sub, scale = T)

#summary
summary(mcf_sub_pca)
mcf_sub_pca

#plot
mcf_sub_pca_plot <- autoplot(mcf_sub_pca, 
                          colour = NA, # HAVE to spell it color this way (colour) 
                          loadings.label = T,
                          loadings.label.size = 3,
                          loadings.label.colour = "black",
                          loadings.label.repel = T) +
  scale_y_continuous(lim=c(-0.1, 0.15)) +
  labs(title = "Zaca Fire MCF PCA")

mcf_sub_pca_plot

```

### Random Forest

Which variables contribute to mortality?

Literature review:

```{r}
# model parameters---------------------------------------------

# set sample size 
nsample = 125

# lists to save outputs from for loop
test_accuracy_list <- list()

```

```{r}

#random forest code---------------------------------------------------
for (i in 1:500)
{

  # divide data into equal # of observations --------------------

  mcf_high <- mcf[which(mcf$mortality_qual == "High"),]
  mcf_low <- mcf[which(mcf$mortality_qual == "Low"),]
  mcf_moderate <- mcf[which(mcf$mortality_qual == "Moderate"),]
  mcf_none <- mcf[which(mcf$mortality_qual == "None"),]

  # pull random observations from the high class and separate
  x_high <- sample(1:nrow(mcf_high), nsample , replace = FALSE)
  polygon_data_high_train <- mcf_high[x_high,]
  polygon_data_high_test <- mcf_high[-x_high,]
  # pull random observations from the low class and separate
  x_low <- sample(1:nrow(mcf_low), nsample , replace = FALSE)
  polygon_data_low_train <- mcf_low[x_low,]
  polygon_data_low_test <- mcf_low[-x_low,]
  # pull random observations from the moderate class and separate
  x_moderate <- sample(1:nrow(mcf_moderate), nsample , replace = FALSE)
  polygon_data_moderate_train <- mcf_moderate[x_moderate,]
  polygon_data_moderate_test <- mcf_moderate[-x_moderate,]
  # pull random observations from the none class and separate
  x_none <- sample(1:nrow(mcf_none), nsample , replace = FALSE)
  polygon_data_none_train <- mcf_none[x_none,]
  polygon_data_none_test <- mcf_none[-x_none,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_none_train, polygon_data_low_train, polygon_data_moderate_train, polygon_data_high_train))
  # recombine classes that were not pulled
  test <- do.call("rbind", list(polygon_data_none_test, polygon_data_low_test, polygon_data_moderate_test, polygon_data_high_test))

  #random forest model
  rfm<-randomForest(mortality_qual~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI, data = train, ntree=500, importance=T, nscale=0)

  #plot to visualize variable importance
  #varImpPlot(rfm)
  model_importance <- importance(rfm)

  #predcit test data
  pp<-predict(rfm, test)

  #test data accuracy
  tabb<- table(pp, test$mortality_qual)
  acc<-1-sum(diag(tabb))/sum(tabb) #% missclassified 
  test_accuracy <-(1-acc)*100 #model accuracy

  # lists to save model outputs
  test_accuracy_list <- append(test_accuracy_list, test_accuracy)
  write.table(model_importance, file="rfm model importance.csv", sep=",",col.names=NA, append=T) 
  write.table(tabb, file = "mcf test accuracy table.csv", sep=",",col.names=NA, append=T)
}

```

Note: when running the model, have to remember to delete the rfm model importance.csv and mcf test accuracy table.csv files otherwise R will keep adding the values onto the orignial file and not make a new one for each model succession. I've been doing that or renaming the file, but maybe theres an easier method

#### Model Results
```{r}

# overall average model accuracy -----------------------------------------------------

accuracy_df <- as.data.frame(unlist(test_accuracy_list))
mean_test_accuracy_1 <- accuracy_df %>% 
  mutate(test_accuracy_list = as.numeric(test_accuracy_list)) %>% 
  summarize(mean_accuracy = mean(unlist(test_accuracy_list)))

#average model importance ---------------------------------------------------------

model_importance_tidy <- read_csv(here("Model Data Sheets and Code", "rfm model importance.csv")) %>% 
  clean_names() %>% 
  filter(none != "None") %>% 
  rename (variables = x1) %>% 
  mutate (variables = str_to_lower(variables)) %>% 
  mutate (none = as.numeric(none),
          low = as.numeric(low),
          moderate = as.numeric(moderate),
          high = as.numeric(high),
          mean_decrease_accuracy = as.numeric(mean_decrease_accuracy),
          mean_decrease_gini = as.numeric(mean_decrease_gini)
          )

# model importance summary statistics
model_importance_summary <- model_importance_tidy %>% 
  group_by(variables) %>% 
  summarize (none_mean = mean(none),
             low_mean = mean(low),
             moderate_mean = mean(moderate),
             high_mean = mean(high),
             mean_dec_accuracy_mean = mean(mean_decrease_accuracy),
             mean_decrease_gini_mean = mean(mean_decrease_gini)
  )

# average class accuracy ---------------------------------------------------------

high <- read_csv("mcf test accuracy table.csv")
moderate <- read_csv("mcf test accuracy table.csv") 
low <- read_csv("mcf test accuracy table.csv")
none <- read_csv("mcf test accuracy table.csv")

high_mean_acc <- high %>% 
  select (X1, High) %>% 
  drop_na() %>% 
  filter (X1 == "High") %>% 
  mutate (High = as.numeric(High),
          class_acc = High / 33) %>% 
  summarise(mean = mean(class_acc))

moderate_mean_acc <- moderate %>% 
  select (X1, Moderate) %>% 
  drop_na() %>% 
  filter (X1 == "Moderate") %>%
  mutate (Moderate = as.numeric(Moderate),
          class_acc = Moderate / 35)%>% 
  summarise(mean = mean(class_acc))

low_mean_acc <- low %>% 
  select (X1, Low) %>% 
  drop_na() %>% 
  filter (X1 == "Low") %>%
  mutate (Low = as.numeric(Low),
          class_acc = Low / 41) %>% 
  summarise(mean = mean(class_acc))

none_mean_acc <- none %>% 
  select (X1, None) %>% 
  drop_na() %>% 
  filter (X1 == "None") %>%
  mutate (Low = as.numeric(None),
          class_acc = Low / 126) %>% 
  summarise(mean = mean(class_acc))

```

#### Partial Dependence Plots
```{r}

```

### Random Forest Model with 3 mortality classes

No and low mortality are pretty similar and both would be ideal for land managers.

```{r}
# model parameters and sub data -----------------------------------------------------------

# sub data 
mcf_3 <- mcf %>%
  mutate(three_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% "Moderate" ~ "Moderate",
    mortality_qual %in% c("Low", "None") ~ "Low"
  )) %>% 
  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
                                     levels=c("Low", "Moderate", "High")))

# set sample size 
nsample1 = 125

# lists to save outputs from for loop
test_accuracy_list <- list()
```

```{r}

#random forest code---------------------------------------------------
for (i in 1:500)
{

  # divide data into equal # of observations --------------------

  mcf_high <- mcf_3[which(mcf_3$three_class_mortality == "High"),]
  mcf_low <- mcf_3[which(mcf_3$three_class_mortality == "Low"),]
  mcf_moderate <- mcf_3[which(mcf_3$three_class_mortality == "Moderate"),]

  # pull random observations from the high class and separate
  x_high <- sample(1:nrow(mcf_high), nsample1 , replace = FALSE)
  polygon_data_high_train <- mcf_high[x_high,]
  polygon_data_high_test <- mcf_high[-x_high,]
  # pull random observations from the low class and separate
  x_low <- sample(1:nrow(mcf_low), nsample1 , replace = FALSE)
  polygon_data_low_train <- mcf_low[x_low,]
  polygon_data_low_test <- mcf_low[-x_low,]
  # pull random observations from the moderate class and separate
  x_moderate <- sample(1:nrow(mcf_moderate), nsample1 , replace = FALSE)
  polygon_data_moderate_train <- mcf_moderate[x_moderate,]
  polygon_data_moderate_test <- mcf_moderate[-x_moderate,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_low_train, polygon_data_moderate_train, polygon_data_high_train))
  # recombine classes that were not pulled
  test <- do.call("rbind", list(polygon_data_low_test, polygon_data_moderate_test, polygon_data_high_test))

  #random forest model
  rfm<-randomForest(three_class_mortality~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI, data = train, ntree=500, importance=T, nscale=0)

  #plot to visualize variable importance
  model_importance <- importance(rfm)

  #varImpPlot(rfm)

  #predcit test data
  pp<-predict(rfm, test)

  #test data accuracy
  tabb<- table(pp, test$three_class_mortality)
  acc<-1-sum(diag(tabb))/sum(tabb) #% missclassified
  test_accuracy <-(1-acc)*100 #model accuracy

  # lists to save model accuracys
  test_accuracy_list <- append(test_accuracy_list, test_accuracy)
  write.table(model_importance, file="rfm model importance_3 class mortality.csv", sep=",",col.names=NA, append=T) 
  write.table(tabb, file = "mcf test accuracy table_3 class mortality.csv", sep=",",col.names=NA, append=T)

}

```

#### Model Results
```{r}

# overall average model accuracy -----------------------------------------------------

accuracy_df <- as.data.frame(unlist(test_accuracy_list))
mean_test_accuracy_1 <- accuracy_df %>% 
  summarize(mean_accuracy = mean(unlist(test_accuracy_list)))

#average model importance ---------------------------------------------------------

model_importance_tidy <- read_csv(here("Model Data Sheets and Code", "rfm model importance_3 class mortality.csv")) %>% 
  clean_names() %>% 
  filter(low != "Low") %>% 
  rename (variables = x1) %>% 
  mutate (variables = str_to_lower(variables)) %>% 
  mutate (low = as.numeric(low),
          moderate = as.numeric(moderate),
          high = as.numeric(high),
          mean_decrease_accuracy = as.numeric(mean_decrease_accuracy),
          mean_decrease_gini = as.numeric(mean_decrease_gini)
          )

# model importance summary statistics
model_importance_summary <- model_importance_tidy %>% 
  group_by(variables) %>% 
  summarize (low_mean = mean(low),
             moderate_mean = mean(moderate),
             high_mean = mean(high),
             mean_dec_accuracy_mean = mean(mean_decrease_accuracy),
             mean_decrease_gini_mean = mean(mean_decrease_gini)
  )

# average class accuracy ---------------------------------------------------------

high <- read_csv("mcf test accuracy table_3 class mortality.csv") 
moderate <- read_csv("mcf test accuracy table_3 class mortality.csv")
low <- read_csv("mcf test accuracy table_3 class mortality.csv")

high_mean_acc <- high %>% 
  select (X1, High) %>% 
  drop_na() %>% 
  filter (X1 == "High") %>% 
  mutate (High = as.numeric(High),
          class_acc = High / 33) %>% 
  summarise(mean = mean(class_acc))

moderate_mean_acc <- moderate %>% 
  select (X1, Moderate) %>% 
  drop_na() %>% 
  filter (X1 == "Moderate") %>%
  mutate (Moderate = as.numeric(Moderate),
          class_acc = Moderate / 35)%>% 
  summarise(mean = mean(class_acc))

low_mean_acc <- low %>% 
  select (X1, Low) %>% 
  drop_na() %>% 
  filter (X1 == "Low") %>%
  mutate (Low = as.numeric(Low),
          class_acc = Low / 292) %>% 
  summarise(mean = mean(class_acc))

```

### Random Forest via Cross Validation

code from:  https://www.youtube.com/watch?v=84JSk36og34

```{r}
# model parameters---------------------------------------------

# set sample size 
nsample2 = 125
```

```{r}

# equal number of observations
  mcf_high <- mcf[which(mcf$mortality_qual == "High"),]
  mcf_low <- mcf[which(mcf$mortality_qual == "Low"),]
  mcf_moderate <- mcf[which(mcf$mortality_qual == "Moderate"),]
  mcf_none <- mcf[which(mcf$mortality_qual == "None"),]

  # pull random observations from the high class and separate
  x_high <- sample(1:nrow(mcf_high), nsample2 , replace = FALSE)
  polygon_data_high_train <- mcf_high[x_high,]
  polygon_data_high_test <- mcf_high[-x_high,]
  # pull random observations from the low class and separate
  x_low <- sample(1:nrow(mcf_low), nsample2 , replace = FALSE)
  polygon_data_low_train <- mcf_low[x_low,]
  polygon_data_low_test <- mcf_low[-x_low,]
  # pull random observations from the moderate class and separate
  x_moderate <- sample(1:nrow(mcf_moderate), nsample2 , replace = FALSE)
  polygon_data_moderate_train <- mcf_moderate[x_moderate,]
  polygon_data_moderate_test <- mcf_moderate[-x_moderate,]
  # pull random observations from the none class and separate
  x_none <- sample(1:nrow(mcf_none), nsample2 , replace = FALSE)
  polygon_data_none_train <- mcf_none[x_none,]
  polygon_data_none_test <- mcf_none[-x_none,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_none_train, polygon_data_low_train, polygon_data_moderate_train, polygon_data_high_train))
  # recombine classes that were not pulled
  test <- do.call("rbind", list(polygon_data_none_test, polygon_data_low_test, polygon_data_moderate_test, polygon_data_high_test))
# make sure y variable is recgonized as a factor
mortality_qual <- as.factor(mcf$mortality_qual)

#add explanatory variables to a single data frame
explantory_variables <- mcf %>% 
  select(rain, tmax,TPI_five,dem_10m_zaca,flow_accumulation,sr,slope,TWI)

# create folds
folds <-  createMultiFolds(mortality_qual, k=10, times=10)

# train Control
control <- trainControl(method = "repeatedcv", number = 10, repeats = 10, index = folds, savePredictions = T)

# cross validation model for random forest
rf.cv <- train(x= explantory_variables, y=mortality_qual, method="rf", tuneLength = 2, ntree=500, trControl=control)

# summary statistics
rf.cv
```

Accuracy pretty much the same without using cross validation.

### Chris' modified binomial model (original code??)

```{r, eval = F, echo = F}
#Prepare data
all_env_variables <- brick(stack(list.files(path = "E:/zaca/environmental_data/all_variables/stack", full.names = TRUE)))
pine <- shapefile("E:/zaca/chris/ground_truth/all_mixed_pine", stringsAsFactors = TRUE)
all_env_variables <- crop(all_env_variables, extent(pine))
environment_extract <- raster::extract(all_env_variables, pine, method = "simple", fun = mean, na.rm = TRUE, sp = TRUE)

env_df <- as.data.frame(environment_extract) %>% 
  filter(prefire != 0) %>% 
  mutate(dem_scaled = dem_10m_zaca/1000, insolation_scaled = insolation_10m_zaca/100000, tmean_scaled = tmean_10m_zaca/10, southness = abs(180 - aspect_10m_zaca), precip_mm = rain_10m_zaca/100) %>% 
  mutate(postfire_success = as.numeric(as.character(car::recode(postfire, "'0' = 0; '1_25' = 125; '25_75' = 500; '75_100' = 875")))) %>% 
  mutate(postfire_failure = 1000 - postfire_success) %>% 
  droplevels(.)

env_df$prefire <- relevel(env_df$prefire, ref = "75_100")

blr <- glm(cbind(postfire_success, postfire_failure) ~ TPI_five_10m_zaca + insolation_scaled + prefire, family = binomial(link = "logit"), data = env_df)

blr

#ifelse(condition, yes, no)
# example ifelse code to get probabilities
class.out <- ifelse(prob < 0.25, "None", 
                    ifelse(prob < 0.5, "Low", 
                           ifelse(...)))

table(class.out, actual.classes)
```

### Chris' modified binomial model (modified by amp): 4 classes

```{r}
# model parameters---------------------------------------------

# set sample size 
nsample4 = 125

#set number of foor loop iterations
iteration = 100

# lists to save outputs from for loop
test_accuracy_list <- list()
```

```{r}

# for loop binomial model -----------------------------------------------------------------

for (i in 1:iteration) {
# train and test data 
  mcf_high <- mcf[which(mcf$mortality_qual == "High"),]
  mcf_low <- mcf[which(mcf$mortality_qual == "Low"),]
  mcf_moderate <- mcf[which(mcf$mortality_qual == "Moderate"),]
  mcf_none <- mcf[which(mcf$mortality_qual == "None"),]

  # pull random observations from the high class and separate
  x_high <- sample(1:nrow(mcf_high), nsample4 , replace = FALSE)
  polygon_data_high_train <- mcf_high[x_high,]
  polygon_data_high_test <- mcf_high[-x_high,]
  # pull random observations from the low class and separate
  x_low <- sample(1:nrow(mcf_low), nsample4 , replace = FALSE)
  polygon_data_low_train <- mcf_low[x_low,]
  polygon_data_low_test <- mcf_low[-x_low,]
  # pull random observations from the moderate class and separate
  x_moderate <- sample(1:nrow(mcf_moderate), nsample4 , replace = FALSE)
  polygon_data_moderate_train <- mcf_moderate[x_moderate,]
  polygon_data_moderate_test <- mcf_moderate[-x_moderate,]
  # pull random observations from the none class and separate
  x_none <- sample(1:nrow(mcf_none), nsample4 , replace = FALSE)
  polygon_data_none_train <- mcf_none[x_none,]
  polygon_data_none_test <- mcf_none[-x_none,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_none_train, polygon_data_low_train, polygon_data_moderate_train,  polygon_data_high_train)) 
  # recombine classes that were not pulled
  test <- do.call("rbind", list(polygon_data_none_test, polygon_data_low_test, polygon_data_moderate_test, polygon_data_high_test))

#model
blr <- glm(cbind(canopy_loss_success, canopy_loss_failure) ~ dem_10m_zaca + rain + prefire + slope + TPI_five, family = binomial(link = "logit"), data = train)

#model outputs
blr
summary(blr)

#predict 
prob <- predict(blr, newdata = test, type = "response")

# ifelse code to get probabilities
class_out <- ifelse(prob < 0.25, "None", 
                    ifelse(prob < 0.5, "Low", 
                           ifelse(prob < 0.75, " Moderate", 
                                  ifelse(prob <1, "High"))))

# convert probability outputs from vector to data frame and reorder the mortality classes. Having trouble
model_predictions <- data.frame(class_out)

write.table(model_predictions, "test.csv", sep = ",", col.names = NA, append=F)
model_test <- read_csv(here::here("Model Data Sheets and Code", "test.csv"))
model_test <- model_test %>% 
  mutate(class_out = fct_relevel(class_out, levels = c("None", "Low", "Moderate", "High")))
file.remove(here::here("Model Data Sheets and Code", "test.csv"))

# test data accuracy 
tabb <- table(Predicted = model_test$class_out, Actual = test$canopy_loss_success)
acc<-1-sum(diag(tabb))/sum(tabb) #% missclassified

test_accuracy <-(1-acc)*100 #model accuracy

# lists to save model outputs
test_accuracy_list <- append(test_accuracy_list, test_accuracy)
write.table(tabb, file = "binomial model test accuracy table.csv", sep=",",col.names=NA, append=T)

}

```

#### Model Results
```{r}
# model output results ---------------------------------------------------------------------------------------

# average model accuracy
accuracy_df <- as.data.frame(unlist(test_accuracy_list)) %>% 
  summarise(accuracy_mean = mean(unlist(test_accuracy)))

# average class accuracy
high <- read_csv("binomial model test accuracy table.csv") %>%  rename("High" = "750")
moderate <- read_csv("binomial model test accuracy table.csv")  %>%  rename("Moderate" = "500")
low <- read_csv("binomial model test accuracy table.csv")  %>%  rename("Low" = "250")
none <- read_csv("binomial model test accuracy table.csv") %>%  rename("None" = "0")

high_mean_acc <- high %>% 
  select (X1, High) %>% 
  drop_na() %>% 
  filter (X1 == "High") %>% 
  mutate (High = as.numeric(High),
          class_acc = High / 33) %>% 
  summarise(mean = (sum(class_acc)/iteration))

moderate_mean_acc <- moderate %>% 
  select (X1, Moderate) %>% 
  drop_na() %>% 
  filter (X1 == "Moderate") %>%
  mutate (Moderate = as.numeric(Moderate),
          class_acc = Moderate / 35)%>% 
  summarise(mean = (sum(class_acc)/iteration))

low_mean_acc <- low %>% 
  select (X1, Low) %>% 
  drop_na() %>% 
  filter (X1 == "Low") %>%
  mutate (Low = as.numeric(Low),
          class_acc = Low / 41) %>% 
  summarise(mean = (sum(class_acc)/iteration))

none_mean_acc <- none %>% 
  select (X1, None) %>% 
  drop_na() %>% 
  filter (X1 == "None") %>%
  mutate (None = as.numeric(None),
          class_acc = None / 126) %>% 
  summarise(mean = (sum(class_acc)/iteration))

```

Notes:

- I changed the response varaible from the difference in pre and post fire canopy cover (in your data its the postfire success/failure, in my data its the change in canopy success/failure), to fire induced canopy loss (i.e. mortality) success/failure

- also split up the data into train and test data with an equal number of observations in each mortality class for the train data. The model kept confusing the mortality with low/none since none had such a higher number of observations

- I don't know why I am having so much trouble with reordering the levels of the test data predictions (after converting the list from vector data to a data frame). I keep getting this error: *Outer Names are only allowed for unnamed scalar atomic* and the fct_relevel won't actually reorder. I tried several different codes (factor, fct_relevel, fct-reorder, etc) to order the class, but the only thing that worked was exporting the data into a csv file then reimporting it back into r, then deleting the file (bc its in a for loop). fct_relevel works now. I think R just doesn't like vector data.

- An issue I was having was when there are no accurate predictions for a mortality class(ie High), that 0 value doesnt get stored in the output table, so when I tried to get the average accuracy for that mortality class I don't have a complete list of the values. I fixed this by suming the acuracy each iteration that appeared in the data set, then dividing that by the total number of iterations run. works.

- i dont know why I am having so much trouble with converting the test data predictions into a data frame and then reordering the levels. If I get the data from to work then the factor works but moderate is registered as an NA when it wasnt before. If i use mutate fct_relevel then I get this error: fct_relvel bc keep getting this error: Outer names are only allowed for unnamed scalar atomic inputs (even when used mutate to convert the data from atomic data and is.atomic() did not register the data as atomic; use is.atomic(data_name$column_name) to check if r recognizes the data as atomic. I think this issue is occuring bc the data I'm converting to a data frame was originally vector data). Convertign the class_out to a data frame, exporting it to a csv, and then reimporting it was the only way to reorder the levels without having moderate converted to NA

### Chris' modified binomial model (modified by amp): 3 classes-combine low and none classes

```{r}
# model parameters---------------------------------------------

# set sample size 
nsample5 = 125

#set number of foor loop iterations
iteration1 = 100

# sub-data: combine low and no mortality classes
mcf_3 <- mcf %>%
  mutate(mortality_qual_3 = case_when(
    mortality_qual %in% c("None", "Low") ~ "Low",
    mortality_qual == "Moderate" ~ "Moderate",
    mortality_qual == "High" ~ "High"
  )) %>% 
  mutate(canopy_loss_success_3 = case_when(
    canopy_loss_success %in% c("0", "250") ~ "250",
    canopy_loss_success == "500" ~"500",
    canopy_loss_success == "750" ~ "750"
  )) %>% 
  mutate(canopy_loss_failure_3 = case_when(
    canopy_loss_failure %in% c("1000", "750") ~ "750",
    canopy_loss_failure == "500" ~"500",
    canopy_loss_failure == "250" ~ "250"
  ))

mcf_3$canopy_loss_success_3 <- factor(mcf_3$canopy_loss_success_3)
mcf_3$canopy_loss_failure_3 <- factor(mcf_3$canopy_loss_failure_3)

# lists to save outputs from for loop

test_accuracy_list <- list()
```

```{r}

# for loop binomial model -----------------------------------------------------------------

for (i in 1:iteration1) {
# train and test data 
  mcf_high <- mcf_3[which(mcf_3 $ mortality_qual_3 == "High"),]
  mcf_low <- mcf_3[which(mcf_3 $ mortality_qual_3 == "Low"),]
  mcf_moderate <- mcf_3[which(mcf_3 $ mortality_qual_3 == "Moderate"),]

  # pull random observations from the high class and separate
  x_high <- sample(1:nrow(mcf_high), nsample5, replace = FALSE)
  polygon_data_high_train <- mcf_high[x_high,]
  polygon_data_high_test <- mcf_high[-x_high,]
  # pull random observations from the low class and separate
  x_low <- sample(1:nrow(mcf_low), nsample5 , replace = FALSE)
  polygon_data_low_train <- mcf_low[x_low,]
  polygon_data_low_test <- mcf_low[-x_low,]
  # pull random observations from the moderate class and separate
  x_moderate <- sample(1:nrow(mcf_moderate), nsample5 , replace = FALSE)
  polygon_data_moderate_train <- mcf_moderate[x_moderate,]
  polygon_data_moderate_test <- mcf_moderate[-x_moderate,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_low_train, polygon_data_moderate_train,  polygon_data_high_train)) 
  # recombine classes that were not pulled
  test <- do.call("rbind", list(polygon_data_low_test, polygon_data_moderate_test, polygon_data_high_test))

#model
blr <- glm(cbind(canopy_loss_success_3, canopy_loss_failure_3) ~ dem_10m_zaca + rain + prefire + slope + TPI_five, family = binomial(link = "logit"), data = train)

#model outputs
blr
summary(blr)

#predict 
prob <- predict(blr, newdata = test, type = "response")

# ifelse code to get probabilities
class_out <- ifelse(prob < 0.25, "Low", 
                     ifelse(prob < 0.75, " Moderate", 
                             ifelse(prob <1, "High")))

# convert probability outputs from vector to data frame and reorder the mortality classes. 
model_predictions <- data.frame(class_out)

write.table(model_predictions, "test.csv", sep = ",", col.names = NA, append=F)
model_test <- read_csv(here::here("Model Data Sheets and Code", "test.csv"))
model_test <- model_test %>% 
  mutate(class_out = fct_relevel(class_out, levels = c("Low", "Moderate", "High")))
file.remove(here::here("Model Data Sheets and Code", "test.csv"))

# test data accuracy 
tabb <- table(Predicted = model_test$class_out, Actual = test$canopy_loss_success_3)
acc<-1-sum(diag(tabb))/sum(tabb)
test_accuracy <-(1-acc)*100 #model accuracy

# lists to save model outputs
test_accuracy_list <- append(test_accuracy_list, test_accuracy)
write.table(tabb, file = "binomial model test accuracy table.csv", sep=",",col.names=NA, append=T)

}

```

#### Model Results
```{r}
# model output results ---------------------------------------------------------------------------------------

# average model accuracy
accuracy_df <- as.data.frame(unlist(test_accuracy_list)) %>% 
  summarise(accuracy_mean = mean(unlist(test_accuracy)))

# average class accuracy
high <- read_csv("binomial model test accuracy table.csv") %>%  rename("High" = "750")
moderate <- read_csv("binomial model test accuracy table.csv")  %>%  rename("Moderate" = "500")
low <- read_csv("binomial model test accuracy table.csv")  %>%  rename("Low" = "250")

high_mean_acc <- high %>% 
  select (X1, High) %>% 
  drop_na() %>% 
  filter (X1 == "High") %>% 
  mutate (High = as.numeric(High),
          class_acc = High / 33) %>% 
  summarise(mean = (sum(class_acc)/iteration))

moderate_mean_acc <- moderate %>% 
  select (X1, Moderate) %>% 
  drop_na() %>% 
  filter (X1 == "Moderate") %>%
  mutate (Moderate = as.numeric(Moderate),
          class_acc = Moderate / 35)%>% 
  summarise(mean = (sum(class_acc)/iteration))

low_mean_acc <- low %>% 
  select (X1, Low) %>% 
  drop_na() %>% 
  filter (X1 == "Low") %>%
  mutate (Low = as.numeric(Low),
          class_acc = Low / 292) %>% 
 summarise(mean = (sum(class_acc)/iteration)) # used this code before to get mean: summarise(mean = mean(class_acc))
```

### Future Models

Chose the GFDL-CM3 model which is from the NOAA Geophysical Fluid Dynamics Laboratory. Wasn't sure which one to pick, but NOAA is a reputable organization.

- LOCA VIR RUN (Scripps): Has monthly data for RCP 4.5 and 8.5 for precipitation and "air temperature"" 
- LOCA Downscaled CMIP5 Climate Projections (Scripps): daily and annual data for RCP 4.5 and 8.5 (no monthly data) for min temp, max temp, precip, solar radiation, relative humidity
- LOCA Derived Products: mixture of things. RCP 4.5 and 8.5. No monthly data for climate variables

```{r}

```

### 2020-03-18 Shade Fraction Data Sheets and Code/Shade Fractions TC Model.R

@@ -1,172 +1,0 @@
-# load packages ----------------------------------------------------------------------------------------------
-
-library(tidyverse)
-library(here)
-library(dplyr)
-library(pROC)
-library(ROCR)
-
-# add data ---------------------------------------------------------------------------------------------------
-
-sf2 <-read_csv(here::here("Shade Fraction Data Sheets and Code", "Final Zaca MCF SF_MainData.csv")) %>% 
-  dplyr::mutate(mortality_qualitative= fct_relevel(mortality_qualitative, levels=c("None", "Low", "Moderate", "High"))) %>% 
-  dplyr::mutate(other_mortality = fct_relevel(other_mortality, levels = c("None", "Low", "Moderate", "High", "Type Converted")))
-
-
-#You can predict values with the model using the predict() function. This enables you to graph the fit line. You'll need to create a data frame with the values you want to input for each predictor variable. You may want to set all of the variables except one to a constant value.
-# create lists ------------------------------------------------------------------------------------------------------------------------------------------
-acc_list <- list()
-table_values <- list()
-model_aic <-list()
-auc_list <- list ()
-#prediction_list <- list()
-
-#type conversion model ----------------------------------------------------------------------------------------------------------------------------
-for (i in 1:500)
-  {
-  #divide the response variable into equal number of observations
-  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
-  tc_no <- sf2[which(sf2$new_model_class == "0"),]
-  
-  # pull 175 random observations from the high class and separate
-  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
-  polygon_data_high_train <- tc_yes[yes,]
-  # pull 175 random observations from the low class and separate
-  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
-  polygon_data_low_train <- tc_no[no,]
-  
-  # combine all random observations
-  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))
-  
-  tc <- glm(new_model_class~shade_2018+max_2000_2006, data=train, family="binomial")
-  sum <-summary(tc)
-  sum
[…]

### 2020-03-18 Shade Fraction Data Sheets and Code/predcit canopy cover with RS.R

@@ -1,189 +1,0 @@
-# load packages and data -----------------------------------------------------
-
-library(randomForest)
-library(tidyverse)
-library(here)
-
-# load data -----------------------------------------------------------------
-
-#shade fraction time series data
-canopy_sf <- read_csv(here::here("Shade Fraction Data Sheets and Code", "Final Zaca MCF SF_MainData.csv")) %>% 
-  mutate(prefire_qual = case_when(
-    prefire %in% '1_25' ~ "Sparse (1-25%) Canopy" ,
-    prefire %in% '25_75' ~ "Open (25-75%) Canopy" ,
-    prefire %in% '75_100' ~ "Dense (75-100%) Canopy"
-  )) %>% 
-  mutate(prefire_qual = as.factor(prefire_qual)) %>% 
-  mutate(postfire_qual = case_when(
-    postfire %in% c("0", '1_25') ~ "Sparse (1-25%) Canopy" ,
-    postfire %in% '25_75' ~ "Open (25-75%) Canopy" ,
-    postfire %in% '75_100' ~ "Dense (75-100%) Canopy"
-  )) %>% 
-  mutate(postfire_qual = as.factor(postfire_qual)) %>% 
-  dplyr::mutate(mortality_qualitative= fct_relevel(mortality_qualitative, levels=c("None", "Low", "Moderate", "High"))) %>% 
-  mutate(other_mortality = fct_relevel(other_mortality, levels = c("None", "Low", "Moderate", "High", "Type Converted")))
-
-# green vegation time series data
-canopy_gv <- read_csv(here::here("Shade Fraction Data Sheets and Code", "Zaca MCF variables for model_plus GV.csv")) %>% 
-  mutate(prefire_qual = case_when(
-    prefire %in% '1_25' ~ "Sparse (1-25%) Canopy" ,
-    prefire %in% '25_75' ~ "Open (25-75%) Canopy" ,
-    prefire %in% '75_100' ~ "Dense (75-100%) Canopy"
-  )) %>% 
-  dplyr::mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
-  rename(mortality_qualitative = mortality_qual)
-
-# sample size --------------------------------------------------------------------------------------
-
-# sample size will be the same for shade fraction and gv fractions (and dnbr when thats added)
-mortality_sample_size <- canopy_sf %>% 
-  group_by(mortality_qualitative) %>% 
[…]

### 2020-03-18 Shade Fraction Data Sheets and Code/type conversion model trials and results.Rmd
copy over files from laptop

---
title: "Type Conversion Model"
author: "Anne-Marie Parkinson"
date: "February 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r}
library(tidyverse)
library(here)
library(dplyr)
library(pROC)
library(ROCR)
```

# add data
```{r}
sf2 <-read_csv(here::here("Shade Fraction Data Sheets and Code", "Final Zaca MCF SF_MainData.csv")) %>% 
  dplyr::mutate(mortality_qualitative= fct_relevel(mortality_qualitative, levels=c("None", "Low", "Moderate", "High"))) %>% 
  dplyr::mutate(other_mortality = fct_relevel(other_mortality, levels = c("None", "Low", "Moderate", "High", "Type Converted")))

gv2 <- read_csv(here::here("Shade Fraction Data Sheets and Code", "Zaca MCF variables for model_plus GV.csv")) %>% 
  dplyr::mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
  rename(mortality_qualitative = mortality_qual)

```

# sample size 
```{r}
count <- sf2 %>% 
  group_by(new_model_class) %>% 
  count()
```

# type conversion model trials

## shade fractions: one response variable

```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
  tc_no <- sf2[which(sf2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~shade_2018, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## Shade fractions: two response variables
```{r}

# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
  tc_no <- sf2[which(sf2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~shade_2018+max_2000_2006, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))

```

## shade fractions: three response variables
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
  tc_no <- sf2[which(sf2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~shade_2018+max_2000_2006+min_2000_200, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## shade fractions: four response variables

```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- sf2[which(sf2$new_model_class == "1"),]
  tc_no <- sf2[which(sf2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~shade_2018+max_2000_2006+min_2000_200+prefire, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV fractions: one response variable
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~gv_2018, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV: two response variables 
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~gv_2018 + gv_min_2000_2006, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV: three response variables

```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~gv_2018 + gv_min_2000_2006 + gv_max_2000_2006, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV: four response variables 
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~gv_2018 + gv_min_2000_2006, gv_max_2000_2006 + gv_avg_2000_2006, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

## GV: four response variables 
```{r}
# make lists to later store values
acc_list <- list()
table_values <- list()
model_aic <-list()
auc_list <- list ()
#prediction_list <- list()

#type conversion model ------------------------------------
for (i in 1:500)
  {
  #divide the response variable into equal number of observations
  tc_yes <- gv2[which(gv2$new_model_class == "1"),]
  tc_no <- gv2[which(gv2$new_model_class == "0"),]

  # pull 175 random observations from the high class and separate
  yes <- sample(1:nrow(tc_yes), 99 , replace = FALSE)
  polygon_data_high_train <- tc_yes[yes,]
  # pull 175 random observations from the low class and separate
  no <- sample(1:nrow(tc_no), 99 , replace = FALSE)
  polygon_data_low_train <- tc_no[no,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_high_train, polygon_data_low_train))

  tc <- glm(new_model_class~gv_2018 + gv_min_2000_2006, gv_max_2000_2006 + prefire, data=train, family="binomial")
  sum <-summary(tc)
  sum
  aic<- sum$aic
  aic
  model_aic <- append(model_aic, aic)

  # predict data 
  pred<-predict(tc, train, type="response")

  # error rate for data
  pred1 <- ifelse (pred>0.5,1,0)
  tab<- table(Predicted=pred1, Actual=train$new_model_class)
  write.table(tab, file = "type conversion train response variable accuracy.csv", sep=",",col.names=NA, append=T)

  acc <-1-sum(diag(tab))/sum(tab)   #error calculation. gives % observations predicted incorrectly 
  acc <- 1- acc 
  acc_list <- append(acc_list, acc)

  #auc score
  pred <- predict(tc,train)
  pred1 <-prediction(pred, train$new_model_class)
  roc <- performance(pred1, "tpr", "fpr")
  auc <-performance(pred1, "auc")
  auc1 <- unlist(slot(auc, "y.values"))
  auc_list <- append(auc_list, auc1)

}

model_aic <- as.data.frame(unlist(model_aic))
acc_list <- as.data.frame(unlist(acc_list))
auc_list <-as.data.frame(unlist(auc_list))

combo <- cbind(model_aic, acc_list, auc_list)
write.table(model_aic, file="model_aic.csv", sep=",")

# table of mean model outputs ---------------------------------------------------

sf_combo_data <- read_csv(here::here("Shade Fraction Data Sheets and Code", "model_aic.csv"))
# the column headers are off, they

# read in data and make it R friendly 
type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% 
  rename(type_converted = '0')
not_type_converted <- read_csv(here::here("Shade Fraction Data Sheets and Code", "type conversion train response variable accuracy.csv")) %>% rename(not_type_converted = '1')

type_converted <- type_converted %>% 
  select (X1, type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 0) %>% 
  mutate(class_acc = type_converted/99)
type_converted_accuracy <- type_converted %>% 
   summarise(mean = mean(class_acc))

not_type_converted <- not_type_converted %>% 
  select (X1, not_type_converted) %>% 
  drop_na() %>% 
  filter (X1 == 1) %>% 
  mutate(class_acc = not_type_converted/99)
not_type_converted_accuracy <- not_type_converted %>% 
   summarise(mean = mean(class_acc))
```

### 2020-03-18 Shade Fraction Data Sheets and Code/predcit canopy cover with RS.R

@@ -1,0 +1,189 @@
+# load packages and data -----------------------------------------------------
+
+library(randomForest)
+library(tidyverse)
+library(here)
+
+# load data -----------------------------------------------------------------
+
+#shade fraction time series data
+canopy_sf <- read_csv(here::here("Shade Fraction Data Sheets and Code", "Final Zaca MCF SF_MainData.csv")) %>% 
+  mutate(prefire_qual = case_when(
+    prefire %in% '1_25' ~ "Sparse (1-25%) Canopy" ,
+    prefire %in% '25_75' ~ "Open (25-75%) Canopy" ,
+    prefire %in% '75_100' ~ "Dense (75-100%) Canopy"
+  )) %>% 
+  mutate(prefire_qual = as.factor(prefire_qual)) %>% 
+  mutate(postfire_qual = case_when(
+    postfire %in% c("0", '1_25') ~ "Sparse (1-25%) Canopy" ,
+    postfire %in% '25_75' ~ "Open (25-75%) Canopy" ,
+    postfire %in% '75_100' ~ "Dense (75-100%) Canopy"
+  )) %>% 
+  mutate(postfire_qual = as.factor(postfire_qual)) %>% 
+  dplyr::mutate(mortality_qualitative= fct_relevel(mortality_qualitative, levels=c("None", "Low", "Moderate", "High"))) %>% 
+  mutate(other_mortality = fct_relevel(other_mortality, levels = c("None", "Low", "Moderate", "High", "Type Converted")))
+
+# green vegation time series data
+canopy_gv <- read_csv(here::here("Shade Fraction Data Sheets and Code", "Zaca MCF variables for model_plus GV.csv")) %>% 
+  mutate(prefire_qual = case_when(
+    prefire %in% '1_25' ~ "Sparse (1-25%) Canopy" ,
+    prefire %in% '25_75' ~ "Open (25-75%) Canopy" ,
+    prefire %in% '75_100' ~ "Dense (75-100%) Canopy"
+  )) %>% 
+  dplyr::mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
+  rename(mortality_qualitative = mortality_qual)
+
+# sample size --------------------------------------------------------------------------------------
+
+# sample size will be the same for shade fraction and gv fractions (and dnbr when thats added)
+mortality_sample_size <- canopy_sf %>% 
+  group_by(mortality_qualitative) %>% 
[…]

### 2020-03-28 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
ALE plots!!! :)

@@
 library(ggfortify)
+library(ALEPlot)
+library(iml)
 
@@
 ```{r}
-#sub-data
 mcf_sub <- mcf %>% 
-  select (dem_10m_zaca:sr, rain:tmax, TPI_five, TWI, cwd:vpd_jan)
+  select (dem_10m_zaca:sr, rain:tmax, TPI_five, TWI, cwd:vpd_aug) %>% 
+  rename ('Elevation' = 'dem_10m_zaca',
+          'Slope' = 'slope',
+          'TPI' = 'TPI_five',
+          'Precip' = 'rain',
+          'Solar Radiation' = 'sr',
+          'Flow Accumulation' = 'flow_accumulation',
+          'CWD' = 'cwd',
+          'Temperature' = 'tmax',
+          'VPD (Jan)' = 'vpd_jan',
+          'VPD (Aug)' = 'vpd_aug')
 
@@
 mcf_sub_pca_plot <- autoplot(mcf_sub_pca, 
-                          colour = NA, # HAVE to spell it color this way (colour) 
+                          colour = NA, # HAVE to spell it color this way (colour) bc thats the only spelling form autoplot recognizes
                           loadings.label = T,
@@
                           loadings.label.repel = T) +
-  scale_y_continuous(lim=c(-0.1, 0.15)) +
-  labs(title = "Zaca Fire MCF PCA")
+  scale_y_continuous(lim=c(-0.15, 0.15)) +
+  labs(title = "Zaca Fire MCF PCA") +
+  theme_minimal()
 
@@
   #random forest model
-  rfm<-randomForest(mortality_qual~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI, data = train, ntree=500, importance=T, nscale=0)
+  rfm<-randomForest(mortality_qual~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug, data = train, ntree=500, importance=T, nscale=0)
 
@@
 
-Note: when running the model, have to remember to delete the rfm model importance.csv and mcf test accuracy table.csv files otherwise R will keep adding the values onto the orignial file and not make a new one for each model succession.
+- Note: when running the model, have to remember to delete the rfm model importance.csv and mcf test accuracy table.csv files otherwise R will keep adding the values onto the orignial file and not make a new one for each model succession.
 I've been doing that or renaming the file, but maybe theres an easier method
+- QUESTION: Should I be using set seed?
 
@@
 
+file.remove(here::here("Model Data Sheets and Code", "rfm model importance.csv"))
 # average class accuracy ---------------------------------------------------------
@@
 
+file.remove(here::here("Model Data Sheets and Code", "mcf test accuracy table.csv"))
+
 ```
 
-#### Partial Dependence Plots
+#### ALE Plots
+
+So I was using partial dependence plots to visualize the variables in RF, but then found out you shouldn't use PDP when the variables are correlated with other variables it's one of the assumptions of PDP https://christophm.github.io/interpretable-ml-book/pdp.html ; https://arxiv.org/pdf/1612.08468.pdf).
+Instead, I should be using Accumulated Local Effects (ALE) plots (https://christophm.github.io/interpretable-ml-book/ale.html).
+Unfortunatly, there is not a lot of information out there on how to set up the code.
+Maybe that stats friend you mentioned way back when will know how and we can offer co-authorship.
+
+- code example of how to make ALE using iml package:
+-- https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
+-- https://cran.r-project.org/web/packages/iml/vignettes/intro.html
+
+- webpage: https://christophm.github.io/interpretable-ml-book/ale.html
+- github code: https://github.com/christophM/interpretable-ml-book/blob/6153b2ac0143f879bd4a856a1048545a40cd1a74/manuscript/05.4-agnostic-ale.Rmd
+
 ```{r}
+train_x <- train[-which(names(train) == "mortality_qual")]
+
+sr_predictor_high <- Predictor$new(rfm, data = train_x, type = "prob", class = "High")
+sr_high <- plot(FeatureEffect$new(sr_predictor_high, feature = "sr")) + ggtitle("High")
+
+sr_predictor_moderate <- Predictor$new(rfm, data = train_x, type = "prob", class = "Moderate")
+sr_moderate <- plot(FeatureEffect$new(sr_predictor_moderate, feature = "sr")) + ggtitle("Moderate")
+
+sr_predictor_low <- Predictor$new(rfm, data = train_x, type = "prob", class = "Low")
+sr_low <- plot(FeatureEffect$new(sr_predictor_low, feature = "sr")) + ggtitle("Low")
+
+sr_predictor_none <- Predictor$new(rfm, data = train_x, type = "prob", class = "None")
+sr_none <- plot(FeatureEffect$new(sr_predictor_none, feature = "sr")) + ggtitle("None")
+
+ggarrange(sr_high, sr_moderate, sr_low, sr_none, ncol = 2, nrow = 2)
+
+sr_df <- aml_sr$results
+ggplot(sr_df, aes(x = sr, y = .ale)) + geom_line()
 
 ```
+- get error when try to run the model on the test data instead of the train data. Think the data used in the model, needs to be the same data to do ALE plots.
+
+- How to interpret ALE plots:
+  - ale plots let us determine the effect that each individual input, isolated/regardless of all others, has on our output.
+  - By localizing the measurements via the use of windows, we are able to avoid including practically unlikely or impossible situations.
+For instance, in our model to predict the number of joggers in a day, temperature and time of year are likely highly correlated.
+By limiting our window in our ALE plot, we can ensure that we're not factoring in a scenario where the time of year is winter and the temperature is 35 degrees Celsius.
+We shouldn't be factoring in the model's prediction for this situation because it is certainly outside of any data set we will be using.
+ALE plots are able to avoid such situations and give us much more accurate results.
 
@@
   )
-
+file.remove(here::here("Model Data Sheets and Code", "rfm model importance_3 class mortality.csv"))
 # average class accuracy ---------------------------------------------------------
@@
   summarise(mean = mean(class_acc))
+
+file.remove(here::here("Model Data Sheets and Code", "mcf test accuracy table_3 class mortality.csv"))
+
+```
+
+```{r}
+train_x <- train[-which(names(train) == "three_class_mortality")]
+
+sr_predictor_high <- Predictor$new(rfm, data = train_x, type = "prob", class = "High")
+sr_high <- plot(FeatureEffect$new(sr_predictor_high, feature = "sr")) + ggtitle("High")
+
+sr_predictor_moderate <- Predictor$new(rfm, data = train_x, type = "prob", class = "Moderate")
+sr_moderate <- plot(FeatureEffect$new(sr_predictor_moderate, feature = "sr")) + ggtitle("Moderate")
+
+sr_predictor_low <- Predictor$new(rfm, data = train_x, type = "prob", class = "Low")
+sr_low <- plot(FeatureEffect$new(sr_predictor_low, feature = "sr")) + ggtitle("Low")
+
+ggarrange(sr_high, sr_moderate, sr_low, ncol = 2, nrow = 2)
 
@@
 
-### Chris' modified binomial model (original code??)
+### Chris' modified binomial model (original code)
 
@@
 
-#set number of foor loop iterations
-iteration = 100
+# set # of iterations to run
+
+iteration = 50
 
@@
 
+- I looked at the PCA to pick variables. It was the easiest way to determine which variables were strongly correlated with other variables
+
 #### Model Results
@@
   mutate (High = as.numeric(High),
-          class_acc = High / 33) %>% 
+          class_acc = High /33) %>% 
   summarise(mean = (sum(class_acc)/iteration))
@@
   mutate (Moderate = as.numeric(Moderate),
-          class_acc = Moderate / 35)%>% 
+          class_acc = Moderate / 35) %>% 
   summarise(mean = (sum(class_acc)/iteration))
@@
           class_acc = None / 126) %>% 
-  summarise(mean = (sum(class_acc)/iteration))
+  summarise(mean = ((sum(class_acc))/iteration))
+
+file.remove(here::here("Model Data Sheets and Code", "binomial model test accuracy table.csv"))
 
@@
 - I don't know why I am having so much trouble with reordering the levels of the test data predictions (after converting the list from vector data to a data frame).
-I keep getting this error: *Outer Names are only allowed for unnamed scalar atomic* and the fct_relevel won't actually reorder.
+I keep getting this error: *Outer Names are only allowed for unnamed scalar atomic* and the fct_relevel won't actually reorder correctly.
 I tried several different codes (factor, fct_relevel, fct-reorder, etc) to order the class, but the only thing that worked was exporting the data into a csv file then reimporting it back into r, then deleting the file (bc its in a for loop).
@@
 - LOCA Derived Products: mixture of things. RCP 4.5 and 8.5. No monthly data for climate variables
+- units (for CNRM-CM5, CanESM2): temperature = Kelvin, precipitation = kg/m2/s, solar radiation = W/m2
+- unsure about units from HadGEM2, GFDL-CM3, or MIROC5 (may be same as above)
 
-```{r}
-
-```
+- *** so I started downloading data to start this analysis and then realized the precipitation values are in flux format.
+Worse, in units of seconds, not days.
+I may learn how to deal with this type of data for future modelingspring quarter, but not there yet.

### 2020-03-30 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
final draft code for chris

@@
 Instead, I should be using Accumulated Local Effects (ALE) plots (https://christophm.github.io/interpretable-ml-book/ale.html).
-Unfortunatly, there is not a lot of information out there on how to set up the code.
-Maybe that stats friend you mentioned way back when will know how and we can offer co-authorship.
+May also want to look at ICE plots (https://towardsdatascience.com/an-overview-of-model-explainability-in-modern-machine-learning-fc0f22c8c29a)
 
 - code example of how to make ALE using iml package:
--- https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
--- https://cran.r-project.org/web/packages/iml/vignettes/intro.html
+  - https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
+  - https://cran.r-project.org/web/packages/iml/vignettes/intro.html
 
-- webpage: https://christophm.github.io/interpretable-ml-book/ale.html
-- github code: https://github.com/christophM/interpretable-ml-book/blob/6153b2ac0143f879bd4a856a1048545a40cd1a74/manuscript/05.4-agnostic-ale.Rmd
+- webpage explaining ALE: https://christophm.github.io/interpretable-ml-book/ale.html
+- github code for webpage above: https://github.com/christophM/interpretable-ml-book/blob/6153b2ac0143f879bd4a856a1048545a40cd1a74/manuscript/05.4-agnostic-ale.Rmd
+
+other sources (1-2 journal article)
+- https://towardsdatascience.com/an-overview-of-model-explainability-in-modern-machine-learning-fc0f22c8c29a
+- https://www.moodysanalytics.com/-/media/presentation/2019/mas-t1d1-transparent-and-interpretable-ml-in-risk-modeling.pdf
+- https://www.enjine.com/blog/interpreting-machine-learning-models-accumulated-local-effects/ --> it would be cool to be able to make the plot in this website
+- https://arxiv.org/pdf/1612.08468.pdf
+- https://cran.r-project.org/web/packages/iml/iml.pdf
+- https://cran.r-project.org/web/packages/iml/vignettes/intro.html
+- https://cran.r-project.org/web/packages/ALEPlot/ALEPlot.pdf (iml package and documentation a lot more useful)
+
+So, I just wanted to get the code started.
+I think you will be better at streamlining makingthe rest of the graphs, especially with the same axis scales.
+One thing I did notice is that ALE plots for the same mortality class resulted in
 
 ```{r}
+
+## sample code to make ALE plos (using solar radiation as example) used combo of: https://cran.r-project.org/web/packages/iml/vignettes/intro.html and https://cran.r-project.org/web/packages/iml/iml.pdf
+
+#do something with the response variable column (removes it but R can still recognize it)
 train_x <- train[-which(names(train) == "mortality_qual")]
 
+# this does something to create values that will be plotted in the ALE plot: inputs include the random forest model (rfm) run above, the trainig data with the modified response column (created above), specifying we want output to be a probability, looking at a specific mortality class
 sr_predictor_high <- Predictor$new(rfm, data = train_x, type = "prob", class = "High")
+
+# plot the outputs in an ALE plot using the predictor info created above just for solar radiation variable(feature = "explanatory variable").
+Doesnt work if do featureeffects$new() and then plot() on separate lines even though thats how one of the smaple codes showed it.
 sr_high <- plot(FeatureEffect$new(sr_predictor_high, feature = "sr")) + ggtitle("High")
 
+#predict values for the other mortality classes for the same variable
 sr_predictor_moderate <- Predictor$new(rfm, data = train_x, type = "prob", class = "Moderate")
@@
 
-sr_df <- aml_sr$results
-ggplot(sr_df, aes(x = sr, y = .ale)) + geom_line()
+#plot all variables(for just one mortality class) : for some reason the function makes graphs for all the variables in the data frame, not just the ones in the random forest model. 
+tt <- FeatureEffects$new(sr_predictor_high, method = "ale")
+plot(tt)
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+all_rfm_variables <- FeatureEffects$new(sr_predictor_low, method = "ale", features = c("cwd", "rain", 'tmax' , 'TPI_five' , 'dem_10m_zaca' , 'flow_accumulation' , 'sr' ,'slope', 'TWI', 'vpd_aug'))
+all_rfm_variables$plot()
+ggsave("ale_low.jpg", height = 9, width = 9)
+
+# plot a variables in ALE plot: https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
+plot(FeatureEffect$new(sr_predictor_low, feature = c("cwd", "dem_10m_zaca")))
+
+# another way to make ale plts with iml package (doesnt work): https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
+sr_predictor_low <- Predictor$new(rfm, data = train_x, type = "prob", class = "Low")
+ale_iml <- FeatureEffect$new(sr_predictor_low, feature = "sr")
+sr_df <- ale_iml$results
+ggplot(sr_df, aes(x = sr, y = .ale)) + geom_line() # not sure what .ale is. the samle code doesn't say, but the above codes work so whatever
+
+# another way to write the code: https://github.com/christophM/interpretable-ml-book/blob/6153b2ac0143f879bd4a856a1048545a40cd1a74/manuscript/05.4-agnostic-ale.Rmd
+FeatureEffects$new(sr_predictor_low, feature = "sr", method = "ale")$plot() +ggtitle("Low") #produces the same graph as above
 
@@
 
-- How to interpret ALE plots:
 - ale plots let us determine the effect that each individual input, isolated/regardless of all others, has on our output.
+
+- How to interpret ALE plots: not really sure since the papers/tutorials I've seen show how to interpret grpahs for quantitative response variables, but I assume since we indicate "type = prob" in the code that the y axis is the probability of a stand having that mortality class at specific value of the x variable
+
 - By localizing the measurements via the use of windows, we are able to avoid including practically unlikely or impossible situations.
@@
   #random forest model
-  rfm<-randomForest(three_class_mortality~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI, data = train, ntree=500, importance=T, nscale=0)
+  rfm<-randomForest(three_class_mortality~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug, data = train, ntree=500, importance=T, nscale=0)
 
@@
 
+#### ALE Plots
 ```{r}
@@
 ggarrange(sr_high, sr_moderate, sr_low, ncol = 2, nrow = 2)
+
+# all rfm variables
+all_rfm_variables <- FeatureEffects$new(sr_predictor_low, method = "ale", features = c("cwd", "rain", 'tmax' , 'TPI_five' , 'dem_10m_zaca' , 'flow_accumulation' , 'sr' ,'slope', 'TWI', 'vpd_aug'))
+all_rfm_variables$plot()
+ggsave("ale_low_3.jpg", height = 9, width = 9)
 
@@
 
-iteration = 50
+iteration = 500
 
@@
 #set number of foor loop iterations
-iteration1 = 100
+iteration1 = 500
 

### 2020-03-30 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
another final draft code (negligable change) plus ale graph images for chris

@@
  summarise(mean = (sum(class_acc)/iteration)) # used this code before to get mean: summarise(mean = mean(class_acc))
+
+file.remove(here::here("Model Data Sheets and Code", "binomial model test accuracy table.csv"))
+
 ```

### 2020-04-01 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
added one more random forest trial. didnt add it to a binomial trial tho

@@
 mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
-  select('drainage_assoication', 'ridge', 'mortality', 
+  select('drainage_association', 'ridge', 'mortality', 
          'prefire','mortality_qual', 'canopy_loss_success', 
@@
          'rain', 'slope', 'tmax', 'tmean', 'tmin', 'TPI_five', 
-         'TPI_queen', 'TWI', 'fri', 'dem_scaled', 'southness', "cwd", "vpd_jan", "vpd_aug" ) 
+         'TPI_queen', 'TWI', 'fri', 'dem_scaled', 'southness', "cwd", "vpd_jan", "vpd_aug", "tc_mortality" ) 
  #mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High")))
@@
 ### Random Forest
-
-Which variables contribute to mortality?
-
-Literature review:
 
@@
 ```
+
+### Random Forest
+
+```{r}
+# model parameters---------------------------------------------
+
+mcf_mod <- mcf %>% 
+  mutate(tc_mortality = case_when(
+    tc_mortality == 1 ~ "tc", #tc = type converted
+    tc_mortality == 0 ~ "not_tc"
+  )) %>% 
+  mutate(tc_mortality = as.factor(tc_mortality))
+
+# set sample size 
+nsample = 75
+
+# lists to save outputs from for loop
+test_accuracy_list <- list()
+
+```
+
+```{r}
+
+#random forest code---------------------------------------------------
+for (i in 1:50)
+{
+
+  # divide data into equal # of observations --------------------
+  mcf_high <- mcf_mod[which(mcf_mod$tc_mortality == "tc"),]
+  mcf_low <- mcf_mod[which(mcf_mod$tc_mortality == "not_tc"),]
+
+  # pull random observations from the high class and separate
+  x_high <- sample(1:nrow(mcf_high), nsample , replace = FALSE)
+  polygon_data_high_train <- mcf_high[x_high,]
+  polygon_data_high_test <- mcf_high[-x_high,]
+  # pull random observations from the low class and separate
+  x_low <- sample(1:nrow(mcf_low), nsample , replace = FALSE)
+  polygon_data_low_train <- mcf_low[x_low,]
+  polygon_data_low_test <- mcf_low[-x_low,]
+
+  # combine all random observations
+  train <- do.call("rbind", list(polygon_data_low_train, polygon_data_high_train))
+  # recombine classes that were not pulled
+  test <- do.call("rbind", list(polygon_data_low_test, polygon_data_high_test))
+
+  #random forest model
+  rfm<-randomForest(tc_mortality~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ridge + drainage_association, data = train, ntree=500, importance=T, nscale=0)
+
+  #plot to visualize variable importance
+  #varImpPlot(rfm)
+  model_importance <- importance(rfm)
+
+  #predcit test data
+  pp<-predict(rfm, test)
+
+  #test data accuracy
+  tabb<- table(pp, test$tc_mortality)
+  acc<-1-sum(diag(tabb))/sum(tabb) # % missclassified 
+  test_accuracy <-(1-acc)*100 #model accuracy
+
+  # lists to save model outputs
+  test_accuracy_list <- append(test_accuracy_list, test_accuracy)
+  write.table(model_importance, file="rfm model tc importance.csv", sep=",",col.names=NA, append=T) 
+  write.table(tabb, file = "mcf tc test accuracy table.csv", sep=",",col.names=NA, append=T)
+}
+
+```
+
+#### Model Results
+```{r}
+
+# overall average model accuracy -----------------------------------------------------
+
+accuracy_df <- as.data.frame(unlist(test_accuracy_list))
+mean_test_accuracy_1 <- accuracy_df %>% 
+  summarize(mean_accuracy = mean(unlist(test_accuracy_list)))
+
+#average model importance ---------------------------------------------------------
+
+model_importance_tidy <- read_csv(here("Model Data Sheets and Code", "rfm model tc importance.csv")) %>% 
+  clean_names() %>% 
+  filter(not_tc != "not_tc") %>% 
+  rename (variables = x1) %>% 
+  mutate (variables = str_to_lower(variables)) %>% 
+  mutate (not_tc = as.numeric(not_tc),
+          tc = as.numeric(tc),
+          mean_decrease_accuracy = as.numeric(mean_decrease_accuracy),
+          mean_decrease_gini = as.numeric(mean_decrease_gini)
+          )
+
+# model importance summary statistics
+model_importance_summary <- model_importance_tidy %>% 
+  group_by(variables) %>% 
+  summarize (not_tc_mean = mean(not_tc),
+             tc_mean = mean(tc),
+             mean_dec_accuracy_mean = mean(mean_decrease_accuracy),
+             mean_decrease_gini_mean = mean(mean_decrease_gini)
+  )
+file.remove(here::here("Model Data Sheets and Code", "rfm model tc importance.csv"))
+# average class accuracy ---------------------------------------------------------
+
+not_tc <- read_csv("mcf tc test accuracy table.csv") 
+tc <- read_csv("mcf tc test accuracy table.csv")
+
+not_tc_acc <- not_tc %>% 
+  select (X1, not_tc) %>% 
+  drop_na() %>% 
+  filter (X1 == "not_tc") %>% 
+  mutate (not_tc = as.numeric(not_tc),
+          class_acc = not_tc / (636-nsample)) %>% 
+  summarise(mean = mean(class_acc))
+
+tc_acc <- tc %>% 
+  select (X1, tc) %>% 
+  drop_na() %>% 
+  filter (X1 == "tc") %>%
+  mutate (tc = as.numeric(tc),
+          class_acc = tc / (99-nsample)) %>% 
+  summarise(mean = mean(class_acc))
+
+file.remove(here::here("Model Data Sheets and Code", "mcf tc test accuracy table.csv"))
+
+```
+ Didnt get to ALE plots.
+But model importance shows that slope and elevation are highly important for the type converted class.
+Not surprised because a lot of stands with the highest mortality were at the top of Big pine mountain which was at the highest elevation in our study area.
+This could indicate that this is trend is specific to this fire situation and does not necessarily predict that stands with high mortality are highly likely to occur in high elevation zones.
 

### 2020-06-08 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
worked on rfm model and added ndvi to the data sheet

@@
 library(iml)
+library(ROCR)
+library(pROC)
 
 mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
-  select('drainage_association', 'ridge', 'mortality', 
-         'prefire','mortality_qual', 'canopy_loss_success', 
-         'canopy_loss_failure', 'postfire_canopy', 'postfire', 
-         'Change_in_canopy_success', 'Change_in_canopy_failure', 
-         'Change_Qualitative','high_mortality', 'area_scaled', 
-         'dem_10m_zaca', 'flow_accumulation', 'sr', 'sr_new', 
-         'rain', 'slope', 'tmax', 'tmean', 'tmin', 'TPI_five', 
-         'TPI_queen', 'TWI', 'fri', 'dem_scaled', 'southness', "cwd", "vpd_jan", "vpd_aug", "tc_mortality" ) 
+  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
+  # mutate(dnbr_qualitative = case_when(
+  #   dnbr < 0.1 ~ "unburned",
+  #   dnbr < 0.27 ~ "low",
+  #   dnbr < 0.66 ~ "moderate",
+  #   dnbr > 0.66 ~ "high")) %>% 
+  # mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
+  #        dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned"))) %>% 
+  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
+  mutate(three_class_mortality = case_when(
+    mortality_qual %in% "High" ~ "High",
+    mortality_qual %in% "Moderate" ~ "Moderate",
+    mortality_qual %in% c("Low", "None") ~ "Low")) %>% 
+  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
+  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
+                                     levels=c("Low", "Moderate", "High")))
+
  #mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High")))
@@
 The code directly below works to reorder the levels without error)
-mcf$mortality_qual  <- factor(mcf$mortality_qual, levels = c("None", "Low", "Moderate", "High"))
-
-```
+
+```
+for dnbr classes see: https://www.int-arch-photogramm-remote-sens-spatial-inf-sci.net/XLI-B8/763/2016/isprs-archives-XLI-B8-763-2016.pdf
+
+for aspect transformation: https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/07-0539.1 appendix B; transformed Aspect = [1 - cos(2piAspect/360)]/2.
+This formula is similar (but not identical) to that used by Roberts and Cooper (1989).
+The transformed aspect values lie in the interval from 0 to1.
+Values near 0 represent aspects close to due north, while values near 1 represent aspects close to due south.
+East and west are treated identically with transformed aspect values of 0.5.
 
@@
 
-```
+tc_count <- mcf %>% 
+  group_by(tc_mortality) %>% 
+  count() # 99 stands with type conversion, 636 with no type conversion
+
+```
+None: 251
+Low: 166
+mod: 160
+high: 158
 
@@
   labs(x="Mortality", y="Solar Radiation") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="Max Aug Temp (celcius)") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="Precipitation") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="TPI") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="TWI") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="Slope") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="Elevation") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="CWD") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="Aug VPD") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
   labs(x="Mortality", y="Jan VPD") +
-  scale_fill_manual(values=c("red", "orange", "yellow", "green")) +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
+  coord_flip()
+
+ggplot(mcf, aes(x=mortality_qual, y=ndvi, fill=mortality_qual))+
+  geom_violin(alpha=0.5, show.legend=F) +
+  theme_classic() +
+  labs(x="Mortality", y="NDVI") +
+  scale_fill_manual(values=c("green", "yellow", "orange", "red")) +
   coord_flip()
@@
 mcf_sub <- mcf %>% 
-  select (dem_10m_zaca:sr, rain:tmax, TPI_five, TWI, cwd:vpd_aug) %>% 
+  select (dem_10m_zaca:sr, rain:tmax, TPI_five, TWI, cwd:aspect_transformed) %>% 
   rename ('Elevation' = 'dem_10m_zaca',
@@
 
-### Random Forest
-
+### Random Forest: ALL observations
+
+to get a stable variable importance. The other models result in very unstable variable importance cmeasures which shuldnt happen with RF but is occuring bc the training data used for each run changes.
+
+```{r}
+#four class mortality
+for (i in 1:10) {
+
+rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ndvi, data = mcf, ntree=1500, importance=T, nscale=0)
+
+rfm
+
+#plot to visualize variable importance
+varImpPlot(rfm)
+model_importance <- importance(rfm)
+
+}
+
+#3 class mortality
+for (i in 1:10) {
+
+rfm <- randomForest(three_class_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ndvi, data = mcf, ntree=1500, importance=T, nscale=0)
+
+rfm
+
+#plot to visualize variable importance
+varImpPlot(rfm)
+model_importance <- importance(rfm)
+
+}
+
+```
+So 1500 trees reuslts in more stable importance values for each model. The model with 3 mortality classes has a better OOB.
+
+# Random Forest: 4 mortality classes
 ```{r}
@@
   #random forest model
-  rfm<-randomForest(mortality_qual~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug, data = train, ntree=500, importance=T, nscale=0)
+  rfm<-randomForest(mortality_qual~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ndvi, data = train, ntree=500, importance=T, nscale=0)
 
@@
 
-### Random Forest Model with 3 mortality classes
+### Random Forest Model: 3 mortality classes
 
@@
 ```{r}
-# model parameters and sub data -----------------------------------------------------------
-
-# sub data 
-mcf_3 <- mcf %>%
-  mutate(three_class_mortality = case_when(
-    mortality_qual %in% "High" ~ "High",
-    mortality_qual %in% "Moderate" ~ "Moderate",
-    mortality_qual %in% c("Low", "None") ~ "Low"
-  )) %>% 
-  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
-  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
-                                     levels=c("Low", "Moderate", "High")))
+# model parameters ------------------------------
 
@@
 #random forest code---------------------------------------------------
-for (i in 1:500)
+for (i in 1:5)
 {
@@
 
-  mcf_high <- mcf_3[which(mcf_3$three_class_mortality == "High"),]
-  mcf_low <- mcf_3[which(mcf_3$three_class_mortality == "Low"),]
-  mcf_moderate <- mcf_3[which(mcf_3$three_class_mortality == "Moderate"),]
+  mcf_high <- mcf[which(mcf$three_class_mortality == "High"),]
+  mcf_low <- mcf[which(mcf$three_class_mortality == "Low"),]
+  mcf_moderate <- mcf[which(mcf$three_class_mortality == "Moderate"),]
 
@@
   #random forest model
-  rfm<-randomForest(three_class_mortality~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug, data = train, ntree=500, importance=T, nscale=0)
+  rfm<-randomForest(three_class_mortality~ rain + tmax + TPI_five + dem_10m_zaca + ndvi + sr + slope + TWI + cwd + vpd_aug, data = train, ntree=500, importance=T, nscale=0)
 
@@
   model_importance <- importance(rfm)
-
-  #varImpPlot(rfm)
+  varImpPlot(rfm)
 
@@
   #random forest model
-  rfm<-randomForest(tc_mortality~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ridge + drainage_association, data = train, ntree=500, importance=T, nscale=0)
+  rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ridge + drainage_association + ndvi, data = train, ntree=1500, importance=T, nscale=0)
 
   #plot to visualize variable importance
-  #varImpPlot(rfm)
+  varImpPlot(rfm)
   model_importance <- importance(rfm)
@@
 accuracy_df <- as.data.frame(unlist(test_accuracy_list))
-mean_test_accuracy_1 <- accuracy_df %>% 
+mean_model_accuracy_1 <- accuracy_df %>% 
   summarize(mean_accuracy = mean(unlist(test_accuracy_list)))
@@
 This could indicate that this is trend is specific to this fire situation and does not necessarily predict that stands with high mortality are highly likely to occur in high elevation zones.
+
+#### ROC and AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(train$tc_mortality, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+
+```
 
@@
 # sub-data: combine low and no mortality classes
-mcf_3 <- mcf %>%
+mcf <- mcf %>%
   mutate(mortality_qual_3 = case_when(
@@
 
-mcf_3$canopy_loss_success_3 <- factor(mcf_3$canopy_loss_success_3)
-mcf_3$canopy_loss_failure_3 <- factor(mcf_3$canopy_loss_failure_3)
+mcf$canopy_loss_success_3 <- factor(mcf$canopy_loss_success_3)
+mcf$canopy_loss_failure_3 <- factor(mcf$canopy_loss_failure_3)
 
@@
 # train and test data 
-  mcf_high <- mcf_3[which(mcf_3 $ mortality_qual_3 == "High"),]
-  mcf_low <- mcf_3[which(mcf_3 $ mortality_qual_3 == "Low"),]
-  mcf_moderate <- mcf_3[which(mcf_3 $ mortality_qual_3 == "Moderate"),]
+  mcf_high <- mcf[which(mcf $ mortality_qual_3 == "High"),]
+  mcf_low <- mcf[which(mcf $ mortality_qual_3 == "Low"),]
+  mcf_moderate <- mcf[which(mcf $ mortality_qual_3 == "Moderate"),]
 

### 2020-06-23 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
explored sampsize fucntion

@@
   mutate(three_class_mortality= fct_relevel(three_class_mortality, 
-                                     levels=c("Low", "Moderate", "High")))
+                                     levels=c("Low", "Moderate", "High"))) %>% 
+  mutate(tc_mortality = as.factor(tc_mortality))
 
@@
 
+```
+
+### Random Forest forced sample sizes
+
+https://stats.stackexchange.com/questions/457451/how-r-randomforest-sampsize-works
+
+tc_mortality 1 = type converted
+tc_mortality 0 = "not_tc"
+
+using for loop to generate several graphs so its easier to compare them
+```{r}
+
+for (i in 1:10) {
+
+rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + drainage_association + ndvi, 
+                    data = mcf, 
+                    sampsize = c('1' = 99, '0' = 99), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+
+```
+variable importance seems stable AND class accuracy is high and equal between classes!!
+
+#### ROC and AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(mcf$tc_mortality, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+
+```
+
+## PDP v. ALE
+
+```{r}
+#PDP
+a <- partialPlot(rfm, as.data.frame(mcf), ndvi, "0") #0=not type converted
+b <- partialPlot(rfm, as.data.frame(mcf), ndvi, "1", col="red") #1=type converted
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(mcf), slope, "0") #0=not type converted
+b <- partialPlot(rfm, as.data.frame(mcf), slope, "1", col="red") #1=type converted
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(mcf), sr, "0") #0=not type converted
+b <- partialPlot(rfm, as.data.frame(mcf), sr, "1", col="red") #1=type converted
+lines(a, col="green")
+
+#ale (see detailed notes in ALE Plots code chunk below; not working right now)
+train_x <- mcf[-which(names(mcf) == "tc_mortality")]
+ndvi_predictor_notTC <- Predictor$new(rfm, data = train_x, y = mcf$tc_mortality, type = "prob", class = "0")
+ndvi_not <- plot(FeatureEffect$new(ndvi_predictor_notTC, feature = "ndvi")) + ggtitle("not TC")
+
+#predict values for the other mortality classes for the same variable
+ndvi_predictor_tc <- Predictor$new(rfm, data = train_x, type = "prob", class = "1")
+sr_moderate <- plot(FeatureEffect$new(ndvi_predictor_tc, feature = "ndvi")) + ggtitle("TC")
+```
+so looking at the 3 most important variables for type conversino plots, all of the probabilities are below zero.
+only the not type converted stands have probabilities above 0, with some values of the explanatory variable having very high probabilities.
+So what this tells me is the model CANNOT distinguish type converted stands from non type converted stands.
+
+### Random Forest and forced sample sizes with training and test data
+
+hmmmm.
+not sure how to do this with training and test data since the train/test data frames will have different quantities of the type conversion classes each time.
+if i do it the way I have been doing it below, then variable importance is too unstable.
+even using my equal distribution sample size code and adding in the sampsize function results in unstable variable importance.
+```{r}
+#train and test data 
+sample <- sample(1:nrow(mcf), size = 0.75*nrow(mcf))
+train <- mcf[sample, ]
+test  <- mcf[-sample, ]
+
+for (i in 1:10) {
+rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + drainage_association + ndvi, 
+                    data = mcf, 
+                    sampsize = c('1' = 99, '0' = 99), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+rfm
+
+#plot to visualize variable importance
+varImpPlot(rfm)
+#model_importance <- importance(rfm)
+importance(rfm)
+}
+#predcit test data
+pp<-predict(rfm, test)
+
+#test data accuracy
+tabb<- table(pp, test$tc_mortality)
+acc<-1-sum(diag(tabb))/sum(tabb) # % missclassified 
+test_accuracy <-(1-acc)*100 #model accuracy
 ```
@@
 #random forest code---------------------------------------------------
-for (i in 1:50)
+for (i in 1:15)
 {
@@
   #random forest model
-  rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ridge + drainage_association + ndvi, data = train, ntree=1500, importance=T, nscale=0)
+  rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ridge + drainage_association + ndvi, data = train, ntree=1500, importance=T, nscale=0, sampsize = c('tc' = 75, 'not_tc' = 75), replace = T)
 
@@
   # lists to save model outputs
-  test_accuracy_list <- append(test_accuracy_list, test_accuracy)
-  write.table(model_importance, file="rfm model tc importance.csv", sep=",",col.names=NA, append=T) 
-  write.table(tabb, file = "mcf tc test accuracy table.csv", sep=",",col.names=NA, append=T)
+  # test_accuracy_list <- append(test_accuracy_list, test_accuracy)
+  # write.table(model_importance, file="rfm model tc importance.csv", sep=",",col.names=NA, append=T) 
+  # write.table(tabb, file = "mcf tc test accuracy table.csv", sep=",",col.names=NA, append=T)
 }

### 2020-07-02 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
added methods/goals section

@@
 
+### Methods background
+
+using forced sample sizes (see https://stats.stackexchange.com/questions/457451/how-r-randomforest-sampsize-works) because the distribution of observations in each mortality class is very uneven and results in a model that severely underpredicts observations in the mortality class with few observations.
+Bc of the small sample size, we will not be using training and test data (and bc RF basically creates its own train/test data).
+
+Different RF trials:
+
+  - Mixed confier forests
+    + response variable: type conversion
+    + response variable: 4 mortality classes (none, low, momderate, high)
+    + response variable: 3 mortality classes (none/low, moderate, high)
+    + response variable: 2 mortality classes (none/low/moderate, high)
+    + response variable: rdnbr
+  - Coulter pines 
+    + response variable: rdnbr
+  - mixed conifer forests and coulter pines combined
+    + response variable: rdnbr
+
+Note: manual classification of mortality and type conversion for the coulter pines was not done. Aint no one got time for that
+
+The type conversion mortality classes-
+tc_mortality 0 = "not_tc" (n=99)
+tc_mortality 1 = type converted (n=>600)
+
+### Goals/Questions
+
+1. Can the mcf model distinguish between type conversion and the other mortality types? 
+  - so far, no! so variable importance is not meaningful for that model
+
+2. Which variables contribute to type conversion of MCFs?
+
+3. How well does rdnbr compare to manual estimates of mortality?
+
+4. How does using rdnbr and manually derived mortality compare as a response variable?
+
+5. Which variables contribute to high/low mortality? high/low rdnbr?
+
+5. Does conifer type contribute to different variable importances?
+
 ### Sample Size for each mortality class
@@
   coord_flip()
+
+ggplot(mcf, aes(x=ndvi, y=rdnbr_2007)) +
+  geom_point()
+
+ndvi_rdnbr_corr <- lm(rdnbr_2007~ndvi, data=mcf)
+summary(ndvi_rdnbr_corr)
 
@@
 
-### Random Forest forced sample sizes
-
-https://stats.stackexchange.com/questions/457451/how-r-randomforest-sampsize-works
-
-tc_mortality 1 = type converted
-tc_mortality 0 = "not_tc"
-
-using for loop to generate several graphs so its easier to compare them
+### RF manual mortality: 4 classes (none, low, moderate, high)
+
+Just using for loop to generate several graphs so its easier to compare them and make sure importance is stable
 ```{r}

### 2020-07-07 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
more model trials, updates/new data, prelim results

@@
 mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
-  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
-  # mutate(dnbr_qualitative = case_when(
-  #   dnbr < 0.1 ~ "unburned",
-  #   dnbr < 0.27 ~ "low",
-  #   dnbr < 0.66 ~ "moderate",
-  #   dnbr > 0.66 ~ "high")) %>% 
-  # mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
-  #        dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned"))) %>% 
+  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("High", "Moderate", "Low", "None"))) %>% 
+  mutate(dnbr_qualitative = case_when(
+    dnbr_2007 < 0.1 ~ "unburned",
+    dnbr_2007 < 0.27 ~ "low",
+    dnbr_2007 < 0.66 ~ "moderate",
+    dnbr_2007 > 0.66 ~ "high")) %>%
+  mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
+         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned"))) %>%
   mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
@@
                                      levels=c("Low", "Moderate", "High"))) %>% 
-  mutate(tc_mortality = as.factor(tc_mortality))
+  mutate(two_class_mortality = case_when(
+    mortality_qual %in% "High" ~ "High",
+    mortality_qual %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
+  mutate(two_class_mortality = as.factor(two_class_mortality)) %>% 
+  mutate(two_class_mortality= fct_relevel(two_class_mortality, 
+                                     levels=c("Other", "High"))) %>% 
+  mutate(tc_mortality = as.factor(tc_mortality)) %>% 
+  mutate(rdnbr_qualitative = case_when(
+    rdnbr_2007 < 69 ~ "None",
+    rdnbr_2007 < 315 ~ "Low",
+    rdnbr_2007 < 640 ~ "Moderate",
+    rdnbr_2007 > 640 ~ "High")) %>%
+  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
+         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, 
+                                         levels=c("High", "Moderate", "Low", "None"))) %>% 
+    mutate(rdnbr_qualitative2 = case_when(
+    rdnbr_qualitative %in% "High" ~ "High",
+    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
+  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
+  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, 
+                                     levels=c("Other", "High")))
 
@@
 The code directly below works to reorder the levels without error)
+
+pico <- read_csv("coulter_env_modified.csv") %>% 
+  mutate(rdnbr_qualitative = case_when(
+    rdnbr_2007 < 69 ~ "None",
+    rdnbr_2007 < 315 ~ "Low",
+    rdnbr_2007 < 640 ~ "Moderate",
+    rdnbr_2007 > 640 ~ "High")) %>%
+  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
+         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, 
+                                         levels=c("High", "Moderate", "Low", "None"))) %>% 
+  na.omit(pico) %>%  # 56 stands have NA values (probs all very small stands)
+  mutate(rdnbr_qualitative2 = case_when(
+    rdnbr_qualitative %in% "High" ~ "High",
+    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
+  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
+  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, 
+                                     levels=c("Other", "High"))) %>% 
+  mutate(aspect_transformed = (1-cos((2*pi*aspect_30m)/360))/2) %>% 
+  rename_all(funs(str_replace(., "_30m", ""))) %>% 
+  mutate(dnbr_qualitative = case_when(
+    dnbr_2007 < 0.1 ~ "unburned",
+    dnbr_2007 < 0.27 ~ "low",
+    dnbr_2007 < 0.66 ~ "moderate",
+    dnbr_2007 > 0.66 ~ "high")) %>%
+  mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
+         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned")))
+
+  #document name is "modified" bc i manually removed all the unnesessary arc columns
 
@@
     + response variable: 2 mortality classes (none/low/moderate, high)
-    + response variable: rdnbr
+    + response variable: rdnbr (regression)
   - Coulter pines 
-    + response variable: rdnbr
+    + response variable: rdnbr (regression)
   - mixed conifer forests and coulter pines combined
-    + response variable: rdnbr
+    + response variable: rdnbr (regression)
 
@@
 ```{r}
+#mcf -------------------------------------------------------------------
 mortality_count <- mcf %>% 
@@
 
-```
-None: 251
-Low: 166
-mod: 160
-high: 158
+rdnbr_mcf <- mcf %>% 
+  group_by(rdnbr_qualitative) %>% 
+  count()
+
+dnbr_mcf <- mcf %>% 
+  group_by(dnbr_qualitative) %>% 
+  count()
+
+#coulter ------------------------------------------------------------------- 
+rdnbr_pico <- pico %>% 
+  group_by(rdnbr_qualitative) %>% 
+  count()
+
+dnbr_pico <- pico %>% 
+  group_by(dnbr_qualitative) %>% 
+  count()
+
+# all -------------------------------------------------------------------
+all_sample_size <- cbind(mortality_count, rdnbr_mcf, dnbr_mcf, rdnbr_pico, dnbr_pico)
+```
+Yah, dnbr is not a good index to use. Rdnbr is better.
+
+### mortality v rdnbr 
+```{r}
+rdnbr_qual_n <- mcf %>% 
+  group_by(rdnbr_qualitative) %>% 
+  count() %>% 
+  mutate(mort_type = "rdnbr") %>% 
+  rename(mortality_qual = rdnbr_qualitative)
+
+mortality_n <- mcf %>% 
+  group_by(mortality_qual) %>% 
+  count() %>% 
+  mutate(mort_type = "manual")
+
+mortality_count <- rbind(rdnbr_qual_n, mortality_n)
+
+ggplot(mortality_count, aes(x = mortality_qual, y = n, fill = mort_type)) + geom_col(position = "dodge") +
+  geom_text(aes(label=n), position = position_dodge(width=0.9), vjust=-0.5) +
+  scale_fill_discrete(name = "How was mortality determined")
+
+```
+
+## type conversion threshold for rdnbr
+```{r}
+##plot -------------------------------------------------------------------------------
+
+# all
+ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
+  geom_jitter()
+
+#by prefire cover
+ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
+  geom_jitter() +
+  facet_wrap(~prefire)
+
+## use lm to get a rdnbr threshold for the type converter stands? ---------------------
+
+threshold <- lm(rdnbr_2007 ~ tc_mortality, data = mcf)
+summary(threshold) #r2=0.53
+
+threshold1 <- lm(rdnbr_2007 ~ tc_mortality + ndvi, data = mcf)
+summary(threshold1) #r2=0.54
+
+threshold2 <- lm(rdnbr_2007 ~ tc_mortality + prefire, data = mcf)
+summary(threshold2) #r2=0.54
+
+# max type converted values ----------------------------------------------
+
+max_nonTC <- mcf %>% 
+  group_by(tc_mortality) %>% 
+  summarise(max = max(rdnbr_2007),
+            min = min(rdnbr_2007)) 
+  #non-type converted=1160.433 ; type converted=1414.205
+
+# potential type conversion threshold --------------------------------------------
+
+tc_threshold <- mcf %>% 
+  mutate(rdnbr_tc_threshold = case_when(
+    rdnbr_2007 <1160 ~ "not_tc",
+    rdnbr_2007 >1160 ~ "tc"
+  ))
+
+```
+well there some overlap with the the 2 classes, but there is also a clear point that the non type converted stands reach (rdnbr=1160) that many type converted stands surpass.
+
+##high mortality v rdnbr
+```{r}
+rdnbr_qual_n <- mcf %>% 
+  group_by(rdnbr_qualitative2) %>% 
+  count() %>% 
+  mutate(mort_type = "rdnbr") %>% 
+  rename(two_class_mortality = rdnbr_qualitative2)
+
+mortality_n <- mcf %>% 
+  group_by(two_class_mortality) %>% 
+  count() %>% 
+  mutate(mort_type = "manual")
+
+mortality_count <- rbind(rdnbr_qual_n, mortality_n)
+
+ggplot(mortality_count, aes(x = two_class_mortality, y = n, fill = mort_type)) + geom_col(position = "dodge") +
+  geom_text(aes(label=n), position = position_dodge(width=0.9), vjust=-0.5) +
+  scale_fill_discrete(name = "How was mortality determined")
+
+```
 
@@
 ndvi_rdnbr_corr <- lm(rdnbr_2007~ndvi, data=mcf)
-summary(ndvi_rdnbr_corr)
+summary(ndvi_rdnbr_corr) #3%
+ndmi_rdnbr_corr <- lm(rdnbr_2007~ndmi_2007, data=mcf)
+summary(ndmi_rdnbr_corr) #82.7%
 
@@
 ### RF manual mortality: 4 classes (none, low, moderate, high)
-
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+                    data = mcf, 
+                    sampsize = c('None' = 158, 'Low' = 158, 'Moderate' = 158, 'High' = 158), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+
+PDP
+```{r}
+#PDP
+a <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "None") 
+c <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Low") 
+d <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Moderate") 
+b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(a, c, d, col= "green", "yellow", "orange") # eh, not working. forgot how to set this up 2 show all 4 off the top of my head
+
+a <- partialPlot(rfm, as.data.frame(mcf), slope, "None") 
+c <- partialPlot(rfm, as.data.frame(mcf), slope, "Low") 
+d <- partialPlot(rfm, as.data.frame(mcf), slope, "Moderate") 
+b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(mcf), sr, "None") 
+c <- partialPlot(rfm, as.data.frame(mcf), sr, "Low") 
+d <- partialPlot(rfm, as.data.frame(mcf), sr, "Moderate")  
+b <- partialPlot(rfm, as.data.frame(mcf), sr, "High", col="red") 
+lines(a, col="green")
+
+```
+
+ALE
+```{r}
+#sub data
+mcf_ale <- mcf %>% 
+  select(mortality_qual, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- mcf_ale[-which(names(mcf_ale) == "mortality_qual")]
+
+# predictors 
+none_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$mortality_qual, type = "prob", class="None")
+low_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$mortality_qual, type = "prob", class="Low")
+moderate_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$mortality_qual, type = "prob", class="Moderate")
+high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$mortality_qual, type = "prob", class="High")
+
+#vpd_jan
+none <- plot(FeatureEffect$new(none_predictor, feature = "vpd_jan")) + ggtitle("None")
+low <- plot(FeatureEffect$new(low_predictor, feature = "vpd_jan")) + ggtitle("Low")
+moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "vpd_jan")) + ggtitle("Moderate")
+high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
+ggarrange(none, low, moderate, high)
+
+#slope
+none <- plot(FeatureEffect$new(none_predictor, feature = "slope")) + ggtitle("None")
+low <- plot(FeatureEffect$new(low_predictor, feature = "slope")) + ggtitle("Low")
+moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "slope")) + ggtitle("Moderate")
+high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High")
+ggarrange(none, low, moderate, high)
+
+#sr
+none <- plot(FeatureEffect$new(none_predictor, feature = "sr")) + ggtitle("None")
+low <- plot(FeatureEffect$new(low_predictor, feature = "sr")) + ggtitle("Low")
+moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "sr")) + ggtitle("Moderate")
+high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High")
+ggarrange(none, low, moderate, high)
+
+```
+
+### RF manual mortality: 3 classes (none/low, moderate, high)
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(three_class_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+                    data = mcf, 
+                    sampsize = c('Low' = 158, 'Moderate' = 158, 'High' = 158),
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+
+PDP
+```{r}
+#PDP
+c <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Low") 
+d <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Moderate") 
+b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "High", col="red", ylim=c(-0.8,0.8))
+
+c <- partialPlot(rfm, as.data.frame(mcf), slope, "Low") 
+d <- partialPlot(rfm, as.data.frame(mcf), slope, "Moderate") 
+b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8))
+
+c <- partialPlot(rfm, as.data.frame(mcf), sr, "Low") 
+d <- partialPlot(rfm, as.data.frame(mcf), sr, "Moderate")  
+b <- partialPlot(rfm, as.data.frame(mcf), sr, "High", col="red")
+
+```
+
+ALE
+```{r}
+#sub data
+mcf_ale <- mcf %>% 
+  select(three_class_mortality, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- mcf_ale[-which(names(mcf_ale) == "three_class_mortality")]
+
+# predictors 
+low_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$three_class_mortality, type = "prob", class="Low")
+moderate_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$three_class_mortality, type = "prob", class="Moderate")
+high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$three_class_mortality, type = "prob", class="High")
+
+#vpd_jan
+low <- plot(FeatureEffect$new(low_predictor, feature = "vpd_jan")) + ggtitle("Low")
+moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "vpd_jan")) + ggtitle("Moderate")
+high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
+ggarrange(low, moderate, high)
+
+#slope
+low <- plot(FeatureEffect$new(low_predictor, feature = "slope")) + ggtitle("Low")
+moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "slope")) + ggtitle("Moderate")
+high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High")
+ggarrange(low, moderate, high)
+
+#sr
+low <- plot(FeatureEffect$new(low_predictor, feature = "sr")) + ggtitle("Low")
+moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "sr")) + ggtitle("Moderate")
+high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High")
+ggarrange(low, moderate, high)
+
+```
+
+### RF manual mortality: 2 classes (none/low/moderate (ie other), high)
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(two_class_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+                    data = mcf, 
+                    sampsize = c('Other' = 158,'High' = 158),
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+ROC and AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(mcf$two_class_mortality, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+```
+PDP
+```{r}
+#PDP
+c <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(c, col="green")
+
+c <- partialPlot(rfm, as.data.frame(mcf), slope, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8))
+
+c <- partialPlot(rfm, as.data.frame(mcf), sr, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf), sr, "High", col="red")
+
+```
+
+### RF manual mortality: 2 classes (none/low (ie other), high)
+removed moderate class
+
+```{r}
+
+mcf_2_class_mod <- mcf %>% 
+  filter(mortality_qual != "Moderate") %>% 
+  mutate(two_class_mortality_mod = case_when(
+    mortality_qual %in% c("None", "Low") ~ "Other",
+    mortality_qual %in% c("High") ~ "High"
+  ))
+mcf_2_class_mod$two_class_mortality_mod <- factor(mcf_2_class_mod$two_class_mortality_mod)  # RF still recognizes that there were 4 mortality classes even though filtered one of them out.
+This allows rf to not recognize the missing class.
+
+for (i in 1:10) {
+
+rfm <- randomForest(two_class_mortality_mod ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+                    data = mcf_2_class_mod, 
+                    sampsize = c('Other' = 158,'High' = 158),
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+
+ROC/AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(mcf_2_class_mod$two_class_mortality_mod, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+```
+
+PDP
+```{r}
+#PDP
+a <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), vpd_jan, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), slope, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), slope, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), sr, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), sr, "High", col="red") 
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), TPI_five, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf_2_class_mod), TPI_five, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(a, col="green")
+
+```
+
+ALE
+```{r}
+#sub data
+mcf_ale <- mcf_2_class_mod %>% 
+  select(two_class_mortality_mod, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- mcf_ale[-which(names(mcf_ale) == "two_class_mortality_mod")]
+
+# predictors 
+other_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$two_class_mortality_mod, type = "prob", class="Other")
+high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$two_class_mortality_mod, type = "prob", class="High")
+
+#vpd_jan
+other <- plot(FeatureEffect$new(other_predictor, feature = "vpd_jan")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
+ggarrange(other, high)
+
+#slope
+other <- plot(FeatureEffect$new(other_predictor, feature = "slope")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High")
+ggarrange(other, high)
+
+#sr
+other <- plot(FeatureEffect$new(other_predictor, feature = "sr")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High")
+ggarrange(other, high)
+
+```
+
+### RF manual: type conversion
 Just using for loop to generate several graphs so its easier to compare them and make sure importance is stable
+
+0 = not type converted
+1 = type converted
 ```{r}
@@
 
-rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + drainage_association + ndvi, 
+rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
                     data = mcf, 
@@
 ```
-variable importance seems stable AND class accuracy is high and equal between classes!!
-
-#### ROC and AUC
+variable importance is KINDA/MOSTLY stable.
+There arent big changes between runs, some variables flip 1 space, but some move 3 spaces.
+Variable importance was more stable when ndvi was included in the model.
+
+ROC and AUC
 ```{r}
@@
 auc(rfm_roc)
-
-```
-
-## PDP v. ALE
-
+```
+
+PDP
 ```{r}
 #PDP
-a <- partialPlot(rfm, as.data.frame(mcf), ndvi, "0") #0=not type converted
-b <- partialPlot(rfm, as.data.frame(mcf), ndvi, "1", col="red") #1=type converted
+a <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "0") #0=not type converted
+b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "1", col="red") #1=type converted
 lines(a, col="green")
@@
 
-#ale (see detailed notes in ALE Plots code chunk below; not working right now)
-train_x <- mcf[-which(names(mcf) == "tc_mortality")]
-ndvi_predictor_notTC <- Predictor$new(rfm, data = train_x, y = mcf$tc_mortality, type = "prob", class = "0")
-ndvi_not <- plot(FeatureEffect$new(ndvi_predictor_notTC, feature = "ndvi")) + ggtitle("not TC")
-
-#predict values for the other mortality classes for the same variable
-ndvi_predictor_tc <- Predictor$new(rfm, data = train_x, type = "prob", class = "1")
-sr_moderate <- plot(FeatureEffect$new(ndvi_predictor_tc, feature = "ndvi")) + ggtitle("TC")
 ```
@@
 So what this tells me is the model CANNOT distinguish type converted stands from non type converted stands.
+
+ALE
+```{r}
+#sub data
+mcf_ale <- mcf %>% 
+  select(tc_mortality, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- mcf_ale[-which(names(mcf_ale) == "tc_mortality")]
+
+# predictors 
+nottc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$tc_mortality, type = "prob", class="0")
+tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$tc_mortality, type = "prob", class="1")
+
+#vpd_jan
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "vpd_jan")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_jan")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+#slope
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "slope")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+#sr
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "sr")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+```
+
+# rdnbr: classification (2 classes)
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_qualitative2 ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+                    data = mcf, 
+                    sampsize = c('High' = 150, 'Other' = 150), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+get very similar oob and class accuracy as the manual mortality trial with 2 classes (none/low/moderate, high).
+So if we just look at high mortality v other mortality then we can use either rdnbr or manual mortality but if we chose to show the other classes it makes sense to use manual mortality since it is a lot more accurate, and thus more representative, than rdnbr.
+
+ROC and AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(mcf$rdnbr_qualitative2, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+```
+
+PDP
+```{r}
+#PDP
+c <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(c, col="green")
+
+c <- partialPlot(rfm, as.data.frame(mcf), slope, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8))
+
+c <- partialPlot(rfm, as.data.frame(mcf), sr, "Other") 
+b <- partialPlot(rfm, as.data.frame(mcf), sr, "High", col="red")
+
+```
+
+ALE
+```{r}
+#sub data
+mcf_ale <- mcf %>% 
+  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- mcf_ale[-which(names(mcf_ale) == "rdnbr_qualitative2")]
+
+# predictors 
+other_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative2, type = "prob", class="Other")
+high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative2, type = "prob", class="High")
+
+#vpd_jan
+other <- plot(FeatureEffect$new(other_predictor, feature = "vpd_jan")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
+ggarrange(other, high)
+
+#slope
+other <- plot(FeatureEffect$new(other_predictor, feature = "slope")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High")
+ggarrange(other, high)
+
+#sr
+other <- plot(FeatureEffect$new(other_predictor, feature = "sr")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+```
+
+# rdnbr: regression
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_2007 ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan, 
+                    data = mcf, 
+                    #sampsize = c('1' = 99, '0' = 99), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+
+PDP
+```{r}
+#PDP
+partialPlot(rfm, as.data.frame(mcf), vpd_jan) 
+partialPlot(rfm, as.data.frame(mcf), slope) 
+partialPlot(rfm, as.data.frame(mcf), sr)
+
+```
+not sure why the y axis doesnt go higher (the max rdnbr value is over 1400)
+
+ALE
+```{r}
+#sub data
+mcf_ale <- mcf %>% 
+  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- mcf_ale[-which(names(mcf_ale) == "rdnbr_qualitative")]
+
+# predictors 
+predictor <- Predictor$new(rfm, data = mcf_ale)
+
+#explanatory variables
+vpd_jan <- plot(FeatureEffect$new(predictor, feature = "vpd_jan")) + ggtitle("vpd_jan")
+slope <- plot(FeatureEffect$new(predictor, feature = "slope")) + ggtitle("slope")
+sr <- plot(FeatureEffect$new(predictor, feature = "sr")) + ggtitle("sr")
+ggarrange(vpd_jan, slope, sr)
+
+```
+not sure why the y axis doesnt go higher (the max rdnbr value is over 1400)
+
+# rdnbr: regression PICO
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_2007 ~ rain + tmax + TPI_five + dem  + insolation + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, #missing fow accumulation
+                    data = pico, 
+                    #sampsize = c('1' = 99, '0' = 99), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+
+PDP
+```{r}
+#PDP
+partialPlot(rfm, as.data.frame(pico), rain) 
+partialPlot(rfm, as.data.frame(pico), slope) 
+partialPlot(rfm, as.data.frame(pico), insolation)
+
+```
+not sure why the y axis doesnt go higher (the max rdnbr value is over 1400)
+
+ALE
+```{r}
+#sub data
+mcf_ale <- mcf %>% 
+  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- mcf_ale[-which(names(mcf_ale) == "rdnbr_qualitative")]
+
+# predictors 
+predictor <- Predictor$new(rfm, data = mcf_ale)
+
+#explanatory variables
+vpd_jan <- plot(FeatureEffect$new(predictor, feature = "vpd_jan")) + ggtitle("vpd_jan")
+slope <- plot(FeatureEffect$new(predictor, feature = "slope")) + ggtitle("slope")
+sr <- plot(FeatureEffect$new(predictor, feature = "sr")) + ggtitle("sr")
+ggarrange(vpd_jan, slope, sr)
+
+```
+not sure why the y axis doesnt go higher (the max rdnbr value is over 1400)
+
+# rdnbr: classification (2 classes) PICO
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_qualitative2 ~ rain + tmax + TPI_five + dem + insolation + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + years_since_burn,
+                    data = pico, 
+                    sampsize = c('High' = 319, 'Other' = 319), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+get very similar oob and class accuracy as the manual mortality trial with 2 classes (none/low/moderate, high).
+So if we just look at high mortality v other mortality then we can use either rdnbr or manual mortality but if we chose to show the other classes it makes sense to use manual mortality since it is a lot more accurate, and thus more representative, than rdnbr.
+
+ROC and AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(pico$rdnbr_qualitative2, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+```
+
+PDP
+```{r}
+#PDP
+c <- partialPlot(rfm, as.data.frame(pico), vpd_jan, "Other") 
+b <- partialPlot(rfm, as.data.frame(pico), vpd_jan, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(c, col="green")
+
+c <- partialPlot(rfm, as.data.frame(pico), insolation, "Other") 
+b <- partialPlot(rfm, as.data.frame(pico), insolation, "High", col="red", ylim=c(-0.8,0.8))
+
+c <- partialPlot(rfm, as.data.frame(pico), tmax, "Other") 
+b <- partialPlot(rfm, as.data.frame(pico), tmax, "High", col="red")
+
+```
+
+ALE
+```{r}
+#sub data
+pico_ale <- pico %>% 
+  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem, insolation, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- pico_ale[-which(names(pico_ale) == "rdnbr_qualitative2")]
+
+# predictors 
+other_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative2, type = "prob", class="Other")
+high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative2, type = "prob", class="High")
+
+#rain
+other <- plot(FeatureEffect$new(other_predictor, feature = "rain")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High")
+ggarrange(other, high)
+
+#temp
+other <- plot(FeatureEffect$new(other_predictor, feature = "tmax")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High")
+ggarrange(other, high)
+
+#sr
+other <- plot(FeatureEffect$new(other_predictor, feature = "insolation")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(high_predictor, feature = "insolation")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+```
+
+# rdnbr: regression (mcf + pico)
+
+sub data 
+```{r}
+# sub data: combine mcf and pico dfs ----------------------
+
+# pico sub
+pico_sub <- pico %>% 
+  rename(sr = insolation,
+         dem_10m_zaca = dem) %>% 
+  select(-ORIG_FID, -aspect,-years_since_burn, -tmin, -tmean, -TPI_queen, -dnbr_2007, -ndmi_2007, -dnbr_qualitative) #missing flow_accumulation
+
+# mcf sub
+mcf_sub <- mcf %>% 
+  select(rain, tmax,  TPI_five , dem_10m_zaca , sr , slope , TWI , cwd , vpd_aug , vpd_jan , aspect_transformed, rdnbr_2007, rdnbr_qualitative2, rdnbr_qualitative)
+
+# combine
+all <- rbind(pico_sub, mcf_sub)
+
+# sample size
+rdnbr_all_4 <- all %>% 
+  group_by(rdnbr_qualitative) %>% 
+  count()
+```
+high: 863, moderate: 459, low: 437, none: 8
+
+regression
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_2007 ~ rain + tmax + TPI_five + dem_10m_zaca + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+                    data = all, 
+                    #sampsize = c('1' = 99, '0' = 99), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+
+PDP
+```{r}
+#PDP
+partialPlot(rfm, as.data.frame(all), rain) 
+partialPlot(rfm, as.data.frame(all), tmax) 
+partialPlot(rfm, as.data.frame(all), sr)
+
+```
+not sure why the y axis doesnt go higher (the max rdnbr value is over 1400)
+
+ALE
+```{r}
+#sub data
+all_ale <- all %>% 
+  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- all_ale[-which(names(all_ale) == "rdnbr_qualitative")]
+
+# predictors 
+predictor <- Predictor$new(rfm, data = all_ale)
+
+#explanatory variables
+vpd_jan <- plot(FeatureEffect$new(predictor, feature = "vpd_jan")) + ggtitle("vpd_jan")
+slope <- plot(FeatureEffect$new(predictor, feature = "slope")) + ggtitle("slope")
+sr <- plot(FeatureEffect$new(predictor, feature = "sr")) + ggtitle("sr")
+ggarrange(vpd_jan, slope, sr)
+
+```
+
+# rdnbr: classification (mcf + pico)
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_qualitative2 ~ rain + tmax + TPI_five + dem_10m_zaca + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed,
+                    data = all, 
+                    sampsize = c('High' = 863, 'Other' = 863), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+
+ROC and AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(all$rdnbr_qualitative2, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+```
+
+PDP
+```{r}
+#PDP
+c <- partialPlot(rfm, as.data.frame(all), tmax, "Other") 
+b <- partialPlot(rfm, as.data.frame(all), tmax, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(c, col="green")
+
+c <- partialPlot(rfm, as.data.frame(all), rain, "Other") 
+b <- partialPlot(rfm, as.data.frame(all), rain, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(c, col="green")
+
+c <- partialPlot(rfm, as.data.frame(all), sr, "Other") 
+b <- partialPlot(rfm, as.data.frame(all), sr, "High", col="red") 
+lines(c, col="green")
+
+```
+
+ALE
+```{r}
+#sub data
+all_ale <- all %>% 
+  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- all_ale[-which(names(all_ale) == "rdnbr_qualitative2")]
+
+# predictors 
+other_predictor <- Predictor$new(rfm, data = all_ale, y = all_ale$rdnbr_qualitative2, type = "prob", class="Other")
+high_predictor <- Predictor$new(rfm, data = all_ale, y = all_ale$rdnbr_qualitative2, type = "prob", class="High")
+
+#vpd_jan
+other <- plot(FeatureEffect$new(other_predictor, feature = "tmax")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High")
+ggarrange(other, high)
+
+#slope
+other <- plot(FeatureEffect$new(other_predictor, feature = "rain")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High")
+ggarrange(other, high)
+
+#sr
+other <- plot(FeatureEffect$new(other_predictor, feature = "sr")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High")
+ggarrange(other, high)
+
+```
+
+### ALE Plots
+
+So I was using partial dependence plots to visualize the variables in RF, but then found out you shouldn't use PDP when the variables are correlated with other variables it's one of the assumptions of PDP https://christophm.github.io/interpretable-ml-book/pdp.html ; https://arxiv.org/pdf/1612.08468.pdf).
+Instead, I should be using Accumulated Local Effects (ALE) plots (https://christophm.github.io/interpretable-ml-book/ale.html).
+May also want to look at ICE plots (https://towardsdatascience.com/an-overview-of-model-explainability-in-modern-machine-learning-fc0f22c8c29a)
+
+- code example of how to make ALE using iml package:
+  - https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
+  - https://cran.r-project.org/web/packages/iml/vignettes/intro.html
+
+- webpage explaining ALE: https://christophm.github.io/interpretable-ml-book/ale.html
+- github code for webpage above: https://github.com/christophM/interpretable-ml-book/blob/6153b2ac0143f879bd4a856a1048545a40cd1a74/manuscript/05.4-agnostic-ale.Rmd
+
+other sources (1-2 journal article)
+- https://towardsdatascience.com/an-overview-of-model-explainability-in-modern-machine-learning-fc0f22c8c29a
+- https://www.moodysanalytics.com/-/media/presentation/2019/mas-t1d1-transparent-and-interpretable-ml-in-risk-modeling.pdf
+- https://www.enjine.com/blog/interpreting-machine-learning-models-accumulated-local-effects/ --> it would be cool to be able to make the plot in this website
+- https://arxiv.org/pdf/1612.08468.pdf
+- https://cran.r-project.org/web/packages/iml/iml.pdf
+- https://cran.r-project.org/web/packages/iml/vignettes/intro.html
+- **https://rdrr.io/cran/iml/src/R/FeatureEffect.R --> has citation
+- https://cran.r-project.org/web/packages/ALEPlot/ALEPlot.pdf (iml package and documentation a lot more useful)
+
+So, I just wanted to get the code started.
+I think you will be better at streamlining makingthe rest of the graphs, especially with the same axis scales.
+One thing I did notice is that ALE plots for the same mortality class resulted in
+
+```{r}
+
+## sample code to make ALE plos (using solar radiation as example) used combo of: https://cran.r-project.org/web/packages/iml/vignettes/intro.html and https://cran.r-project.org/web/packages/iml/iml.pdf
+
+#do something with the response variable column (removes it but R can still recognize it)
+train_x <- train[-which(names(train) == "mortality_qual")]
+
+# this does something to create values that will be plotted in the ALE plot: inputs include the random forest model (rfm) run above, the trainig data with the modified response column (created above), specifying we want output to be a probability, looking at a specific mortality class
+sr_predictor_high <- Predictor$new(rfm, data = train_x, type = "prob", class = "High")
+
+# plot the outputs in an ALE plot using the predictor info created above just for solar radiation variable(feature = "explanatory variable").
+Doesnt work if do featureeffects$new() and then plot() on separate lines even though thats how one of the smaple codes showed it.
+sr_high <- plot(FeatureEffect$new(sr_predictor_high, feature = "sr")) + ggtitle("High")
+
+#predict values for the other mortality classes for the same variable
+sr_predictor_moderate <- Predictor$new(rfm, data = train_x, type = "prob", class = "Moderate")
+sr_moderate <- plot(FeatureEffect$new(sr_predictor_moderate, feature = "sr")) + ggtitle("Moderate")
+
+sr_predictor_low <- Predictor$new(rfm, data = train_x, type = "prob", class = "Low")
+sr_low <- plot(FeatureEffect$new(sr_predictor_low, feature = "sr")) + ggtitle("Low")
+
+sr_predictor_none <- Predictor$new(rfm, data = train_x, type = "prob", class = "None")
+sr_none <- plot(FeatureEffect$new(sr_predictor_none, feature = "sr")) + ggtitle("None")
+
+ggarrange(sr_high, sr_moderate, sr_low, sr_none, ncol = 2, nrow = 2)
+
+#plot all variables(for just one mortality class) : for some reason the function makes graphs for all the variables in the data frame, not just the ones in the random forest model. 
+tt <- FeatureEffects$new(sr_predictor_high, method = "ale")
+plot(tt)
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+all_rfm_variables <- FeatureEffects$new(sr_predictor_low, method = "ale", features = c("cwd", "rain", 'tmax' , 'TPI_five' , 'dem_10m_zaca' , 'flow_accumulation' , 'sr' ,'slope', 'TWI', 'vpd_aug'))
+all_rfm_variables$plot()
+ggsave("ale_low.jpg", height = 9, width = 9)
+
+# plot a variables in ALE plot: https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
+plot(FeatureEffect$new(sr_predictor_low, feature = c("cwd", "dem_10m_zaca")))
+
+# another way to make ale plts with iml package (doesnt work): https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
+sr_predictor_low <- Predictor$new(rfm, data = train_x, type = "prob", class = "Low")
+ale_iml <- FeatureEffect$new(sr_predictor_low, feature = "sr")
+sr_df <- ale_iml$results
+ggplot(sr_df, aes(x = sr, y = .ale)) + geom_line() # not sure what .ale is. the samle code doesn't say, but the above codes work so whatever
+
+# another way to write the code: https://github.com/christophM/interpretable-ml-book/blob/6153b2ac0143f879bd4a856a1048545a40cd1a74/manuscript/05.4-agnostic-ale.Rmd
+FeatureEffects$new(sr_predictor_low, feature = "sr", method = "ale")$plot() +ggtitle("Low") #produces the same graph as above
+
+```
+- get error when try to run the model on the test data instead of the train data. Think the data used in the model, needs to be the same data to do ALE plots.
+
+- ale plots let us determine the effect that each individual input, isolated/regardless of all others, has on our output.
+
+- How to interpret ALE plots: not really sure since the papers/tutorials I've seen show how to interpret grpahs for quantitative response variables, but I assume since we indicate "type = prob" in the code that the y axis is the probability of a stand having that mortality class at specific value of the x variable
+
+- By localizing the measurements via the use of windows, we are able to avoid including practically unlikely or impossible situations.
+For instance, in our model to predict the number of joggers in a day, temperature and time of year are likely highly correlated.
+By limiting our window in our ALE plot, we can ensure that we're not factoring in a scenario where the time of year is winter and the temperature is 35 degrees Celsius.
+We shouldn't be factoring in the model's prediction for this situation because it is certainly outside of any data set we will be using.
+ALE plots are able to avoid such situations and give us much more accurate results.
+
+####  (ignore below)
 
@@
 
-#### Model Results
+Model Results
 ```{r}
@@
 
-#### ALE Plots
-
-So I was using partial dependence plots to visualize the variables in RF, but then found out you shouldn't use PDP when the variables are correlated with other variables it's one of the assumptions of PDP https://christophm.github.io/interpretable-ml-book/pdp.html ; https://arxiv.org/pdf/1612.08468.pdf).
-Instead, I should be using Accumulated Local Effects (ALE) plots (https://christophm.github.io/interpretable-ml-book/ale.html).
-May also want to look at ICE plots (https://towardsdatascience.com/an-overview-of-model-explainability-in-modern-machine-learning-fc0f22c8c29a)
-
-- code example of how to make ALE using iml package:
-  - https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
-  - https://cran.r-project.org/web/packages/iml/vignettes/intro.html
-
-- webpage explaining ALE: https://christophm.github.io/interpretable-ml-book/ale.html
-- github code for webpage above: https://github.com/christophM/interpretable-ml-book/blob/6153b2ac0143f879bd4a856a1048545a40cd1a74/manuscript/05.4-agnostic-ale.Rmd
-
-other sources (1-2 journal article)
-- https://towardsdatascience.com/an-overview-of-model-explainability-in-modern-machine-learning-fc0f22c8c29a
-- https://www.moodysanalytics.com/-/media/presentation/2019/mas-t1d1-transparent-and-interpretable-ml-in-risk-modeling.pdf
-- https://www.enjine.com/blog/interpreting-machine-learning-models-accumulated-local-effects/ --> it would be cool to be able to make the plot in this website
-- https://arxiv.org/pdf/1612.08468.pdf
-- https://cran.r-project.org/web/packages/iml/iml.pdf
-- https://cran.r-project.org/web/packages/iml/vignettes/intro.html
-- https://cran.r-project.org/web/packages/ALEPlot/ALEPlot.pdf (iml package and documentation a lot more useful)
-
-So, I just wanted to get the code started.
-I think you will be better at streamlining makingthe rest of the graphs, especially with the same axis scales.
-One thing I did notice is that ALE plots for the same mortality class resulted in
-
-```{r}
-
-## sample code to make ALE plos (using solar radiation as example) used combo of: https://cran.r-project.org/web/packages/iml/vignettes/intro.html and https://cran.r-project.org/web/packages/iml/iml.pdf
-
-#do something with the response variable column (removes it but R can still recognize it)
-train_x <- train[-which(names(train) == "mortality_qual")]
-
-# this does something to create values that will be plotted in the ALE plot: inputs include the random forest model (rfm) run above, the trainig data with the modified response column (created above), specifying we want output to be a probability, looking at a specific mortality class
-sr_predictor_high <- Predictor$new(rfm, data = train_x, type = "prob", class = "High")
-
-# plot the outputs in an ALE plot using the predictor info created above just for solar radiation variable(feature = "explanatory variable").
-Doesnt work if do featureeffects$new() and then plot() on separate lines even though thats how one of the smaple codes showed it.
-sr_high <- plot(FeatureEffect$new(sr_predictor_high, feature = "sr")) + ggtitle("High")
-
-#predict values for the other mortality classes for the same variable
-sr_predictor_moderate <- Predictor$new(rfm, data = train_x, type = "prob", class = "Moderate")
-sr_moderate <- plot(FeatureEffect$new(sr_predictor_moderate, feature = "sr")) + ggtitle("Moderate")
-
-sr_predictor_low <- Predictor$new(rfm, data = train_x, type = "prob", class = "Low")
-sr_low <- plot(FeatureEffect$new(sr_predictor_low, feature = "sr")) + ggtitle("Low")
-
-sr_predictor_none <- Predictor$new(rfm, data = train_x, type = "prob", class = "None")
-sr_none <- plot(FeatureEffect$new(sr_predictor_none, feature = "sr")) + ggtitle("None")
-
-ggarrange(sr_high, sr_moderate, sr_low, sr_none, ncol = 2, nrow = 2)
-
-#plot all variables(for just one mortality class) : for some reason the function makes graphs for all the variables in the data frame, not just the ones in the random forest model. 
-tt <- FeatureEffects$new(sr_predictor_high, method = "ale")
-plot(tt)
-
-# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
-all_rfm_variables <- FeatureEffects$new(sr_predictor_low, method = "ale", features = c("cwd", "rain", 'tmax' , 'TPI_five' , 'dem_10m_zaca' , 'flow_accumulation' , 'sr' ,'slope', 'TWI', 'vpd_aug'))
-all_rfm_variables$plot()
-ggsave("ale_low.jpg", height = 9, width = 9)
-
-# plot a variables in ALE plot: https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
-plot(FeatureEffect$new(sr_predictor_low, feature = c("cwd", "dem_10m_zaca")))
-
-# another way to make ale plts with iml package (doesnt work): https://medium.com/the-die-is-forecast/shining-a-light-on-the-black-box-of-machine-learning-2b49fe471cee
-sr_predictor_low <- Predictor$new(rfm, data = train_x, type = "prob", class = "Low")
-ale_iml <- FeatureEffect$new(sr_predictor_low, feature = "sr")
-sr_df <- ale_iml$results
-ggplot(sr_df, aes(x = sr, y = .ale)) + geom_line() # not sure what .ale is. the samle code doesn't say, but the above codes work so whatever
-
-# another way to write the code: https://github.com/christophM/interpretable-ml-book/blob/6153b2ac0143f879bd4a856a1048545a40cd1a74/manuscript/05.4-agnostic-ale.Rmd
-FeatureEffects$new(sr_predictor_low, feature = "sr", method = "ale")$plot() +ggtitle("Low") #produces the same graph as above
-
-```
-- get error when try to run the model on the test data instead of the train data. Think the data used in the model, needs to be the same data to do ALE plots.
-
-- ale plots let us determine the effect that each individual input, isolated/regardless of all others, has on our output.
-
-- How to interpret ALE plots: not really sure since the papers/tutorials I've seen show how to interpret grpahs for quantitative response variables, but I assume since we indicate "type = prob" in the code that the y axis is the probability of a stand having that mortality class at specific value of the x variable
-
-- By localizing the measurements via the use of windows, we are able to avoid including practically unlikely or impossible situations.
-For instance, in our model to predict the number of joggers in a day, temperature and time of year are likely highly correlated.
-By limiting our window in our ALE plot, we can ensure that we're not factoring in a scenario where the time of year is winter and the temperature is 35 degrees Celsius.
-We shouldn't be factoring in the model's prediction for this situation because it is certainly outside of any data set we will be using.
-ALE plots are able to avoid such situations and give us much more accurate results.
-
 ### Random Forest Model: 3 mortality classes
@@
 
-#### ALE Plots
+ALE Plots
 ```{r}
@@
 
-#### ROC and AUC
+ROC and AUC
 ```{r}

### 2020-07-09 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
final-ish draft of preliminary model runs

@@
   group_by(mortality_qual) %>% 
-  count()
+  count() %>% 
+  rename(manual_mortality=mortality_qual)
 
@@
   group_by(rdnbr_qualitative) %>% 
-  count()
+  count() %>% 
+  rename(mcf_rdnbr = rdnbr_qualitative)
 
@@
   group_by(dnbr_qualitative) %>% 
-  count()
+  count() %>% 
+  rename(mcf_dnbr = dnbr_qualitative)
 
@@
   group_by(rdnbr_qualitative) %>% 
-  count()
+  count() %>% 
+  rename(pico_rdnbr =rdnbr_qualitative)
 
@@
   group_by(dnbr_qualitative) %>% 
-  count()
+  count() %>% 
+  rename(pico_dnbr = dnbr_qualitative)
 
@@
 
-# all
+# mcf
 ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
@@
 
-#by prefire cover
+##by prefire cover
 ggplot(mcf, aes(x = tc_mortality, y = rdnbr_2007)) +
@@
 
-## use lm to get a rdnbr threshold for the type converter stands? ---------------------
+## use lm to get a rdnbr threshold for the type converter stands? pretty sure thats not what you meant chris ---------------------
 
@@
 # potential type conversion threshold --------------------------------------------
-
+## pretty sure i did NOT followed the correct steps to pick this threshold. 
 tc_threshold <- mcf %>% 
@@
     rdnbr_2007 >1160 ~ "tc"
-  ))
+  )) %>% 
+  mutate(rdnbr_tc_threshold = as.factor(rdnbr_tc_threshold))
+
+tc_threshold_count <- tc_threshold %>% 
+  group_by(rdnbr_tc_threshold) %>% 
+  count() # not tc = 702, tc = 33
 
@@
 
-rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi, 
                     data = mcf, 
@@
 
-### RF manual: type conversion
-Just using for loop to generate several graphs so its easier to compare them and make sure importance is stable
-
-0 = not type converted
-1 = type converted
-```{r}
-
-for (i in 1:10) {
-
-rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
-                    data = mcf, 
-                    sampsize = c('1' = 99, '0' = 99), 
-                    ntree=1500, 
-                    importance=T,
-                    replace = T)
-# OOB and importance
-rfm
-importance(rfm)
-#plot to visualize variable importance
-varImpPlot(rfm)
-
-}
-
-```
-variable importance is KINDA/MOSTLY stable.
-There arent big changes between runs, some variables flip 1 space, but some move 3 spaces.
-Variable importance was more stable when ndvi was included in the model.
-
-ROC and AUC
-```{r}
-# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
-rfm_roc <-  roc(mcf$tc_mortality, rfm$votes[,1])
-plot(rfm_roc) # plot comes out a little weird
-auc(rfm_roc)
-```
-
-PDP
-```{r}
-#PDP
-a <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "0") #0=not type converted
-b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "1", col="red") #1=type converted
-lines(a, col="green")
-
-a <- partialPlot(rfm, as.data.frame(mcf), slope, "0") #0=not type converted
-b <- partialPlot(rfm, as.data.frame(mcf), slope, "1", col="red") #1=type converted
-lines(a, col="green")
-
-a <- partialPlot(rfm, as.data.frame(mcf), sr, "0") #0=not type converted
-b <- partialPlot(rfm, as.data.frame(mcf), sr, "1", col="red") #1=type converted
-lines(a, col="green")
-
-```
-so looking at the 3 most important variables for type conversino plots, all of the probabilities are below zero.
-only the not type converted stands have probabilities above 0, with some values of the explanatory variable having very high probabilities.
-So what this tells me is the model CANNOT distinguish type converted stands from non type converted stands.
-
-ALE
-```{r}
-#sub data
-mcf_ale <- mcf %>% 
-  select(tc_mortality, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
-
-train_x <- mcf_ale[-which(names(mcf_ale) == "tc_mortality")]
-
-# predictors 
-nottc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$tc_mortality, type = "prob", class="0")
-tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$tc_mortality, type = "prob", class="1")
-
-#vpd_jan
-other <- plot(FeatureEffect$new(nottc_predictor, feature = "vpd_jan")) + ggtitle("Not Type Converted")
-high <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_jan")) + ggtitle("Type Converted")
-ggarrange(other, high)
-
-#slope
-other <- plot(FeatureEffect$new(nottc_predictor, feature = "slope")) + ggtitle("Not Type Converted")
-high <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted")
-ggarrange(other, high)
-
-#sr
-other <- plot(FeatureEffect$new(nottc_predictor, feature = "sr")) + ggtitle("Not Type Converted")
-high <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted")
-ggarrange(other, high)
-
-```
-
 # rdnbr: classification (2 classes)
@@
 high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High")
+ggarrange(other, high)
+
+```
+
+# Type conversion: manual mortality (classification)
+
+0 = not type converted
+1 = type converted
+```{r}
+
+for (i in 1:10) {
+
+rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi, 
+                    data = mcf, 
+                    #sampsize = c('1' = 99, '0' = 99), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+
+```
+
+ROC and AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(mcf$tc_mortality, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+```
+
+PDP
+```{r}
+#PDP
+a <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "0") #0=not type converted
+b <- partialPlot(rfm, as.data.frame(mcf), vpd_jan, "1", col="red") #1=type converted
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(mcf), slope, "0") #0=not type converted
+b <- partialPlot(rfm, as.data.frame(mcf), slope, "1", col="red") #1=type converted
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(mcf), sr, "0") #0=not type converted
+b <- partialPlot(rfm, as.data.frame(mcf), sr, "1", col="red") #1=type converted
+lines(a, col="green")
+
+```
+so looking at the 3 most important variables for type conversino plots, all of the probabilities are below zero.
+only the not type converted stands have probabilities above 0, with some values of the explanatory variable having very high probabilities.
+So what this tells me is the model CANNOT distinguish type converted stands from non type converted stands.
+
+ALE
+```{r}
+#sub data
+mcf_ale <- mcf %>% 
+  select(tc_mortality, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- mcf_ale[-which(names(mcf_ale) == "tc_mortality")]
+
+# predictors 
+nottc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$tc_mortality, type = "prob", class="0")
+tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$tc_mortality, type = "prob", class="1")
+
+#vpd_jan
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "vpd_jan")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_jan")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+#slope
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "slope")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+#sr
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "sr")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+```
+
+# Type conversion: rdnbr threshold (classification)
+
+preliminary threshold: 1160
+```{r}
+
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_tc_threshold ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi, 
+                    data = tc_threshold, 
+                    #sampsize = c('1' = 99, '0' = 99), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+```
+ROC and AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(tc_threshold$rdnbr_tc_threshold, rfm$votes[,1])
+plot(rfm_roc) # plot comes out a little weird
+auc(rfm_roc)
+```
+
+PDP
+```{r}
+#PDP
+a <- partialPlot(rfm, as.data.frame(tc_threshold), rain, "not_tc") 
+b <- partialPlot(rfm, as.data.frame(tc_threshold), rain, "tc", col="red") 
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(tc_threshold), slope, "not_tc") 
+b <- partialPlot(rfm, as.data.frame(tc_threshold), slope, "tc", col="red") 
+lines(a, col="green")
+
+a <- partialPlot(rfm, as.data.frame(tc_threshold), vpd_aug, "not_tc") 
+b <- partialPlot(rfm, as.data.frame(tc_threshold), vpd_aug, "tc", col="red") 
+lines(a, col="green")
+
+```
+ALE
+```{r}
+#sub data
+tc_threshold_ale <- tc_threshold %>% 
+  select(rdnbr_tc_threshold, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed, ndvi) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+
+train_x <- tc_threshold_ale[-which(names(tc_threshold_ale) == "rdnbr_tc_threshold")]
+
+# predictors 
+nottc_predictor <- Predictor$new(rfm, data = tc_threshold_ale, y = tc_threshold_ale$rdnbr_tc_threshold, type = "prob", class="not_tc")
+tc_predictor <- Predictor$new(rfm, data = tc_threshold_ale, y = tc_threshold_ale$rdnbr_tc_threshold, type = "prob", class="tc")
+
+#vpd_jan
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "vpd_jan")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_jan")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+#slope
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "slope")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted")
+ggarrange(other, high)
+
+#sr
+other <- plot(FeatureEffect$new(nottc_predictor, feature = "rain")) + ggtitle("Not Type Converted")
+high <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted")
 ggarrange(other, high)

### 2020-07-14 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
minor change

@@
 b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8))
+lines(c, col="green")
 
@@
 b <- partialPlot(rfm, as.data.frame(mcf), slope, "High", col="red", ylim=c(-0.8,0.8)) 
+lines(c, col="green")
 

### 2020-07-26 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
minor changes

@@
   mutate(dnbr_qualitative = case_when(
-    dnbr_2007 < 0.1 ~ "unburned",
-    dnbr_2007 < 0.27 ~ "low",
-    dnbr_2007 < 0.66 ~ "moderate",
-    dnbr_2007 > 0.66 ~ "high")) %>%
+    dnbr_2007 < 0.1 ~ "None",
+    dnbr_2007 < 0.27 ~ "Low",
+    dnbr_2007 < 0.66 ~ "Moderate",
+    dnbr_2007 > 0.66 ~ "High")) %>%
   mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
-         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned"))) %>%
+         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("High", "Moderate", "Low", "None"))) %>%
   mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
@@
                                          levels=c("High", "Moderate", "Low", "None"))) %>%
+  dplyr::select(-Mortality, -type_converted, -ndvi_2006) %>% 
   na.omit(pico) %>%  # 56 stands have NA values (probs all very small stands)
+  #filter(Mortality != "bigcone") %>% # 60 stands actually have big cones not pico
   mutate(rdnbr_qualitative2 = case_when(
@@
   mutate(dnbr_qualitative = case_when(
-    dnbr_2007 < 0.1 ~ "unburned",
-    dnbr_2007 < 0.27 ~ "low",
-    dnbr_2007 < 0.66 ~ "moderate",
-    dnbr_2007 > 0.66 ~ "high")) %>%
+    dnbr_2007 < 0.1 ~ "None",
+    dnbr_2007 < 0.27 ~ "Low",
+    dnbr_2007 < 0.66 ~ "Moderate",
+    dnbr_2007 > 0.66 ~ "High")) %>%
   mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
-         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned")))
+         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("High", "Moderate", "Low", "None")))
 
@@
 East and west are treated identically with transformed aspect values of 0.5.
+
+Need to remove the bigcone stands from the pico dataset but having a little trouble atm.
 
@@
   count() %>% 
-  rename(manual_mortality=mortality_qual)
-
+  #rename(manual_mortality=mortality_qual) %>% 
+  mutate(mortality_type = 'manual')
+
+rdnbr_mcf <- mcf %>% 
+  group_by(rdnbr_qualitative) %>% 
+  count() %>% 
+  rename(mortality_qual = rdnbr_qualitative) %>% 
+  mutate(mortality_type = 'rdnbr')
+
+dnbr_mcf <- mcf %>% 
+  group_by(dnbr_qualitative) %>% 
+  count() %>% 
+  rename(mortality_qual = dnbr_qualitative) %>% 
+  mutate(mortality_type = 'dnbr')
+
+mcf_sample_size <- rbind(mortality_count, rdnbr_mcf, dnbr_mcf)
+
+ggplot(mcf_sample_size, aes(x = mortality_qual, y = n, fill = mortality_type)) + 
+  geom_col(position = "dodge") +
+  geom_text(aes(label=n), position = position_dodge(width=0.9), vjust=-0.5) +
+  scale_fill_discrete(name = "How was mortality determined")
+
+# mcf type conversison ------------------------------------------------------------------- 
 tc_count <- mcf %>% 
@@
   count() # 99 stands with type conversion, 636 with no type conversion
-
-rdnbr_mcf <- mcf %>% 
-  group_by(rdnbr_qualitative) %>% 
-  count() %>% 
-  rename(mcf_rdnbr = rdnbr_qualitative)
-
-dnbr_mcf <- mcf %>% 
-  group_by(dnbr_qualitative) %>% 
-  count() %>% 
-  rename(mcf_dnbr = dnbr_qualitative)
 
@@
   count() %>% 
-  rename(pico_rdnbr =rdnbr_qualitative)
+  rename(mortality_qual =rdnbr_qualitative) %>% 
+  mutate(mortality_type = 'rdnbr')
 
@@
   count() %>% 
-  rename(pico_dnbr = dnbr_qualitative)
-
-# all -------------------------------------------------------------------
-all_sample_size <- cbind(mortality_count, rdnbr_mcf, dnbr_mcf, rdnbr_pico, dnbr_pico)
-```
-Yah, dnbr is not a good index to use. Rdnbr is better.
-
-### mortality v rdnbr 
-```{r}
-rdnbr_qual_n <- mcf %>% 
-  group_by(rdnbr_qualitative) %>% 
-  count() %>% 
-  mutate(mort_type = "rdnbr") %>% 
-  rename(mortality_qual = rdnbr_qualitative)
-
-mortality_n <- mcf %>% 
-  group_by(mortality_qual) %>% 
-  count() %>% 
-  mutate(mort_type = "manual")
-
-mortality_count <- rbind(rdnbr_qual_n, mortality_n)
-
-ggplot(mortality_count, aes(x = mortality_qual, y = n, fill = mort_type)) + geom_col(position = "dodge") +
+  rename(mortality_qual = dnbr_qualitative)%>% 
+  mutate(mortality_type = 'dnbr')
+
+pico_sample_size <- rbind(rdnbr_pico, dnbr_pico)
+
+ggplot(pico_sample_size, aes(x = mortality_qual, y = n, fill = mortality_type)) + 
+  geom_col(position = "dodge") +
   geom_text(aes(label=n), position = position_dodge(width=0.9), vjust=-0.5) +
@@
 ```
+MCF: Yah, dnbr is not a good index to use.
+Rdnbr is better at least when it comes to high severity and high mortality.
+Both dnbr and rdnbr over estimate moderate adn low mortlity and severely underestiamte stands with no mortality.
+Bc of this, we could make the assumption that pico stands with high rdnbr also have high mortality.
 
@@
 
-rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi, 
+rfm <- randomForest(mortality_qual ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi + ndmi_2006 + rdnbr_qualitative,
                     data = mcf, 
@@
 
+plot_fi(rf_feat_importance(rfm, mcf))
+
 }
@@
 mcf_ale <- mcf %>% 
-  select(mortality_qual, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+  select(mortality_qual, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, ndmi_2006, aspect_transformed, ndvi, rdnbr_qualitative) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
 
@@
 
-rfm <- randomForest(two_class_mortality_mod ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+rfm <- randomForest(two_class_mortality_mod ~ rain + tmax + TPI_five + flow_accumulation + sr + slope + TWI  + aspect_transformed + ndvi + ndmi_2006, 
                     data = mcf_2_class_mod, 
@@
 ```
+removed vpd_aug and cwd (rain had greater importance than precip and vpd_jan had greater importance than vpd_aug).
+.Variable importance got worse, so removed vpd_jan bc tmax became more important more often.
+NOW the variable iportance is a lot more stable!
+and lo and behold tmax is more important than slope.
+removed elevation and replaced it with vpd_jan and variabe importance got cray cray again.
+Current variables result in very stable importance values!
 
@@
 mcf_ale <- mcf_2_class_mod %>% 
-  select(two_class_mortality_mod, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+  select(two_class_mortality_mod, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed, ndvi, ndmi_2006) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
 
@@
 #vpd_jan
-other <- plot(FeatureEffect$new(other_predictor, feature = "vpd_jan")) + ggtitle("Other")
-high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_jan")) + ggtitle("High")
+other <- plot(FeatureEffect$new(other_predictor, feature = "ndvi")) + ggtitle("Other")
+high <- plot(FeatureEffect$new(high_predictor, feature = "ndvi")) + ggtitle("High")
 ggarrange(other, high)
@@
 
-rfm <- randomForest(rdnbr_qualitative2 ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed, 
+rfm <- randomForest(rdnbr_qualitative2 ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + aspect_transformed + ndvi + ndmi_2006, 
                     data = mcf, 
@@
 mcf_ale <- mcf %>% 
-  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
+  select(rdnbr_qualitative2, rain, tmax, TPI_five, dem_10m_zaca, flow_accumulation, sr, slope, TWI, cwd, vpd_aug, vpd_jan, aspect_transformed, ndvi, ndmi_2006) # the predictor$new code had issues with using a df with NA values even thought the variabels in the RF do not have NA values.
 
@@
 ```
+interesting. for rdnbr classification, the ale plots show different trends for ndvi. .However, this might change as the explanatory variables change
 
@@
 FeatureEffects$new(sr_predictor_low, feature = "sr", method = "ale")$plot() +ggtitle("Low") #produces the same graph as above
+
+# categorical variable 
+none <-  plot(FeatureEffect$new(none_predictor, feature = "rdnbr_qualitative")) + ggtitle("None")
+low <- plot(FeatureEffect$new(low_predictor, feature = "rdnbr_qualitative")) + ggtitle("Low")
+moderate <- plot(FeatureEffect$new(moderate_predictor, feature = "rdnbr_qualitative")) + ggtitle("Moderate")
+high <- plot(FeatureEffect$new(high_predictor, feature = "rdnbr_qualitative")) + ggtitle("High")
+ggarrange(none, low, moderate, high)
+
+# plot 2 variables
+none <-  plot(FeatureEffect$new(none_predictor, feature = c("tmax", "rain"))) + ggtitle("None")
+low <- plot(FeatureEffect$new(low_predictor, feature = c("tmax", "rain"))) + ggtitle("Low")
+moderate <- plot(FeatureEffect$new(moderate_predictor, feature = c("tmax", "rain"))) + ggtitle("Moderate")
+high <- plot(FeatureEffect$new(high_predictor, feature = c("tmax", "rain"))) + ggtitle("High")
+ggarrange(none, low, moderate, high)
 

### 2020-07-30 Model Data Sheets and Code/Manuscript_Code.Rmd
New document with final code for manuscript

---
title: "Zaca Fire Type Conversion"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Prepare Workspace

```{r load_libraries}

library(tidyverse)
library(raster)
library(VSURF)
library(ALEplot)

setwd("workingdirectory") #change this to your local directory

set.seed(2020) #Set seed so random forests are replicable

```

### Load Shapefiles

```{r load_shapefiles}

zaca <- shapefile("E:/zaca/chris/zaca_scar")

mixed_conifer <- shapefile("all_mixed_pine.shp")

coulter <- shapefile("coulter_pines.shp") %>% 
  spTransform(., crs(mixed_conifer))

```

### Prepare Environmental Variables

```{r}

dem <- raster("E:/zaca/environmental_data/all_variables/dem_10m_zaca.tif")

vpd_jan <- raster("E:/zaca/environmental_data/cwd_vpd_amp/PRISM_vpdmax_30yr_normal_800mM2_01_asc/PRISM_vpdmax_30yr_normal_800mM2_01_asc.asc") %>% 
  projectRaster(., dem_30m) %>% 
  crop(., zaca) %>% 
  mask(., zaca)

#TPI
#Insolation
#Temp
#Rainfall

```

## Mixed Conifer Forest

### Extract Values

```{r}

mixed_conifer_env <- extract()

```

### Stepwise Variable Selection

```{r}

#Extract prediction and response variables
mcf.predictors <- mixed_conifer_env[,2:ncol(mixed_conifer_env)]
mcf.response <- mixed_conifer_env[,1]

#Run stepwise variable selection algorithm
mcf.vsurf <- VSURF(x = predictors, y = response)

#Get names of selected variables
names(mcf.predictors[mcf.vsurf$varselect.pred])

```

### Random Forest

```{r}

#Create random forest using selected variables
mixed_conifer_rf <- randomForest(RdNBR_2007 ~ var1 + var2 + var3, 
                                 data = mixed_conifer_env, 
                                 importance = TRUE, 
                                 na.action = na.omit)

```

### 2020-08-02 Model Data Sheets and Code/Manuscript_Code.Rmd
Calculate fire history

@@
 
-### Prepare Workspace
+## Prepare Workspace
 
@@
 library(raster)
+library(randomForest)
 library(VSURF)
@@
 
-setwd("workingdirectory") #change this to your local directory
+setwd("E:/zaca/type_conversion")
 
@@
 
-```
-
-### Load Shapefiles
-
-```{r load_shapefiles}
-
-zaca <- shapefile("E:/zaca/chris/zaca_scar")
-
-mixed_conifer <- shapefile("all_mixed_pine.shp")
-
-coulter <- shapefile("coulter_pines.shp") %>% 
-  spTransform(., crs(mixed_conifer))
+UTM_11N <- CRS("+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ")
 
@@
 
-### Prepare Environmental Variables
+## Prepare Vector Data
+
+### Fire History
+
+#### Load Fire History Data
 
 ```{r}
+
+#Load Fires
+firep19_1 <- shapefile("./fire19_1/firep19_1.shp")
+
+#Rename Columns
+names(firep19_1@data)[which(names(firep19_1@data) == "ALARM_DATE")] <- "START_DATE"
+names(firep19_1@data)[which(names(firep19_1@data) == "CONT_DATE")] <- "END_DATE"
+names(firep19_1@data)[which(names(firep19_1@data) == "YEAR_")] <- "YEAR"
+
+#Load Prescribed Burns
+rxburn19_1 <- shapefile("./fire19_1/rxburn19_1.shp")
+
+#Rename Columns
+names(rxburn19_1@data)[which(names(rxburn19_1@data) == "TREATMENT1")] <- "FIRE_NAME"
+names(rxburn19_1@data)[which(names(rxburn19_1@data) == "YEAR_")] <- "YEAR"
+
+#Combine Layers
+fire_history <- bind(firep19_1, rxburn19_1)
+
+#Change Column Data Types
+fire_history$YEAR <- as.numeric(fire_history$YEAR)
+
+#Select Columns
+fire_history <- fire_history[, c("FIRE_NAME", "YEAR")]
+
+```
+
+#### Calculate Years Since Fire
+
+```{r}
+
+#Extract Zaca burn scar
+zaca <- subset(fire_history, FIRE_NAME == "ZACA" & YEAR == 2007)
+
+#Extract fires before Zaca
+fires_before_zaca <- subset(fire_history, YEAR < 2007)
+
+#Calculate years since last burn
+fires_within_zaca_scar <- intersect(zaca, fires_before_zaca)
+fires_within_zaca_scar$years_since_fire <- fires_within_zaca_scar$YEAR.1 - fires_within_zaca_scar$YEAR.2
+
+#Reproject to UTM Zone 11N
+fires_within_zaca_scar <- spTransform(fires_within_zaca_scar, UTM_11N)
+
+```
+
+#### Reproject Zaca Boundary
+
+```{r}
+
+zaca <- spTransform(zaca, UTM_11N)
+
+```
+
+### Vegetation Cover
+
+```{r}
+
+south_coast_vegetation <- shapefile("RVMid_R05_SouCoast.shp") %>% 
+  crop(., zaca)
+
+central_coast_vegetation <- shapefile("RVMid_R05_CentralCoast.shp") %>% 
+  crop(., zaca)
+
+zaca_vegetation <- merge(south_coast_vegetation, central_coast_vegetation)
+
+```
+
+## Prepare Environmental Variables
+
+## Create Raster Templates
+
+The templates have the correct projection, extent, origin, and spatial resolution for the environmental variables.
+The environmental variables will be reprojected to line up with the templates.
+The pixel grid is derived from the Landsat images.
+The topographic variables and insolation will be analyzed at a 10 m spatial resolution.
+The other variables will be analyzed at a 30 m spatial resolution.
+
+```{r}
+
+template_30m <- raster(landsat_2006[[1]]) %>% 
+  crop(., y = zaca)
+
+template_10m <- raster(landsat_2006[[1]]) %>% 
+  crop(., y = zaca) %>% 
+  disaggregate(., fact = 3)
+
+```
+
+### Prepare Digital Elevation Model
+
+```{r dem}
+
+dem120 <- raster("C:/Users/Christopher/Downloads/USGS_13_n35w120.tif")
+dem121 <- raster("C:/Users/Christopher/Downloads/USGS_13_n35w121.tif")
+
+usgs_dem <- merge(dem120, dem121) %>% 
+  projectRaster(., to = template_10m, method = "bilinear")
+
+```
+
+### Prepare Topography Data
+
+```{r topography}
+
+```
+
+### Prepare Climate Data
+
+```{r climate}
 

### 2020-08-03 Model Data Sheets and Code/Manuscript_Code.Rmd
Landsat pre-processing

@@
 UTM_11N <- CRS("+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ")
+
+```
+
+## Prepare Imagery Data
+
+### Load Function to Pre-process Landsat
+
+The Landsat pre-processing function is stored in a separate file.
+For each image, it stacks the Landsat bands into a single image file.
+It then masks erroneous reflectance retrievals, masks clouds, and saves the new image.
+The function automatically iterates over every data set in a subdirectory.
+
+```{r}
+
+source("preprocessLandsat.R")
+
+```
+
+### Pre-process Landsat Images
+
+```{r}
+
+preprocessLandsat("E:/zaca/type_conversion/Imagery")
 

### 2020-08-03 Model Data Sheets and Code/preprocessLandsat.R

@@ -1,0 +1,91 @@
+preprocessLandsat <- function(inputPath, 
+                         allBands = TRUE, 
+                         maskClouds = TRUE, 
+                         writeFile = TRUE, 
+                         toMemory = FALSE, 
+                         fileFormat = "GTiff"){
+  
+  require(raster)
+  
+  setwd(inputPath)
+  
+  ### Iterate over every image subdirectory
+  
+  for (dir in list.dirs(path = ".", full.names = FALSE)[nchar(list.dirs(path = ".", full.names = FALSE)) > 0]){
+    
+    setwd(paste(inputPath, "/", dir, sep = ""))
+    
+    ### Create Raster Brick
+    
+    #List image files that are actually Landsat bands
+    band_list <- list.files(path = ".", glob2rx("*band*.tif$"))
+    
+    #Drop aerosol band for Landsat 8
+    if(substr(band_list[1], 1,4) == "LC08"){
+      band_list <- band_list[-1]
+      print("Landsat 8")
+    } else {print("Not Landsat 8")}
+    
+    print(paste("Processing ", substr(band_list[1], 18, 25), sep = ""))
+    
+    print(band_list)
+    
+    #Merge bands into a brick
+    image_brick <- stack(band_list)
+    
+    #Set NA value for output
+    NAvalue(image_brick) <- -9999
+    
+    ### Overwrite Invalid Pixel Values
+    
+    #Set all bad retrievals to NA
+    image_brick <- reclassify(image_brick, c(-Inf, 0, NA))
+    image_brick <- reclassify(image_brick, c(10000, Inf, NA))
+    
+    #Set all bands to NA if any band is NA for a pixel
+    if(allBands == TRUE){
+      
+      vals <- values(image_brick)
+      vals[rowSums(is.na(vals)) > 0, ] <- NA
+      image_brick <- setValues(image_brick, vals)
+      
+    }
+    
+    ### Mask Clouds
+    
+    if(maskClouds == TRUE){
+      
+      pixel_qa_band_name <- list.files(path = ".", glob2rx("*pixel_qa*.tif$"))
+      
+      pixel_qa_band <- raster(pixel_qa_band_name)
+      
+      #Mask medium and high confidence clouds
+      image_brick[pixel_qa_band %in% c(130, 132, 136, 144, 160, 176, 224)] <- NA
+      
+    }
+    
+    ### Export Output
+    
+    setwd(inputPath)
+    
+    if(writeFile == TRUE){
+      
[…]

### 2020-08-04 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
minor changes

@@
 1. Can the mcf model distinguish between type conversion and the other mortality types? 
-  - so far, no! so variable importance is not meaningful for that model
-
-2. Which variables contribute to type conversion of MCFs?
+    yes, if we use ALE plots
+
+2. Which variables contribute to type conversion of MCFs and pico?
 
 3. How well does rdnbr compare to manual estimates of mortality? 
+    high rdnbr matches up very well with high mortality. There is less of a match up between moderate, low,      and no mortality with the corresponding rdnbr severity levels.
 
 4. How does using rdnbr and manually derived mortality compare as a response variable?
-
-5. Which variables contribute to high/low mortality? high/low rdnbr?
-
-5. Does conifer type contribute to different variable importances?
+    this would be a better question for the big cone model
+
+5. Which variables contribute to high/low/type conversion rdnbr? Is there a distinguishable difference between values of the explanatory variables for high/type conversion levels of rdnbr?
+
+6. Does conifer type contribute to different variable importances?
 

### 2020-08-08 chi_sq for vegetation and seedling age classes.Rmd
analysis for field sruvey results section

---
title: "Field Surveys"
author: "Anne-Marie Parkinson"
date: "August 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# create data frame
df <- data.frame(age_class = c("0-3", "4-6", "7-10"),
                 cepa = c(1, 6, 41),
                 chaparral = c(12, 13, 27))

# make contigency table: getting the data in a format that will be used in chi.sq()
df_contigency <- df %>% 
  column_to_rownames('age_class')

# run chi-square test
df_chi <- chisq.test(df_contigency)
df_chi # pvalue = 0.000665

# get Standardized residuals
df_chi$stdres
```
The numbers reported in the above data frame come from the manuscript, which I calcualted from excel sheets.
cepa = the number of recorded sapling/seedlings found in ceanothus palmeri in each corresponding age class 
chaparral = the number of recorded sapling/seedlings found in mixed chaparral in each corresponding age class

Standardized residuals > |2| indicates strong divergence from null hypothesis scenario. So There are significantly more seedlings/saplings in the 7-10 age class that were found in ceanothus palmeri, But both vegetation types had a majority of seedling/saplings in the 7-10 age class. And there were significantly more seedlings/saplings in the 0-3 age class that were found in mixed chaparral.

($\chi$^2^(`r df_chi$parameter`) = `r round(df_chi$statistic,2)`, *p* = `r round(df_chi$p.value,2)`)

chris, later we can combine this with the final manuscript code.

### 2020-08-17 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
vsurf and model selection

@@
   mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, 
-                                     levels=c("Other", "High")))
+                                     levels=c("Other", "High")))  %>% 
+  mutate(rdnbr_tc = case_when(
+    rdnbr_2007 < 1000 ~ "Not_TC",
+    rdnbr_2007 >= 1000 ~ "TC"
+  )) %>% 
+  mutate(rdnbr_tc = as.factor(rdnbr_tc)) %>% 
+  mutate(rdnbr_tc= fct_relevel(rdnbr_tc, 
+                                     levels=c("Not_TC", "TC")))
 
@@
 
-# mcf type conversison ------------------------------------------------------------------- 
+# mcf manual type conversison ------------------------------------------------------------------- 
 tc_count <- mcf %>% 
@@
   count() # 99 stands with type conversion, 636 with no type conversion
+
+# mcf type conversion with rdnbr > 1000 threshold --------------------------------------------
+tc2_coun <- mcf %>% 
+  group_by(rdnbr_tc) %>% 
+  count() #not tc = 680, tc = 55
 
@@
 Bc of this, we could make the assumption that pico stands with high rdnbr also have high mortality.
+
+### Stepwise Variable Selection
+
+```{r}
+
+mcf_sub <- mcf %>% 
+  dplyr::select(rdnbr_qualitative,flow_accumulation, sr, rain, slope, tmax, TPI_five, TWI, fri, aspect_transformed, cwd, ndvi, vpd_aug, vpd_jan,ndmi_2006, dem_10m_zaca)
+
+#Extract prediction and response variables
+mcf.predictors <- mcf_sub[,2:ncol(mcf_sub)]
+mcf.response <- mcf_sub[,1]
+
+#Run stepwise variable selection algorithm
+mcf.vsurf <- VSURF(x = mcf.predictors, y = mcf.response)
+
+#Get names of selected variables
+names(mcf.predictors[mcf.vsurf$varselect.pred])
+
+# does vsurf selected variables improve variance explained and result in stabilized variable importance ranking?
+
+for (i in 1:7) {
+
+rfm <- randomForest(rdnbr_tc ~ rain + tmax + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ndvi + ndmi_2006 + aspect_transformed + vpd_jan, 
+                    data = mcf, 
+                    sampsize = c('Not_TC' = 55, 'TC' = 55), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+rfm_roc <-  roc(mcf$rdnbr_tc, rfm$votes[,1])
+auc(rfm_roc)
+
+```
+all variables: rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ndvi + ndmi_2006 + vpd_jan + aspect_transformed
+
+vsurf variable selection (mcf, regression): slope + sr + vpd_aug + tmax + ndmi_2006 + vpd_jan + ndvi + rain + dem_10m_zaca
+
+vsurf variables: % variance explained = 50%, VI ranking is NOT stable, msr = 44,577
+all variables: % variance explained = 47%, VI ranking is NOT stable, msr = 47,592
+all variables (minus elevation): % variance explained = 47%, VI ranking is mostly stable
+all variables (minus vpd_jan and elevation): % variance explained = 44%, VI ranking IS very stable
+all variables (minus aspect, tpi, twi): % variance explained = 50%, VI ranking is not that stable, msr = 44,699
+all variables (minus aspect, tpi, twi, vpd_jan): % variance explained = 49%, VI ranking is kinda stable
+all variables (minus aspect, tpi, twi, elevation): % variance explained = 49%, VI ranking is kinda stable
+
+vsurf variable selection (mcf, classification, rdnr_tc): vpd_aug + slope + rain + vpd_jan + ndvi
+
+vsurf variables: oob = 14%, not tc class error = 13%, tc class error = 25%, VI ranking IS stable, auc = 89%
+all variables: oob = 15%, not tc class error = 14%, tc class error = 31%, VI ranking is NOT stable, auc = 86%
+all variables (minus elevation, tpi, and vpd_jan): oob = 15%, not tc class error = 14%, tc class error = 35%, VI ranking IS stable, auc = 84%
+all variables(minus elevation, tpi): oob = 15%, not tc class error = 13%, tc class error = 31%, VI ranking is not stable, auc = 86%
+
+vsurf variable selection (mcf, classification, rdnbr 4 mortality classes): sr + ndmi_2006 + vdp_aug + dem_10m_zaca + ndvi + vpd_jan + tmax + cwd + rain + flow_accumulation
+
+### Stepwise Variable Selection - PICO
+
+```{r}
+
+pico_sub <- pico %>% 
+  dplyr::select(flow_accumulation, insolation, rain, slope, tmax, TPI_five, TWI,  aspect_transformed, cwd,  vpd_aug, vpd_jan,ndmi_2006, dem, years_since_burn) # no ndvi
+
+#Extract prediction and response variables
+pico.predictors <- pico_sub[,c(2:14)]
+pico.response <- pico_sub[, 1]
+
+#Run stepwise variable selection algorithm
+pico.vsurf <- VSURF(pico$rdnbr_2007 ~ ., data = pico_sub) #setting up the code the same way as the mcf resulted in an error, but this format also works
+
+#Get names of selected variables
+names(pico_sub[pico.vsurf$varselect.pred])
+
+# does vsurf selected variables improve variance explained and result in stabilized variable importance ranking?
+
+for (i in 1:7) {
+
+rfm <- randomForest(rdnbr_2007 ~ flow_accumulation + insolation + + rain + ndmi_2006 + tmax + dem + vpd_jan + vpd_aug + cwd + years_since_burn, 
+                    data = pico, 
+                    #sampsize = c('Not_TC' = 55, 'TC' = 55), 
+                    ntree=1500, 
+                    importance=T,
+                    replace = T)
+# OOB and importance
+rfm
+importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+rfm_roc <-  roc(mcf$rdnbr_tc, rfm$votes[,1])
+auc(rfm_roc)
+
+```
+VSURF regreession variables selected: flow accumulation + insolation + + rain + ndmi_2006 + tmax + dem + vpd_jan + vpd_aug + cwd + years_since_burn
 

### 2020-09-17 Model Data Sheets and Code/Manuscript_Code.Rmd
minor changes
vsurf and model selection
Calculate RdNBR, import terrain and climate variables

@@
 preprocessLandsat("E:/zaca/type_conversion/Imagery")
+
+```
+
+### Calculate NBR
+
+```{r}
+
+#Directory will all pre-processed Landsat images
+setwd("E:/zaca/type_conversion/Imagery/Level2")
+
+for(img in list.files(path = ".", "GTiff$")){
+
+  print(paste("Processing ", substr(img, 18, 25), sep = ""))
+
+  #Merge bands into a brick
+  image <- brick(img)
+
+  NAvalue(image) <- -9999
+
+  #Check sensor ID, mask using QA layer, and then compute NBR using appropriate formula
+  if(as.numeric(substr(img, 3, 4)) == 8){
+
+    print("Recognized Landsat 8")
+
+    #Band designations change, depending on whether the OLI aerosol band is included
+    if(nlayers(image) == 7){nbr <- (image[[5]] - image[[7]])/(image[[5]] + image[[7]])}
+    else if(nlayers(image) == 6){nbr <- (image[[4]] - image[[6]])/(image[[4]] + image[[6]])}
+    else {print("Incorrect number of bands")}
+
+  } else if(as.numeric(substr(img, 3, 4)) == 5 | as.numeric(substr(img, 3, 4)) == 7) {
+
+    print("Recognized Landsat 5 or 7")
+
+    #Band 7 is the sixth band in the brick for TM/ETM
+    if(nlayers(image) == 6){nbr <- (image[[4]] - image[[6]])/(image[[4]] + image[[6]])}
+    else {print("Incorrect number of bands")}
+
+    } else {print("Error recognizing sensor")}
+
+  writeRaster(nbr, filename = paste("NBR_", substr(img, 18, 25), ".tif", sep = ""), format = "GTiff", NAflag = -9999)
+  assign(paste("NBR_", substr(img, 18, 25)), nbr)
+
+  rm(image, image_brick, nbr)
+
+  print("Image completed")
+
+}
+
+```
+
+### Calculate RdNBR
+
+```{r}
+
+NBR_brick <- brick(NBR_2000, NBR_2001, NBR_2002, NBR_2003, NBR_2004, NBR_2005, NBR_2006, NBR_2007) * 1000
+
+baseline_NBR <- calc(NBR_brick[[1:7]], fun = median)
+
+RdNBR_2007 <- (baseline_NBR - NBR_brick[[8]]) / sqrt(abs(baseline_NBR/1000))
 
@@
 
+#dem <- raster("E:/zaca/environmental_data/all_variables/dem_10m_zaca.tif")
+
 ```
@@
 ```{r topography}
+
+slope <- terrain(x = usgs_dem, opt = "slope", unit = "degrees", neighbors = 8)
+
+aspect <- terrain(x = usgs_dem, opt = "aspect", unit = "degrees", neighbors = 8)
+
+tpi <- terrain(x = usgs_dem, opt = "TPI") #Can calculate larger moving windows manually
 
@@
 
-dem <- raster("E:/zaca/environmental_data/all_variables/dem_10m_zaca.tif")
-
-vpd_jan <- raster("E:/zaca/environmental_data/cwd_vpd_amp/PRISM_vpdmax_30yr_normal_800mM2_01_asc/PRISM_vpdmax_30yr_normal_800mM2_01_asc.asc") %>% 
-  projectRaster(., dem_30m) %>% 
+jan_max_vpd_30yr <- raster("E:/zaca/environmental_data/cwd_vpd_amp/PRISM_vpdmax_30yr_normal_800mM2_01_asc/PRISM_vpdmax_30yr_normal_800mM2_01_asc.asc") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
   crop(., zaca) %>% 
@@
 
-#TPI
-#Insolation
-#Temp
-#Rainfall
+jan_min_temp_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/January_Min_Temperature/PRISM_tmin_30yr_normal_800mM2_01_asc/PRISM_tmin_30yr_normal_800mM2_01_asc.asc") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
+  crop(., zaca) %>% 
+  mask(., zaca)
+
+aug_max_temp_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/August_Max_Temperature/PRISM_tmax_30yr_normal_800mM2_08_asc/PRISM_tmax_30yr_normal_800mM2_08_asc.asc") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
+  crop(., zaca) %>% 
+  mask(., zaca)
+
+aug_max_vpd_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/August_Max_VPD/PRISM_vpdmax_30yr_normal_800mM2_08_asc/PRISM_vpdmax_30yr_normal_800mM2_08_asc.asc") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
+  crop(., zaca) %>% 
+  mask(., zaca)
+
+annual_precip_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/Annual_Precipitation/PRISM_ppt_30yr_normal_800mM2_annual_asc/PRISM_ppt_30yr_normal_800mM2_annual_asc.asc") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
+  crop(., zaca) %>% 
+  mask(., zaca)
+
+#Add Insolation
+
+```
+
+### Prepare Weather Data
+
+```{r}
+
+jan_min_temp_2007 <- raster("E:/zaca/type_conversion/Environmental_Variables/Weather/January_Min_Temp/PRISM_tmin_stable_4kmM3_200701_bil/PRISM_tmin_stable_4kmM3_200701_bil.bil") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
+  crop(., zaca) %>% 
+  mask(., zaca)
+
+aug_max_temp_2006 <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/August_Max_Temperature/PRISM_tmax_30yr_normal_800mM2_08_asc/PRISM_tmax_30yr_normal_800mM2_08_asc.asc") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
+  crop(., zaca) %>% 
+  mask(., zaca)
+
+aug_max_vpd_2006 <- raster("E:/zaca/type_conversion/Environmental_Variables/Weather/August_Max_VPD/PRISM_vpdmax_stable_4kmM3_200608_bil/PRISM_vpdmax_stable_4kmM3_200608_bil.bil") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
+  crop(., zaca) %>% 
+  mask(., zaca)
+
+annual_precip_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/Annual_Precipitation/PRISM_ppt_30yr_normal_800mM2_annual_asc/PRISM_ppt_30yr_normal_800mM2_annual_asc.asc") %>% 
+  projectRaster(., template_30m, method = "ngb") %>% 
+  crop(., zaca) %>% 
+  mask(., zaca)
+
+#Calculate precipitation for preceding year?
 
@@
 #Run stepwise variable selection algorithm
-mcf.vsurf <- VSURF(x = predictors, y = response)
+mcf.vsurf <- VSURF(x = mcf.predictors, y = mcf.response)
 

### 2020-09-20 Model Data Sheets and Code/Manuscript_Code.Rmd
added tuning code

@@
 
+# Tuning Parameters
+
+source: https://arxiv.org/pdf/1804.03515.pdf (see digital page 9 for tuneRanger)
+
+replace: dataframe and response variable 
+```{r}
+set.seed(1127)
+
+# run tuneRanger
+mcf_task <- makeClassifTask(data = dataframe, target = "y variable")
+tune_output <- tuneRanger(mcf_task, 
+                          measure = list(mmce), 
+                          num.trees = 1500, 
+                          tune.parameters = c("mtry", "min.node.size"))
+
+# get best parameter values
+tune_output$model
+
+```
+get error when I try to add "replace" as a tuning parameter :(
+
 ### Random Forest

### 2020-09-20 Model Data Sheets and Code/Manuscript_Code.Rmd
minor change

@@
 
+options for the measure function: https://mlr.mlr-org.com/articles/tutorial/measures.html
+
 replace: dataframe and response variable
+
+tunRanger can also be used for ntree, but studies have shown that after a certain number of trees, usually a couple hundred, there is no benefit on accuracy/model fit of increasing ntree, so i fgiure just using a large number will get the same result as whatever number tuneRangere spits out.
 ```{r}
@@
 
+# recommended parameter settings
 # get best parameter values

### 2020-09-26 Field Surveys.Rmd
field survey code

---
title: "Field Survey Analysis"
author: "Anne-Marie Parkinson"
date: "September 25, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

this code generates statistics for the field survey results section of the manuscript. Do note that not all values will be found here. Some were calculated by hand since hte sample size was so small. 
# packages
```{r}
library(here)
library(tidyverse)
library(janitor)
library(kableExtra)
```

# load data
```{r}
pines <- read_csv(here("Model Data Sheets and Code", "pine_surveys_for_R.csv")) %>% 
  clean_names() %>% 
  mutate(number_pines = parse_number(number_pines),
         pico = parse_number(pico),
         pije = parse_number(pije)) %>% 
  mutate(pico = replace_na(pico, 0),
         pije = replace_na(pije, 0),
         cade = replace_na(cade, 0),
         abco = replace_na(abco, 0))
```

# conifer regen
```{r}

################################ total number of saplings ################################ 
tot_sap <- pines %>% 
  summarise(tot = sum(number_pines))

tot_sap %>% kable()

#################################  percent of species ################################

species <- pines %>% 
  pivot_longer("pico": "abco",
               names_to = "species",
               values_to = "number_saplings") %>% 
  group_by(species) %>% 
  summarise(tot_species = sum(number_saplings),
            percent = tot_species/230)

species %>% kable()

################################# number/percent of plots with no saplings ################################ 
num_zero <- pines %>% 
  group_by(number_pines) %>% 
  count() %>% 
  mutate(percent = n/20)

num_zero %>% kable()

################################# number/percent saplings grouped by classes ################################

pines2 <- pines %>% 
  mutate(class = case_when(
    number_pines <= 0 ~ "Zero",
    number_pines <= 10 ~ "<10",
    number_pines <= 20 ~ "<20",
    number_pines >= 20 ~ ">20")) %>% 
  group_by(class) %>% 
  count() %>% 
  mutate(percent = n/20)

pines2 %>% kable()

```

# dominant vegetation cover

count how many plots have a veg type as the dominant veg type
```{r}
# tot primary dominant vegetation
dom <- pines %>% 
  group_by(dominant_vegetation) %>% 
  count() %>% 
  mutate(percent = n/20)

dom %>% kable()

# primary dominant vegetation by regeneration class
dom_class <- pines %>% 
  mutate(class = case_when(
    number_pines <= 0 ~ "Zero",
    number_pines <= 20 ~ "<20",
    number_pines >= 20 ~ ">20")) %>% 
  group_by(class, dominant_vegetation) %>% 
  count() %>% 
  mutate(percent = n/20)

dom_class %>% kable()

# secondary dom vegetation
dom2 <- pines %>% 
  group_by(x2nd_dominant_vegetation_if_any) %>% 
  count() %>% 
  mutate(percent = n/12) # only dividing by 12 bc 8 plots had no secondary dominant vegetation

dom2 %>% kable()

# secondary dominant vegetation by regeneration class
dom2_class <- pines %>% 
  mutate(class = case_when(
    number_pines <= 0 ~ "Zero",
    number_pines <= 20 ~ "<20",
    number_pines >= 20 ~ ">20")) %>% 
  group_by(class, x2nd_dominant_vegetation_if_any) %>% 
  count() %>% 
  mutate(percent = n/12)

dom2_class %>% kable()

```

### 2020-09-26 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
field survey code

@@
 
-mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
+mcf <- read.csv(here::here("Model Data Sheets and Code", 'Zaca MCF variables for model.csv')) %>% 
   mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("High", "Moderate", "Low", "None"))) %>% 
@@
 
+## Correlation
+
+```{r}
+mcf_corr <- mcf %>% 
+  select(dem_10m_zaca:TPI_five)
+correlation <- cor(mcf_corr$dem_10m_zaca, mcf_corr$TPI_five) # this works but can only compare 2 columns at a time
+cor(mcf_corr[dem_10m_zaca:TPI_five]) # wont work :/ 
+corr.test(mcf_corr[dem_10m_zaca:TPI_five]) # neither does this; library(psych)
+sapply(mcf_corr, cor, y=mcf_corr$sr) # kinda works but only compares correlatoins to one column at atime
+with(mcf_corr, cor(mcf_corr[dem_10m_zaca:TPI_five]))
+```
+idk why this aint workin man. I cleaned the environment and used the minimum number of packages/read_csv()  code necessary. its worked before!
+
 ### Stepwise Variable Selection

### 2020-09-27 Model Data Sheets and Code/Manuscript_Code.Rmd
tried to run the manucript code. No go. it also takes FOREVER

@@
 library(randomForest)
-library(VSURF)
-library(ALEplot)
+#library(VSURF)
+library(ALEPlot)
+library(tuneRanger)
 
@@
 #Directory will all pre-processed Landsat images
-setwd("E:/zaca/type_conversion/Imagery/Level2")
+setwd("E:/zaca/type_conversion/Imagery") #changed from: setwd("E:/zaca/type_conversion/Imagery/Level2")
 
-for(img in list.files(path = ".", "GTiff$")){
+for(img in list.files(path = ".", "tif$")){ # changed from for(img in list.files(path = ".", "GTiff$"))
 

### 2020-10-05 Model Data Sheets and Code/Manuscript_Code.Rmd
final code (messy but will clean up later...hopefully)

@@
 library(randomForest)
-#library(VSURF)
+library(VSURF)
 library(ALEPlot)
 library(tuneRanger)
+library(iml)
+library(ggpubr)
 
@@
 
-### Extract Values
+###... Extract Values
 
@@
 
-### Stepwise Variable Selection
+###... Stepwise Variable Selection
 
@@
 #Extract prediction and response variables
-mcf.predictors <- mixed_conifer_env[,2:ncol(mixed_conifer_env)]
-mcf.response <- mixed_conifer_env[,1]
+mcf.predictors <- mcf_model[,2:ncol(mcf_model)]
+mcf.response <- mcf_model[,1]
 
@@
 #Get names of selected variables
+names(mcf.predictors[mcf.vsurf$varselect.thres])
+names(mcf.predictors[mcf.vsurf$varselect.interp])
 names(mcf.predictors[mcf.vsurf$varselect.pred])
@@
 ```
+prediction step, variables selected:all
 
-# Tuning Parameters
+interp step: variables se;ected: "vpd_aug"         "slope"           "rain"            "tmax"            "tmin_jan"        "sr" "aug_tmax2006"    "tmin_jan2007"    "precip_2006"     "aug_max_vpd2006" "ndmi_2006"       "evi_2006"  "tslf"
+
+prediction step, variables selected: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "sr"           "aug_tmax2006"  "tmin_jan2007" "ndmi_2006"    "evi_2006"     "tslf"
+
+so climate variables are performing better than the weather variables (maybe bc the climate data is avail in higher resolution so there is more variability in values?)
+
+removed aug_max_vpd2006 and precip_2006 and reran VSURF
+
+interp step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "tmin_jan2007"
+
+predicition step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan2007"
+
+###... Load data
+
+the above code wouldnt work, so I got evi and min jan temp in the aca mcf mortality model results.rmd and loaded the old files ive been using
+```{r}
+mcf <- read.csv(here::here("Model Data Sheets and Code", 'Zaca MCF variables for model.csv')) %>% 
+  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("High", "Moderate", "Low", "None"))) %>% 
+  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
+  mutate(two_class_mortality = case_when(
+    mortality_qual %in% "High" ~ "High",
+    mortality_qual %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
+  mutate(two_class_mortality = as.factor(two_class_mortality),
+         two_class_mortality= fct_relevel(two_class_mortality, levels=c("NLM", "High"))) %>% 
+  mutate(tc_mortality = as.factor(tc_mortality)) %>% 
+  mutate(rdnbr_qualitative = case_when(
+    rdnbr_2007 < 640 ~ "NLM",
+    rdnbr_2007 < 850 ~ "High",
+    rdnbr_2007 >= 850 ~ "Type Converted")) %>%
+  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
+         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, levels=c("NLM","High", "Type Converted"))) %>% 
+  #   mutate(rdnbr_qualitative2 = case_when(
+  #   rdnbr_qualitative %in% "High" ~ "High",
+  #   rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
+  # mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2),
+  #        rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, levels=c("NLM", "High")))%>% 
+  mutate(rdnbr_tc = case_when(
+    rdnbr_2007 < 850 ~ "Not_TC",
+    rdnbr_2007 >= 850 ~ "TC"
+  )) %>% 
+  mutate(rdnbr_tc = as.factor(rdnbr_tc)) %>% 
+  mutate(rdnbr_tc= fct_relevel(rdnbr_tc, levels=c("Not_TC", "TC")))
+
+mcf_tc_samplesize <- mcf %>% group_by(tc_mortality) %>%count() #type converted n = 99
+mcf_tc_samplesie <- mcf %>% group_by(rdnbr_tc) %>%dplyr::count() # type converted n = 55 (almost too small to use)
+mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::count() # high=66, low= 585, moderate=272, type converted n = 84 
+mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::summarise(acres = sum(Shape_Area/4047)) %>%dplyr::summarise(tot_acres = sum(acres))
+
+ggplot(mcf, aes(x = rdnbr_tc, y = rdnbr_2007)) +geom_jitter(aes(color=tc_mortality))
+ggplot(mcf, aes(x=dnbr_2007, y=rdnbr_2007)) +geom_point(aes(color=tc_mortality)) + geom_hline(yintercept=875)
+ggplot(mcf, aes(ndmi_2006, y=evi_2006)) +geom_point()
+ggplot(mcf, aes(dem_10m_zaca, y=tmin_jan2007)) +geom_point()
+mod <- lm(ndmi_2006~evi_2006, data=mcf) # R2=0.18
+mod1 <- lm(dem_10m_zaca~tmin_jan2007, data=mcf) # R2=0.04
+mod2 <- lm(dem_10m_zaca~cwd, data=mcf) # R2=0.47
+mod3 <- lm(dem_10m_zaca~rain, data=mcf) # R2=0.498
+mod4 <- lm(dem_10m_zaca~tmax, data=mcf) # R2=0.5
+mod5 <- lm(dem_10m_zaca~vpd_aug, data=mcf) # R2=0.56
+mod6 <- lm(tmax~vpd_aug, data=mcf) # R2=0.95
+mod7 <- lm(cwd~vpd_aug, data=mcf) # R2=0.53
+mod8 <- lm(tmax~cwd, data=mcf) # R2=0.49
+
+mcf %>% dplyr::summarise(min = min(tslf),
+                         mean = mean(tslf),
+                         median = median(tslf),
+                         max = max(tslf), 
+                         min_elev = min(dem_10m_zaca),
+                         mean_elev = mean(dem_10m_zaca),
+                         max_elev = max(dem_10m_zaca))
+
+```
+
+###... Tuning Parameters
 
@@
 
-replace: dataframe and response variable
-
-tunRanger can also be used for ntree, but studies have shown that after a certain number of trees, usually a couple hundred, there is no benefit on accuracy/model fit of increasing ntree, so i fgiure just using a large number will get the same result as whatever number tuneRangere spits out.
+tunRanger can also be used to tune for ntree, but studies have shown that after a certain number of trees, usually a couple hundred, there is no benefit on accuracy/model fit of increasing ntree, so i fgiure just using a large number will get the same result as whatever number tuneRangere spits out.
 ```{r}
@@
 
+# select variables to be used in the model
+mcf_model <- mcf %>% 
+  dplyr::select(rdnbr_tc, tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, tslf) # not sure if fri and years so fire, whic sounds like Time Since Last Fire, are the same thing
+
 # run tuneRanger
-mcf_task <- makeClassifTask(data = dataframe, target = "y variable")
+mcf_task <- makeClassifTask(data = mcf_model, target = "rdnbr_tc")
+tune_output <- tuneRanger(mcf_task, 
+                          measure = list(mmce), 
+                          num.trees = 1500, 
+                          tune.parameters = c("mtry", "min.node.size"))
+
+# recommended parameter settings
+# get best parameter values
+tune_output$model
+
+# 4 rdnbr classes
+mcf_model3 <- mcf %>% 
+  dplyr::select(rdnbr_qualitative, tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, tslf) # not sure if fri and years so fire, whic sounds like Time Since Last Fire, are the same thing
+
+# run tuneRanger
+mcf_task <- makeClassifTask(data = mcf_model3, target = "rdnbr_qualitative")
 tune_output <- tuneRanger(mcf_task, 
@@
 
-### Random Forest
+Trials: (2 classes)
+- rdnbr cut off: 850; variables included: tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TPI_five, TWI, ndmi_2006, evi_2006, fri; optimal parameters: mtry=6, min.node.size=4; acc = 82%, not tc = 75%, tc=62%, auc=84%; stable VI: kinda
+Using this one- rdnbr cut off: 850; variables included: removed tpi, mtry=3,min.node.size=7
+- rdnbr cut off: 850; variables included: removed tmin; optimal parameters: mtry=5, min.node.size=10; stable VI: kinda
+- rdnbr cut off: 850; tmax, vpd_aug, rain, sr, slope, aspect_transformed, evi_2006, tslf, tmin_jan2007, ndmi_2006; mtry=3,min.node.size=7; stable vi: sure!
+- rdnbr cut off: 850; variables included: removed cwd; optimal parameters: mtry=4, min.node.size=2; stable VI: NO! horrible stability. 
+- rdnbr cut off: 850; variables included: removed evi_2006; optimal parameters: mtry=4, min.node.size=3; stable VI:kinda
+- rdnbr cut off: 850; variables included: removed ndmi_2006; optimal parameters: mtry=6, min.node.size=27; stable VI:kinda
+
+- rdnbr cut off: 875; variables included: all; optimal parameters: mtry=6, min.node.size=15; stable VI:eh, kinda
+- rdnbr cut off: 875; variables inlcuded: rdnbr_tc, tmax, vpd_aug, rain, tmin_jan, sr, slope, aspect_transformed, TPI_five, TWI, evi_2006, tslf, cwd, tmin_jan2007, ndmi_2006, precip_2006, aug_max_vpd2006, aug_tmax2006; mtry=7,min.node.size=8; stable VI: NO!
+horrible stability.
+- rdnbr cut off: 875; variables inlcuded: removed precip_2006 and aug_max_vpd2006 bc of VSURF results; mtry=6,min.node.size=5
+
+- rdnbr cut off: 875; variables inlcuded:tmax, vpd_aug, rain, sr, slope, aspect_transformed, TPI_five, TWI, evi_2006, tslf, cwd, tmin_jan2007, ndmi_2006; mtry=4,min.node.size=3; stable VI: NO!
+horrible stability.
+
+4 classes
+mtry=12,min.node.size=5; stable VI!!; oob not great, TC has lots of confusion with moderate class
+
+3 classes
+mtry=9,min.node.size=3; VI is mostly stable!
+###... Random Forest
+```{r}
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_qualitative ~., 
+                    data = mcf_model3, 
+                    #sampsize = c('Not_TC' = 84, 'TC' = 84), 
+                    sampsize = c("NLM" = 84,  "High" = 66, "Type Converted" = 84),
+                    ntree=1500, 
+                    importance=T,
+                    replace = T, 
+                    mtry = 9,
+                    min.node.size = 3)
+# OOB and importance
+rfm
+#importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+
+```
+###... ROC/AUC
+```{r}
+# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
+rfm_roc <-  roc(mcf_model$rdnbr_tc, rfm$votes[,2])
+plot(rfm_roc)
+auc(rfm_roc)
+```
+###... ALE
+```{r}
+
+mcf_ale <- mcf_model3[-which(names(mcf_model3) == "rdnbr_qualitative")]
+
+# predictors 
+nlm_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="NLM")
+high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="High")
+tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="Type Converted")
+
+d = 7
+#ndmi
+nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm1, high1, tc1, nrow=1)
+##ggsave(here::here("Graph Images", "mcf ndmi.jpg"))
+#theme (panel.background = element_rect(width = 5))
+
+#solar radiation
+nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+##ggsave(here::here("Graph Images", "mcf sr.jpg"))
+
+#precip
+nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+##ggsave(here::here("Graph Images", "mcf precip.jpg"))
+
+#evi
+nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf evi.jpg"))
+
+#tmax
+nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf tmax.jpg"))
+
+#tmin jan 2007
+nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf tmin_jan2007.jpg"))
+
+#vpd aug
+nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf vpd_aug.jpg"))
+
+#cwd
+nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf cwd.jpg"))
+
+#slope
+nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf slope.jpg"))
+
+#tslf
+nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf tslf.jpg"))
+
+#twi
+nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf twi.jpg"))
+
+#aspect
+nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf Aspect.jpg"))
+
+ggarrange(nlm1, nlm2,nlm3,nlm4,nlm5,nlm6,nlm7,nlm8,nlm9,nlm10,nlm11,nlm12, nrow = 3, ncol=4)
+ggsave(here::here("Graph Images", "mcf all ulm.jpg"), width=8, height = 8)
+
+ggarrange(high1,high2,high3,high4,high5,high6,high7,high8,high9,high10,high11,high12, nrow=3, ncol=4)
+ggsave(here::here("Graph Images", "mcf all high.jpg"), width=8, height = 8)
+
+ggarrange(tc1,tc2,tc3,tc4,tc5,tc6,tc7,tc8,tc9,tc10,tc11,tc12, nrow=3, ncol=4)
+ggsave(here::here("Graph Images", "mcf all tc.jpg"), width=8, height = 8)
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+# all_rfm_variables <- FeatureEffects$new(tc_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'sr', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'tslf', 'tmin_jan2007'))
+# all_rfm_variables$plot()
+# ggsave("mcf tc.jpg", height = 9, width = 9)
+
+```
+
+###... ALE
+```{r}
+
+mcf_ale <- mcf_model[-which(names(mcf_model) == "rdnbr_tc")]
+
+# predictors 
+noTC_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_tc, type = "prob", class="Not_TC")
+TC_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_tc, type = "prob", class="TC")
+
+#ndvi
+Low <- plot(FeatureEffect$new(noTC_predictor, feature = "vpd_aug")) + ggtitle("Not Type Converted") + labs(x="NDVI")
+high <- plot(FeatureEffect$new(TC_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="NDVI")
+ggarrange(Low, high)
+
+#temp
+Low <- plot(FeatureEffect$new(noTC_predictor, feature = "rain")) + ggtitle("Not Type Converted")+ labs(x="Precipitation")
+high <- plot(FeatureEffect$new(TC_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="Precipitation")
+ggarrange(Low, high)
+
+#precip
+Low <- plot(FeatureEffect$new(noTC_predictor, feature = "evi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Outside")
+high <- plot(FeatureEffect$new(TC_predictor, feature = "evi_2006")) + ggtitle("Type Converted")+ labs(x="Vegetation Outside")
+ggarrange(Low, high)
+
+#precip
+Low <- plot(FeatureEffect$new(noTC_predictor, feature = "ndmi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Inside")
+high <- plot(FeatureEffect$new(TC_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="Vegetation Inside")
+ggarrange(Low, high)
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
+# all_rfm_variables$plot()
+# ggsave("ale4.jpg", height = 9, width = 9)
+```
+interesting graphs for evi and ndmi: high evi (which represents denser veg) is associated with low mortality, which is counterintuative, but higher ndmi (which represents moist veg??) is also associated with low mortality.
+
+## Coulter Pines
+
+#### Load data
 
 ```{r}
+pico <- read_csv("coulter_env_modified.csv") %>% 
+  mutate(rdnbr_qualitative = case_when(
+    rdnbr_2007 < 640 ~ "NLM",
+    rdnbr_2007 < 850 ~ "High",
+    rdnbr_2007 >= 850 ~ "Type Converted")) %>%
+  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
+         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, levels=c("NLM","High", "Type Converted"))) %>% 
+  #filter(Mortality != "bigcone") %>% # 60 stands actually have big cones not pico
+  dplyr::select(-Mortality, -type_converted, -ndvi_2006) %>% 
+  na.omit(pico) %>%  # 56 stands have NA values (probs all very small stands)
+   mutate(rdnbr_qualitative2 = case_when(
+    rdnbr_qualitative %in% "High" ~ "High",
+    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
+  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
+  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, levels=c("NLM", "High"))) %>% 
+  mutate(aspect_transformed = (1-cos((2*pi*aspect_30m)/360))/2) %>% 
+  mutate(rdnbr_tc = case_when(
+    rdnbr_2007 < 850 ~ "Not_TC",
+    rdnbr_2007 >= 850 ~ "TC"
+  )) %>% 
+  mutate(rdnbr_tc = as.factor(rdnbr_tc)) %>% 
+  mutate(rdnbr_tc= fct_relevel(rdnbr_tc, levels=c("Not_TC", "TC"))) %>% 
+  rename_all(funs(str_replace(., "_30m", "")))
 
-#Create random forest using selected variables
-mixed_conifer_rf <- randomForest(RdNBR_2007 ~ var1 + var2 + var3, 
-                                 data = mixed_conifer_env, 
-                                 importance = TRUE, 
-                                 na.action = na.omit)
+pico_tc_samplesize <- pico %>% group_by(rdnbr_tc) %>% dplyr::count() # 
+pico_tc_samplesize <- pico %>% group_by(rdnbr_qualitative) %>% dplyr::count() # nlm = 324, high=120, tc=595
+pico_tc_samplesize <- pico %>% group_by(rdnbr_qualitative) %>% dplyr::summarise(acres = sum(acres)) %>% dplyr::summarise(total_acres = sum(acres))
+mod1 <- lm(dem~tmin_jan2007, data=pico) # R2=0.368
+mod2 <- lm(dem~cwd, data=pico) # R2=0.68
+mod3 <- lm(dem~rain, data=pico) # R2=0.475
+mod4 <- lm(dem~tmax, data=pico) # R2=0.86
+mod5 <- lm(dem~vpd_aug, data=pico) # R2=0.455
+
+pico_elev <- pico %>% dplyr::summarise(elev_mean = mean(dem),
+                                       min_elev = min(dem),
+                                       max_elev =max(dem))
 
 ```
+
+###... Tuning Parameters
+
+source: https://arxiv.org/pdf/1804.03515.pdf (see digital page 9 for tuneRanger)
+
+options for the measure function: https://mlr.mlr-org.com/articles/tutorial/measures.html
+
+```{r}
+set.seed(1127)
+
+# select variables to be used in the model
+pico_model2 <- pico %>% dplyr::select(rdnbr_tc, tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007, ndmi_2006)
+
+# run tuneRanger
+pico_task <- makeClassifTask(data = pico_model2, target = "rdnbr_tc")
+tune_output <- tuneRanger(pico_task, 
+                          measure = list(mmce), 
+                          num.trees = 1500, 
+                          tune.parameters = c("mtry", "min.node.size"))
+
+# recommended parameter settings
+# get best parameter values
+tune_output$model
+
+## other 
+# select variables to be used in the model
+pico_model3 <- pico %>% dplyr::select(rdnbr_qualitative, tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007)
+
+# run tuneRanger
+pico_task <- makeClassifTask(data = pico_model3, target = "rdnbr_qualitative")
+tune_output <- tuneRanger(pico_task, 
+                          measure = list(mmce), 
+                          num.trees = 1500, 
+                          tune.parameters = c("mtry", "min.node.size"))
+
+# recommended parameter settings
+# get best parameter values
+tune_output$model
+
+```
+
+rdnbr = 850: tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007, ndmi_2006; mtry=3,min.node.size=2
+
+3 class: mtry=1,min.node.size=2; VI stable!
+
+###... Random Forest
+```{r}
+
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_qualitative ~., 
+                    data = pico_model3, 
+                    #sampsize = c('Not_TC' = 444, 'TC' = 444), 
+                    sampsize = c("NLM" = 120,  "High" = 120, "Type Converted" = 120),
+                    ntree=1500, 
+                    importance=T,
+                    replace = T, 
+                    mtry = 1,
+                    min.node.size = 2)
+# OOB and importance
+rfm
+#importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+
+#ggplot(pico_model, aes(x=tmin_jan, y=tmin_jan2007)) + geom_point()
+#ggplot(pico_model, aes(x=tmax, y=aug_tmax2006)) + geom_point()
+
+```
+
+###... ALE
+```{r}
+pico_ale <- pico_model3[-which(names(pico_model3) == "rdnbr_qualitative")]
+
+# predictors 
+nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="NLM")
+high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="High")
+tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="Type Converted")
+
+d = 7
+
+#ndmi
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("NLM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("High") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico ndmi.jpg"))
+#theme (panel.background = element_rect(width = 5))
+
+#solar radiation
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "insolation")) + ggtitle("NLM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "insolation")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "insolation")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico sr.jpg"))
+
+#precip
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("NLM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico precip.jpg"))
+
+#tmax
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico tmax.jpg"))
+
+#tmin jan 2007
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico tmin_jan2007.jpg"))
+
+#vpd aug
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico vpd_aug.jpg"))
+
+#cwd
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico cwd.jpg"))
+
+#evi
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico evi.jpg"))
+
+#slope
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico slope.jpg"))
+
+#tslf
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "years_since_burn")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "years_since_burn")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "years_since_burn")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico tslf.jpg"))
+
+#twi
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico twi.jpg"))
+
+#aspect
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "pico Aspect.jpg"))
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
+# all_rfm_variables$plot()
+# ggsave("ale4.jpg", height = 9, width = 9)
+
+```
+
+###... ALE
+```{r}
+
+pico_ale <- pico_model3[-which(names(pico_model3) == "rdnbr_qualitative")]
+
+# predictors 
+nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="NLM")
+high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="High")
+tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="Type Converted")
+
+d = 7
+#ndmi
+nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+##ggsave(here::here("Graph Images", "mcf ndmi.jpg"))
+#theme (panel.background = element_rect(width = 5))
+
+#solar radiation
+nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high2 <- plot(FeatureEffect$new(high_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+##ggsave(here::here("Graph Images", "mcf sr.jpg"))
+
+#precip
+nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+##ggsave(here::here("Graph Images", "mcf precip.jpg"))
+
+#evi
+nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf evi.jpg"))
+
+#tmax
+nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf tmax.jpg"))
+
+#tmin jan 2007
+nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf tmin_jan2007.jpg"))
+
+#vpd aug
+nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf vpd_aug.jpg"))
+
+#cwd
+nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf cwd.jpg"))
+
+#slope
+nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf slope.jpg"))
+
+#tslf
+nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high10 <- plot(FeatureEffect$new(high_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf tslf.jpg"))
+
+#twi
+nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf twi.jpg"))
+
+#aspect
+nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+#ggsave(here::here("Graph Images", "mcf Aspect.jpg"))
+
+ggarrange(nlm1, nlm2,nlm3,nlm4,nlm5,nlm6,nlm7,nlm8,nlm9,nlm10,nlm11,nlm12, nrow = 3, ncol=4)
+ggsave(here::here("Graph Images", "pico all ulm.jpg"), width=8, height = 8)
+
+ggarrange(high1,high2,high3,high4,high5,high6,high7,high8,high9,high10,high11,high12, nrow=3, ncol=4)
+ggsave(here::here("Graph Images", "pico all high.jpg"), width=8, height = 8)
+
+ggarrange(tc1,tc2,tc3,tc4,tc5,tc6,tc7,tc8,tc9,tc10,tc11,tc12, nrow=3, ncol=4)
+ggsave(here::here("Graph Images", "pico all tc.jpg"), width=8, height = 8)
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+# all_rfm_variables <- FeatureEffects$new(tc_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'sr', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'tslf', 'tmin_jan2007'))
+# all_rfm_variables$plot()
+# ggsave("mcf tc.jpg", height = 9, width = 9)
+
+```
+
+###... ALE
+```{r}
+
+pico_ale <- pico_model2[-which(names(pico_model2) == "rdnbr_tc")]
+
+# predictors 
+noTC_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_tc, type = "prob", class="Not_TC")
+TC_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_tc, type = "prob", class="TC")
+
+#ndvi
+Low <- plot(FeatureEffect$new(noTC_predictor, feature = "vpd_aug")) + ggtitle("Not Type Converted") + labs(x="NDVI")
+high <- plot(FeatureEffect$new(TC_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="NDVI")
+ggarrange(Low, high)
+
+#temp
+Low <- plot(FeatureEffect$new(noTC_predictor, feature = "rain")) + ggtitle("Not Type Converted")+ labs(x="Precipitation")
+high <- plot(FeatureEffect$new(TC_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="Precipitation")
+ggarrange(Low, high)
+
+#precip
+Low <- plot(FeatureEffect$new(noTC_predictor, feature = "evi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Outside")
+high <- plot(FeatureEffect$new(TC_predictor, feature = "evi_2006")) + ggtitle("Type Converted")+ labs(x="Vegetation Outside")
+ggarrange(Low, high)
+
+#precip
+Low <- plot(FeatureEffect$new(noTC_predictor, feature = "ndmi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Inside")
+high <- plot(FeatureEffect$new(TC_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="Vegetation Inside")
+ggarrange(Low, high)
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+all_rfm_variables <- FeatureEffects$new(noTC_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
+all_rfm_variables$plot()
+ggsave("pico_no_tc2.jpg", height = 9, width = 9)
+```

### 2020-10-07 Model Data Sheets and Code/Manuscript_Code.Rmd
tidy up code

@@
 ## Mixed Conifer Forest
-
-###... Extract Values
-
-```{r}
-
-mixed_conifer_env <- extract()
-
-```
-
-###... Stepwise Variable Selection
-
-```{r}
-
-#Extract prediction and response variables
-mcf.predictors <- mcf_model[,2:ncol(mcf_model)]
-mcf.response <- mcf_model[,1]
-
-#Run stepwise variable selection algorithm
-mcf.vsurf <- VSURF(x = mcf.predictors, y = mcf.response)
-
-#Get names of selected variables
-names(mcf.predictors[mcf.vsurf$varselect.thres])
-names(mcf.predictors[mcf.vsurf$varselect.interp])
-names(mcf.predictors[mcf.vsurf$varselect.pred])
-
-```
-prediction step, variables selected:all
-
-interp step: variables se;ected: "vpd_aug"         "slope"           "rain"            "tmax"            "tmin_jan"        "sr" "aug_tmax2006"    "tmin_jan2007"    "precip_2006"     "aug_max_vpd2006" "ndmi_2006"       "evi_2006"  "tslf"
-
-prediction step, variables selected: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "sr"           "aug_tmax2006"  "tmin_jan2007" "ndmi_2006"    "evi_2006"     "tslf"
-
-so climate variables are performing better than the weather variables (maybe bc the climate data is avail in higher resolution so there is more variability in values?)
-
-removed aug_max_vpd2006 and precip_2006 and reran VSURF
-
-interp step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "tmin_jan2007"
-
-predicition step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan2007"
 
@@
 
-mcf_tc_samplesize <- mcf %>% group_by(tc_mortality) %>%count() #type converted n = 99
-mcf_tc_samplesie <- mcf %>% group_by(rdnbr_tc) %>%dplyr::count() # type converted n = 55 (almost too small to use)
-mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::count() # high=66, low= 585, moderate=272, type converted n = 84 
+```
+
+### ... summary statistics
+```{r}
+# sample sizes
+mcf_tc_samplesize <- mcf %>% group_by(tc_mortality) %>%count() #type converted manual estimates n = 99
+mcf_tc_samplesie <- mcf %>% group_by(rdnbr_tc) %>%dplyr::count() # rdnbr=1000 for type converted  n = 55 (almost too small to use)
+mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::count() #rdnbr = 850:  high=66, low= 585, moderate=272, type converted n = 84
+
+# acres by burn severity class an dtotal acres
 mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::summarise(acres = sum(Shape_Area/4047)) %>%dplyr::summarise(tot_acres = sum(acres))
 
+# look at correlations/trends in potentially correlated variables
 ggplot(mcf, aes(x = rdnbr_tc, y = rdnbr_2007)) +geom_jitter(aes(color=tc_mortality))
@@
 
+# summary info on TSLF adn elevation for the manuscript
 mcf %>% dplyr::summarise(min = min(tslf),
@@
 ```
+surprisingly, far fewer variables are strongly correlated with each other and elevation than I would've have thought.
+
+###... Extract Values
+
+```{r}
+
+mixed_conifer_env <- extract()
+
+```
+
+###... Stepwise Variable Selection (VSURF)
+
+```{r}
+
+#Extract prediction and response variables
+mcf.predictors <- mcf_model[,2:ncol(mcf_model)]
+mcf.response <- mcf_model[,1]
+
+#Run stepwise variable selection algorithm
+mcf.vsurf <- VSURF(x = mcf.predictors, y = mcf.response)
+
+#Get names of selected variables
+names(mcf.predictors[mcf.vsurf$varselect.thres])
+names(mcf.predictors[mcf.vsurf$varselect.interp])
+names(mcf.predictors[mcf.vsurf$varselect.pred])
+
+```
+run 1 results: 
+prediction step, variables selected:all
+
+interp step: variables se;ected: "vpd_aug"         "slope"           "rain"            "tmax"            "tmin_jan"        "sr" "aug_tmax2006"    "tmin_jan2007"    "precip_2006"     "aug_max_vpd2006" "ndmi_2006"       "evi_2006"  "tslf"
+
+prediction step, variables selected: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "sr"           "aug_tmax2006"  "tmin_jan2007" "ndmi_2006"    "evi_2006"     "tslf"
+
+so climate variables are performing better than the weather variables (maybe bc the climate data is avail in higher resolution so there is more variability in values?)
+
+*removed aug_max_vpd2006 and precip_2006 and reran VSURF*
+
+run 2 results: 
+interp step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "tmin_jan2007"
+
+predicition step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan2007"
 
@@
 # select variables to be used in the model 
-mcf_model <- mcf %>% 
-  dplyr::select(rdnbr_tc, tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, tslf) # not sure if fri and years so fire, whic sounds like Time Since Last Fire, are the same thing
-
-# run tuneRanger
-mcf_task <- makeClassifTask(data = mcf_model, target = "rdnbr_tc")
-tune_output <- tuneRanger(mcf_task, 
-                          measure = list(mmce), 
-                          num.trees = 1500, 
-                          tune.parameters = c("mtry", "min.node.size"))
-
-# recommended parameter settings
-# get best parameter values
-tune_output$model
-
-# 4 rdnbr classes
 mcf_model3 <- mcf %>% 
-  dplyr::select(rdnbr_qualitative, tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, tslf) # not sure if fri and years so fire, whic sounds like Time Since Last Fire, are the same thing
+  dplyr::select(rdnbr_qualitative, tmin_jan2007, vpd_aug, cwd, tmax, rain, sr, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, tslf) # not sure if fri and years so fire, whic sounds like Time Since Last Fire, are the same thing
 
@@
 mtry=9,min.node.size=3; VI is mostly stable!
+
 ###... Random Forest
@@
 ```
+
 ###... ROC/AUC
@@
 ```
+
 ###... ALE
 ```{r}
-
 mcf_ale <- mcf_model3[-which(names(mcf_model3) == "rdnbr_qualitative")]
@@
 # predictors 
-nlm_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="NLM")
+nlm_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf$rdnbr_qualitative, type = "prob", class="NLM")
+high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf$rdnbr_qualitative, type = "prob", class="High")
+tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf$rdnbr_qualitative, type = "prob", class="Type Converted")
+
+d = 7
+
+#ndmi
+nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("High") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm1, high1, tc1, nrow=1)
+ggsave(here::here("Graph Images", "mcf ndmi.jpg"))
+
+#solar radiation
+nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm2, high2, tc2, nrow=1)
+ggsave(here::here("Graph Images", "mcf sr.jpg"))
+
+#precip
+nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm3, high3, tc3, nrow=1)
+ggsave(here::here("Graph Images", "mcf precip.jpg"))
+
+#tmax
+nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high4 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm4, high4, tc4, nrow=1)
+ggsave(here::here("Graph Images", "mcf tmax.jpg"))
+
+#tmin jan 2007
+nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm5, high5, tc5, nrow=1)
+ggsave(here::here("Graph Images", "mcf tmin_jan2007.jpg"))
+
+#vpd aug
+nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high6 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm6, high6, tc6, nrow=1)
+ggsave(here::here("Graph Images", "mcf vpd_aug.jpg"))
+
+#cwd
+nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high7 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm7, high7, tc7, nrow=1)
+ggsave(here::here("Graph Images", "mcf cwd.jpg"))
+
+#evi
+nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high8 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm8, high8, tc8, nrow=1)
+ggsave(here::here("Graph Images", "mcf evi.jpg"))
+
+#slope
+nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm9, high9, tc9, nrow=1)
+ggsave(here::here("Graph Images", "mcf slope.jpg"))
+
+#tslf
+nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm10, high10, tc10, nrow=1)
+ggsave(here::here("Graph Images", "mcf tslf.jpg"))
+
+#twi
+nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm11, high11, tc11, nrow=1)
+ggsave(here::here("Graph Images", "mcf twi.jpg"))
+
+#aspect
+nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm12, high12, tc12, nrow=1)
+ggsave(here::here("Graph Images", "mcf Aspect.jpg"))
+
+ggarrange(nlm1, high1, tc1, tc1,nlm2, high2, tc2,tc1,nlm3, high3, tc3,tc1, nrow =3, ncol =4)
+ggsave(here::here("Graph Images", "mcf ndm_sr_precip6.jpg"))
+
+#ggarrange(nlm5, high5, tc5, nlm6, high6, tc6, nlm7, high7, tc7)
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
+# all_rfm_variables$plot()
+# ggsave("ale4.jpg", height = 9, width = 9)
+
+```
+
+###... ALE
+```{r}
+
+mcf_ale <- mcf_model3[-which(names(mcf_model3) == "rdnbr_qualitative")]
+
+# predictors 
+nlm_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="ULM")
 high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="High")
@@
 #ndmi
-nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + themet.margin = unit(c(d,1,1,1), "lines"))
+tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + themet.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm1, high1, tc1, nrow=1)
@@
 #vpd aug
-nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-ggarrange(nlm, high, tc, nrow=1)
+nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))# + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))# + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm7, high7, tc7, nrow=1)
 #ggsave(here::here("Graph Images", "mcf vpd_aug.jpg"))
@@
 ggarrange(nlm1, nlm2,nlm3,nlm4,nlm5,nlm6,nlm7,nlm8,nlm9,nlm10,nlm11,nlm12, nrow = 3, ncol=4)
-ggsave(here::here("Graph Images", "mcf all ulm.jpg"), width=8, height = 8)
+ggsave(here::here("Graph Images", "mcf all ulm_no sampsize.jpg"), width=8, height = 8)
 
 ggarrange(high1,high2,high3,high4,high5,high6,high7,high8,high9,high10,high11,high12, nrow=3, ncol=4)
-ggsave(here::here("Graph Images", "mcf all high.jpg"), width=8, height = 8)
+#ggsave(here::here("Graph Images", "mcf all high_no sampsize.jpg"), width=8, height = 8)
 
 ggarrange(tc1,tc2,tc3,tc4,tc5,tc6,tc7,tc8,tc9,tc10,tc11,tc12, nrow=3, ncol=4)
-ggsave(here::here("Graph Images", "mcf all tc.jpg"), width=8, height = 8)
+ggsave(here::here("Graph Images", "mcf all tc_no sampsize.jpg"), width=8, height = 8)
 
@@
   mutate(rdnbr_qualitative = case_when(
-    rdnbr_2007 < 640 ~ "NLM",
+    rdnbr_2007 <= 640 ~ "NLM",
     rdnbr_2007 < 850 ~ "High",
@@
 # predictors 
-nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="NLM")
-high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="High")
-tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="Type Converted")
+nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico$rdnbr_qualitative, type = "prob", class="NLM")
+high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico$rdnbr_qualitative, type = "prob", class="High")
+tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico$rdnbr_qualitative, type = "prob", class="Type Converted")
 
@@
 #ndmi
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("NLM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("High") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #solar radiation
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "insolation")) + ggtitle("NLM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "insolation")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "insolation")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #precip
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("NLM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #tmax
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #tmin jan 2007
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #vpd aug
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #cwd
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #evi
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #slope
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #tslf
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "years_since_burn")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "years_since_burn")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "years_since_burn")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #twi
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 #aspect
-nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("NLM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 high <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))

### 2020-10-08 Model Data Sheets and Code/Manuscript_Code.Rmd
minor change. tried running vsurf for pico data and got an error

@@
 
+###... Stepwise Variable Selection (VSURF)
+
+```{r}
+
+#Extract prediction and response variables
+pico.predictors <- pico_model[,2:ncol(pico_model)]
+pico.response <- pico_model[,1]
+
+#Run stepwise variable selection algorithm
+pico.vsurf <- VSURF(x = pico.predictors, y = pico.response)
+
+#Get names of selected variables
+names(pico.predictors[pico.vsurf$varselect.thres])
+names(pico.predictors[pico.vsurf$varselect.interp])
+names(pico.predictors[pico$varselect.pred])
+
+```
+get error for some reason, but not with the MCF df :(
+
 ###... Tuning Parameters
@@
 # select variables to be used in the model
-pico_model2 <- pico %>% dplyr::select(rdnbr_tc, tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007, ndmi_2006)
+pico_model <- pico %>% dplyr::select(rdnbr_qualitative, tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007, ndmi_2006)
 

### 2020-10-09 Model Data Sheets and Code/Manuscript_Code.Rmd
work on combined model

@@
 #solar radiation
-nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm2, high2, tc2, nrow=1)
@@
 #precip
-nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm3, high3, tc3, nrow=1)
@@
 #tmax
-nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high4 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-ggarrange(nlm4, high4, tc4, nrow=1)
-ggsave(here::here("Graph Images", "mcf tmax.jpg"))
 
 #tmin jan 2007
-nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm5, high5, tc5, nrow=1)
@@
 #vpd aug
-nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high6 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high6 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm6, high6, tc6, nrow=1)
@@
 #cwd
-nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high7 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high7 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm7, high7, tc7, nrow=1)
@@
 #evi
-nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high8 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high8 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm8, high8, tc8, nrow=1)
@@
 #slope
-nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm9, high9, tc9, nrow=1)
@@
 #tslf
-nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm10, high10, tc10, nrow=1)
@@
 #twi
-nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm11, high11, tc11, nrow=1)
@@
 #aspect
-nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) # + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))  + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))  + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))  + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm12, high12, tc12, nrow=1)
@@
 nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + themet.margin = unit(c(d,1,1,1), "lines"))
-tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + themet.margin = unit(c(d,1,1,1), "lines"))
+high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm1, high1, tc1, nrow=1)
@@
 #solar radiation
-nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #precip
-nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #evi
-nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #tmax
-nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #tmin jan 2007
-nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #vpd aug
-nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))# + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))# + theme(plot.margin = unit(c(d,1,1,1), "lines"))
-tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm7, high7, tc7, nrow=1)
@@
 #cwd
-nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #slope
-nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #tslf
-nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #twi
-nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #aspect
-nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 
+pico %>% group_by(rdnbr_qualitative) %>% dplyr::count()
+
 pico_tc_samplesize <- pico %>% group_by(rdnbr_tc) %>% dplyr::count() # 
@@
 # predictors 
-nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico$rdnbr_qualitative, type = "prob", class="NLM")
-high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico$rdnbr_qualitative, type = "prob", class="High")
-tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico$rdnbr_qualitative, type = "prob", class="Type Converted")
+nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_model3$rdnbr_qualitative, type = "prob", class="NLM")
+high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_model3$rdnbr_qualitative, type = "prob", class="High")
+tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_model3$rdnbr_qualitative, type = "prob", class="Type Converted")
 
@@
 nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
-high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #solar radiation
-nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high2 <- plot(FeatureEffect$new(high_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high2 <- plot(FeatureEffect$new(high_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #precip
-nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #evi
-nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #tmax
-nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #tmin jan 2007
-nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #vpd aug
-nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #cwd
-nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #slope
-nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #tslf
-nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high10 <- plot(FeatureEffect$new(high_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high10 <- plot(FeatureEffect$new(high_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #twi
-nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 ggarrange(nlm, high, tc, nrow=1)
@@
 #aspect
-nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
-high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ themet.margin = unit(c(d,1,1,1), "lines"))
+nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
 tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
@@
 ```
+
+#COMBINED COulter and MCF data
+
+##... prep data
+```{r}
+# rename some variables to match with MCF
+pico_mod <- pico_model3 %>% rename(sr = insolation, tslf = years_since_burn)
+
+# combine mcf and pico
+
+all_data <- rbind(pico_mod, mcf_model3)
+
+# samp size
+all_data %>%  group_by(rdnbr_qualitative) %>% dplyr::count()
+
+```
+
+###... Tuning Parameters
+
+source: https://arxiv.org/pdf/1804.03515.pdf (see digital page 9 for tuneRanger)
+
+options for the measure function: https://mlr.mlr-org.com/articles/tutorial/measures.html
+
+```{r}
+set.seed(1127)
+
+# run tuneRanger
+all_task <- makeClassifTask(data = all_data, target = "rdnbr_qualitative")
+tune_output <- tuneRanger(all_task, 
+                          measure = list(mmce), 
+                          num.trees = 1500, 
+                          tune.parameters = c("mtry", "min.node.size"))
+
+# recommended parameter settings
+tune_output$model
+
+```
+mtry=3,min.node.size=3
+
+###... Random Forest
+```{r}
+
+for (i in 1:10) {
+
+rfm <- randomForest(rdnbr_qualitative ~., 
+                    data = all_data, 
+                    sampsize = c("NLM" = 186,  "High" = 186, "Type Converted" = 186),
+                    ntree=1500, 
+                    importance=T,
+                    replace = T, 
+                    mtry = 3,
+                    min.node.size = 3)
+# OOB and importance
+rfm
+#importance(rfm)
+#plot to visualize variable importance
+varImpPlot(rfm)
+
+}
+
+```
+
+###... ALE
+```{r}
+all_ale <- all_data[-which(names(all_data) == "rdnbr_qualitative")]
+
+# predictors 
+nlm_predictor <- Predictor$new(rfm, data = all_ale, y = all_data$rdnbr_qualitative, type = "prob", class="NLM")
+high_predictor <- Predictor$new(rfm, data = all_ale, y = all_data$rdnbr_qualitative, type = "prob", class="High")
+tc_predictor <- Predictor$new(rfm, data = all_ale, y = all_data$rdnbr_qualitative, type = "prob", class="Type Converted")
+
+d = 7
+
+#ndmi
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("High") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all ndmi.jpg"))
+#theme (panel.background = element_rect(width = 5))
+
+#solar radiation
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all sr.jpg"))
+
+#precip
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all precip.jpg"))
+
+#tmax
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.21)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High") + labs(x="Max August Temp") + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all tmax.jpg"))
+
+#tmin jan 2007
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all tmin_jan2007.jpg"))
+
+#vpd aug
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all vpd_aug.jpg"))
+
+#cwd
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all cwd.jpg"))
+
+#evi
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all evi.jpg"))
+
+#slope
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all slope.jpg"))
+
+#tslf
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all tslf.jpg"))
+
+#twi
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all twi.jpg"))
+
+#aspect
+nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+high <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+tc <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
+ggarrange(nlm, high, tc, nrow=1)
+ggsave(here::here("Graph Images", "all Aspect.jpg"))
+
+# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
+# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
+# all_rfm_variables$plot()
+# ggsave("ale4.jpg", height = 9, width = 9)
+
+```
